### 2022-02-26 manuscript_covid_transparency.Rmd
First commit

---
title             : "Transparency of COVID-19-related research in medical journals"
shorttitle        : "Transparency of COVID-19-related research"

author: 
    - name        : "Eero Raittio"
    affiliation   : "1"
    corresponding : no    # Define only one corresponding author
    email         : "eero.raittio@uef.fi"
  - name          : "Ahmad Sofi-Mahmudi"
    affiliation   : c("2", "3")
    corresponding : yes    # Define only one corresponding author
    address       : "41, Shanaz St., Seqiz Health Network, Seqiz, Kurdistan, Iran."
    email         : "a.sofimahmudi@gmail.com"

affiliation:
  - id            : "1"
    institution   : "Institute of Dentistry, University of Eastern Finland, Kuopio, Finland."
  - id            : "2"
    institution   : "Seqiz Health Network, Kurdistan University of Medical Sciences, Seqiz, Iran."
  - id            : "3"
    institution   : "Cochrane Iran Associate Centre, National Institute for Medical Research Development (NIMAD), Tehran, Iran."

note: "\\clearpage"

authornote:

abstract:

keywords          : "Hypothesis Testing"
wordcount         : "2,860"

bibliography      : ["references.bib", "r-references.bib"]

floatsintext      : yes #place figures and tables in the text rather than at the end
numbersections    : no #number sections headings
figurelist        : no #create list of figure captions
tablelist         : no #create list of table captions
footnotelist      : no ##create list of footnotes
linenumbers       : no #add line numbers in the margins
mask              : no #omit identifying information from the title page
draft             : no #add "draft" watermark across all pages

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_word
---

```{r setup, include = FALSE}
library(ggplot2)
library(papaja)
library(citr)
library(rpact)
library(dplyr)
r_refs("r-references.bib")
```

"Whatever else is done about null-hypothesis tests, let us stop viewing statistical analysis as a sanctification process. We are awash in a sea of uncertainty, caused by a flood tide of sampling and measurement errors, and there are no objective procedures that avoid human judgment and guarantee correct interpretations of results." [@abelson1997: p.13]

Psychological science is experiencing a period of widespread methodological reflection [@nelson2018]. Starting 2010, a chain of events caused psychological scientists to doubt the credibility of psychological research [@moshontz2018]. These events included a case of widespread academic fraud, a series of failed replications of famous findings, and a paper published in psychology's most prestigious journal claiming that extrasensory perception (ESP) is real [@nelson2018].

According to some, one of the root causes of the crisis of confidence is the misunderstanding and misuse of statistical methods [@goodman2019]. In psychological science, one of the most frequently used approaches to statistical inference is the null-hypothesis significance test (NHST). For decades, researchers have been warned about mindless applications of this tool [e.g., @cohen1994; @gigerenzer2004; @nickerson2000; @rozeboom1960].

It has been argued that, in practice, the NHST has taken on a mechanical nature. According to Gigerenzer et al. [-@gigerenzer1989, p.210-227], scientists have interpreted the NHST as a completely mechanized procedure to arrive at new knowledge while eliminating the need for personal judgment and promising objectivity. A normative view has emerged; statistics can speak with one voice and replace personal judgment by means of formal rules.

Among statisticians, there seems to be widespread agreement about what not do with hypothesis tests and *p*-values. Unfortunately, there is less agreement about what researchers should be doing instead [@wasserstein2019]. Crucially, although the founders of statistics disagreed on many things, they seemed to agree on at least one thing: they all rejected mechanical inference and highlighted the importance of human judgment [@abelson1995; @gigerenzer2015; @mayo2018].

Unfortunately, the literature on NHST seems to have lost sight of the important distinction between the tool itself and its thoughtless application [@mayo2018; @lakens2020a]. It seems the tool has become synonymous with its weakest construal: setting up a nil-null hypothesis and using a single *p*-value below a conventional threshold to infer evidence for a substantive research claim [@mayo2018; @gelman2016]. As a consequence, numerous scholars have called for banning the tool altogether.

It is important to emphasize, however, that NHST has a place in scientific inference [@abelson1997; @lakens2020a] and that it is likely to stay [@goodman2019]. Rather than invalidating the tool on the basis of its weakest construal, an important step forward is to highlight how NHSTs can be thoughtfully applied in practice [@wasserstein2019].

In this paper, I examine the way in which hypothesis tests are conducted in psychological science. I focus on three decisions central to hypothesis testing: (a) the null hypothesis to be tested, (b) the choice of acceptable error rates, and (c) the determination of sample size. I refer to conventional uses of NHST (i.e., choices that are widely shared across the literature, are made without justification, and/or refer to heuristics or conventions) as *defaults*. I refer to thoughtful uses of NHST (i.e., choices that are actively justified) as *active choices*. In addition to examining common practice, I provide a practical overview of active choices that can be made within the traditional framework to strengthen the practice of null-hypothesis significance testing.

In the first section, I introduce default hypothesis tests and their thoughtful alternatives. In the second section, I conduct a literature review to examine the ways in which hypothesis tests are practiced in the psychological research literature. My goal is to assess the choices psychological scientists make when testing a hypothesis, the extent to which justifications are provided for these choices, and the extent to which choices are shared across the research literature (i.e., the prevalence of conventions). Of special interest are the choice of the null hypothesis (i.e., whether to test a nil-null or, for example, a range prediction), the choices of error rates (i.e., how the balance is struck between type I and type II error rates), and the determination of sample size.

In the third section, I provide an in-depth, practical introduction to active choice hypothesis tests: a selection of thoughtful choices that can be made within the traditional framework to strengthen the practice of null-hypothesis significance testing. First, I consider the choice of null hypothesis. As an alternative to testing nil-null hypotheses, I introduce several approaches researchers can take to test range predictions.

Second, I consider the choice of error rates. As an alternative to using conventional error thresholds, I discuss several approaches that researchers can take to choose and justify their error rates. Third, I consider sample size determination. As an alternative to using heuristics and unrealistic power estimates (based on effect sizes researchers deem plausible), I discuss several approaches that researchers can take to justify their sample size. In addition, I present a tool that allows researchers to efficiently perform well-powered experiments: sequential analysis.

# Default vs. Active Choice Hypothesis Tests

"Knowing what not to do with *p*-values is indeed necessary, but it does not suffice. It is as though statisticians were asking users of statistics to tear out the beams and struts holding up the edifice of modern scientific research without offering solid construction materials to replace them. Pointing out old, rotting timbers was a good start, but now we need more." [@wasserstein2019: p. 1]

## On the choice of the null hypothesis
"This type of problem... is the basis of the tests of significance by which we can examine whether or not the data are in harmony with any suggested hypothesis." [@fisher1925: p. 8]

The meaning of 'null' hypothesis is commonly misunderstood [@bakan1967]. In the weakest construal of the NHST, the null hypothesis tested is a hypothesis of no difference [i.e., an effect or difference of exactly zero; @mayo2018]; this null is commonly referred to as the *nil-null* or the *straw-man* null hypothesis. However, the term null was originally used to refer to any hypothesis to be 'nullified' [@bakan1967; @mayo2018].

The null can be any hypothesis a researcher wishes to test. According to Neyman [-@neyman1957: p.10], envisioning possible hypotheses that are relevant to the phenomenon of interest is the work of creative imagination. Many of the problems associated with NHST lessen if the null hypothesis is defined as something other than a hypothesized difference, effect, or correlation of precisely zero [@nickerson2000].

A stronger construal of the NHST is one where the researcher tests a range null hypothesis rather than a nil-null hypothesis [@nickerson2000; @lakens2020a]: a move, referred to in the title, from a straw man to a steel woman. Instead of testing the nil-null hypothesis that the effect is exactly zero, researchers can test whether effects more extreme than a range of predicted or practically important effect sizes can be rejected. Examples of such tests are equivalence tests, minimum-effect tests, and inferiority tests [@lakens2020a; @lakens2018]. In the third section, I will give a practical introduction to the use of range predictions in hypothesis tests.

## On the choice of error rates

"[N]o scientific worker has a fixed level of significance at which from year to year, and in all circumstances, he rejects hypotheses; he rather gives his mind to each particular case in the light of his evidence and his ideas." [@fisher1956: p. 44-45]

"From the point of view of mathematical theory all that we can do is to show how the risk of the errors may be controlled and minimized. The use of these statistical tools in any given case, in determining just how the balance should be struck, must be left to the investigator." [@neyman1933: p.293]

Prior to conducting a Neyman-Pearson hypothesis test, researchers need to decide on a type I error rate (i.e., the long-run rate at which a true null hypothesis is rejected) and a type II error rate (i.e., given a number of participants, the long-run probability of failing to reject the null for a specified effect size of interest). In effect, a researcher needs to decide on the importance of exercising scientific caution (i.e., avoiding a type I error) versus the importance of discovery of an effect (i.e., avoiding a type II error).

However, in practice, the sense of balancing the two types of error has been lost [@oakes1986; @rozeboom1960]. More commonly, researchers tend to use conventional cut-offs for the two error rates ($\alpha$ =.05 and $\beta$ = .20, implying that type I errors are considered 4 times more costly than type II errors). In recent work, researchers have been urged to justify their $\alpha$ [@lakens2018]. In the third section, I will elaborate on several approaches to the choice and justification of error rates.

## On the choice of sample size

"[I]t is an essential characteristic of experimentation that it is carried out with limited resources, and an essential part of the subject of experimental design to ascertain how these should be best applied." [@fisher1935: p. 18]

As early as 1962, the perils of conducting small-sample, low-power experiments were introduced to the psychological literature [@cohen1962]. Unfortunately, more recent work has not been able to paint a rosier picture of the statistical power of hypothesis tests in psychological science [@bakker2012; @morey2016].

A possible reason that hypothesis tests often have low chances of detecting meaningful effects is that researchers use heuristics (e.g., *N* = 50 per cell) or effect size estimates from previous studies (which may very well be inflated) when determining sample size [@lakens2017]. Alternative approaches that have been suggested are safeguard power analysis [@perugini2014], expected power analysis [@biesanz2017], and powering the test to detect the smallest effect size of interest [@lakens2017].

Researchers are often surprised to find out how many observations are typically required for a test to have appropriate power. Sequential analysis, on the other hand, allows researchers to perform well-powered experiments with smaller sample sizes [@wald1973]. Thus, sequential analysis has the potential to significantly improve the practice of hypothesis testing in psychological science [@beffarabret2018; @lakens2014; @schnuerch2020]. In the following section, I give a hands-on introduction to sequential hypothesis testing.

## *Sequential Analysis*

In 1929, Harold Dodge and Harry Romig introduced the idea of sequential analysis to the science of statistical quality control [@dodge1929]. Sequential analysis is a method of statistical inference that allows data to be analyzed several times during its collection, with the goal of optimizing the amount of resources spent in research [@beffarabret2018]. When Abraham Wald further developed this idea in the '40s, his work was deemed so useful to the United States military that it was classified as Restricted under the Espionage act [@wald1973].

```{r, include = FALSE}
gs = function(proc, return = "n") {
  bs = ifelse(proc == "asOF", "bsOF", "bsP")

  # Specify the design
  design = getDesignGroupSequential(sided = 1,
                                    alpha = 0.05, 
                                    beta = 0.2,
                                    kMax = 3, 
                                    typeOfDesign = proc,
                                    typeBetaSpending= bs)

  # Get parameters
  parameters = getSampleSizeMeans(design = design, groups = 2, alternative = 0.5)
  n_gs = ceiling(parameters$maxNumberOfSubjects/3/2) # n per look per group
  alpha = parameters$criticalValuesPValueScale
  futility = parameters$futilityBoundsPValueScale %>% round(2)

  if(return == "alpha") {return(alpha %>% signif(2) %>% as.character() %>% sub(".", "", .,))}
  if(return == "n") {return(n_gs)}
  if(return == "futility") {return(futility %>% as.character() %>% sub(".", "", .,))}
}

n_p = gs(proc = "asP")
alpha_p = gs(proc = "asP", return = "alpha")
futility_p = gs(proc = "asP", return = "futility")
n_of = gs(proc = "asOF")
alpha_of = gs(proc = "asOF", return = "alpha")
futility_of = gs(proc = "asOF", return = "futility")
```

The most common procedure for sequential hypothesis testing is a group-sequential (GS) design [@proschan2006; @schonbrodt2017]. The goal of GS procedures is to efficiently reject a null hypothesis in favor of an alternative hypothesis. In a prototypical group-sequential procedure [e.g., the Pocock correction or the O'Brien-Fleming correction, @pocock1977; @fleming1984], a researcher is allowed to look at the data several times during the course of a study.

It is well-known that multiple looks at data, conventionally, increase the type I error rate [@simmons2011]: i.e., the long-run rate at which a true null hypothesis is rejected. Sequential analysis procedures, however, allow a researcher to take multiple looks at data without inflating the type I error rate. Critical $\alpha$ levels are adjusted for each look so that the overall type I error rate is controlled at the desired level (for instance, at an $\alpha$ of 5%).

A general, sophisticated approach is provided by an "$\alpha$*-spending function*" [@lan1983; @wassmer2016]. This approach allows a researcher to have unequally spaced looks at their data (for instance, after the first 20% and the first 75% of the maximum sample size) and to adapt the number of looks as the experiment continues.

For example, a researcher who wishes to control their overall type I error rate at 5% and take *k* = 3 looks (i.e., two interim looks and one final look) at the data can use the common Pocock-like $\alpha$-spending function to determine what the corrected $\alpha$ level should be for each look (referred to as $\alpha_{1,k}$). In this case, $\alpha_{1,1:2}$ = `r alpha_p[1]` for the first two looks and $\alpha_{1,3}$ = `r alpha_p[3]` for the last look (i.e., for a one-tailed test). If during any of the interim looks $p \le \alpha_1$, the researcher can reject the null hypothesis and end the experiment; this is known as "*stopping for efficacy*".

```{r, include = FALSE}
n_fixed = ceiling(power.t.test(delta = 0.5, sig.level = .05, power = 0.8, 
                                 type = "two.sample", alternative = "one.sided")$n)
```

Because the use of group-sequential procedures allows for early stopping, it reduces the average sample size needed. For example, if a researcher wants to conduct a one-sided, two-sample *t* test ($H_0:~\delta$ = 0 against $H_+:~\delta > 0$) with 80% power to detect a Cohen's *d* (i.e., a standardized mean difference) of 0.5, a conventional Null-Hypothesis Significance Test with an overall $\alpha$ of 5% would require `r n_fixed` participants per group. Using the Pocock-like GS design outlined above, the same study can be conducted with only `r n_p` participants per group per look.

In addition to stopping early for efficacy, a researcher can "*stop for futility*" if the observed effect size is too small to warrant further investigation. This can be done by manually specifying a lower bound (for example, a lower bound at *d* $\le$ 0 or, equivalently, at *p* $\ge$ .50) or by using a "$\beta$*-spending function*" to determine the critical lower bound for each look (referred to as $\alpha_{0,k}$).

The logic of the $\beta$-spending function is similar to that of the $\alpha$-spending function: critical values that warrant stopping for futility are determined a priori while controlling the overall type II error rate at the desired level (for instance, at a $\beta$ of 20%). In other words, the bounds are determined such that the probability of a type II error (i.e., given the number of subjects, the long-run probability of failing to reject the null when the true population effect $\delta$ = 0.5) is $\beta$ %.

Figures 1A and 1B provide a visual overview of two possible group-sequential procedures (the Pocock-like and the O'Brien-Fleming-like designs) using $\alpha$- and $\beta$-spending functions, in the case of a one-tailed, two-sample *t*-test with $\alpha$ = 5%, $\beta$ = 20% for $\delta$ = 0.5, and $k_{max}$ = 3 looks. In our example of a Pocock-like design, the futility bounds are $\alpha_{0,1}$ = `r futility_p[1]` for the first look and $\alpha_{0,2}$ = `r futility_p[2]` for the second look. The researcher continues their study until the observed *p*-value crosses one of the prespecified $\alpha_1$ or $\alpha_0$ thresholds, or until the maximum *N* has been reached.

Scientific experiments can require substantial time, money, and effort. At times, the well-being of research subjects, human or non-human, is at stake [@stefan2020]. Under the Ethics Code of the American Psychological Association, psychological scientists have the ethical obligation to minimize research subjects' exposure to physical, emotional, or psychological harm [@apa2020]. Sequential analysis can contribute to this goal, by allowing researchers to stop an experiment when the interim data show clear evidence that effects are positive, negative, or largely absent [@proschan2006]. Sequential hypothesis testing procedures, it turns out, are highly valuable tools for psychological scientists to add to their statistical toolbox.

\newpage

# In-line code example

```{r}
n_fixed = ceiling(power.t.test(delta = 0.5, sig.level = .05, power = 0.8, 
                                 type = "two.sample", alternative = "one.sided")$n)
```

For example, if a researcher wants to conduct a one-sided, two-sample *t* test ($H_0:~\delta$ = 0 against $H_+:~\delta > 0$) with 80% power to detect a Cohen's *d* (i.e., a standardized mean difference) of 0.5, a conventional Null-Hypothesis Significance Test with an overall $\alpha$ of 5% would require `r n_fixed` participants per group.

\newpage

# Example Table

```{r}
df = read.csv("MARP_data.csv")
```

```{r}
mod1 = lm(wb_overall_mean ~ religiosity + age + gender + ses + education, data = df)
apa_lm = apa_print(mod1)
apa_table(apa_lm$table, caption = "Well-being regression table.")
```

The estimate for religiosity is `r apa_lm$estimate$religiosity`

\newpage

## Acknowledgement
This manuscript was created using `r cite_r("r-references.bib")`.

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup

\newpage

## Figure 1A
## *Pocock-like Group-Sequential Design: A worked example* 
\noindent
```{r, fig.heigth = 1, fig.width = 12}
knitr::include_graphics("Figure1A.png", dpi = 1200)
```

*Note.* Pocock-like group-sequential design for a one-tailed, two-sample *t* test with $\alpha$ = .05, $\beta$ = .20 and $k_{max}$ = 3 looks. Critical values are determined on the basis of Pocock-like $\alpha$- and $\beta$-spending functions in *R* package *rpact* (Wassmer & Pahlke, 2020).

\newpage

## Figure 1B
## *O'Brien-Fleming-like Group-Sequential Design: A worked example* 
\noindent
```{r, fig.heigth = 1, fig.width = 12}
knitr::include_graphics("Figure1B.png", dpi = 1200)
```

*Note.* O'Brien-Fleming-like group-sequential design for a one-tailed, two-sample *t* test with $\alpha$ = .05, $\beta$ = .20 and $k_{max}$ = 3 looks. Critical values are determined on the basis of O'Brien-Fleming-like $\alpha$- and $\beta$-spending functions in *R* package *rpact* (Wassmer & Pahlke, 2020).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2022-02-26 protocol_covid_transparency.qmd
First commit

---
title: "Transparency of COVID-19 medical literature"
author:
  - Ahmad Sofi-Mahmudi^[Seqiz Health Network, Kurdistan University of Medical Sciences, Seqiz, Iran.; a.sofimahmudi@gmail.com]
  - Eero Raittio*^[Institute of Dentistry, University of Eastern Finland, Kuopio, Finland.; eero.raittio@uef.fi]
  - Sergio E. Uribe^[Department of Conservative Dentistry and Oral Health & Bioinformatics Research Unit, Riga Stradins University, Riga, Latvia.; sergio.uribe@rsu.lv]
format: html
editor: visual
---

**Corresponding author:** Eero Raittio; Address: ; Telephone: ; Email: eero.raittio\@uef.fi.

**Authors' contributions:** All authors conceived and designed, and wrote and revised the protocol.

# Abstract

**Objective and aim:** We aim to assess the use of transparency indicators (data availability, code availability, statements of protocol registration and conflicts of interest and funding disclosures) from open access COVID-19-related articles published in medical journals available from the Europe PubMed Central (PMC) database. We will map transparency in medical articles across different journals, publication years, publishers, specialties and journal impact factors.

**Methods:** We will search and export all open access articles from PubMed-indexed journals available in the Europe PMC database. We will detect transparency indicators with a validated and automated tool developed to extract the indicators from the downloaded articles. Basic journal- and article-related information will be retrieved from the PMC database. We will use R for the searches (europepmc package), data handling, analysis (rtransparent package) and reporting (ggplot2 package).

**Discussion:** Our results will establish the current state of the rapidly publishing COVID-19 research reporting transparency and may indicate areas for improvement as well as specific steps or requirements to ensure reporting transparency. This would influence both scientists' and the public's perception of the credibility of results published in peer-reviewed journals.

***Keywords:*** COVID-19; Dentistry; Data; FAIR data principles; Open Science.

# Aims

Our aim is to assess the uptake of transparent practices (data sharing, code sharing, conflict of interest disclosures, funding disclosures, statements of protocol registration) from open access full-text COVID-19-related articles published in medical journals available from the Europe PMC. We map transparency across publication years, publishers and journals.

# Methods

## Data sources and study selection

We will search for open access COVID-19-related articles from dental journals available in the Europe PMC database. We will review dental journals based on a list of PubMed-indexed dental journals provided by the National Library of Medicine (NLM) catalog (https://bit.ly/3tz2ScL). We will restrict our search to papers in English. We will consider papers published from 2020-01-01 to 2021-12-31 with variants of COVID-19 keyword in their title. We will download all identified records in XML full-text format (for full-text evaluation) and in CSV format (for descriptive tabulations) from the database. The search query will be as follows:

'(TITLE:COVID OR COVID19 OR "SARS-CoV-2" OR "SARS-CoV2" OR SARSCoV2 OR"SARSCoV-2" OR "SARS coronavirus 2" OR "2019 nCoV" OR "2019nCoV" OR "2019-novel CoV" OR "nCov 2019" OR "nCov 19" OR "severe acute respiratory syndrome coronavirus 2" OR "novel coronavirus disease" OR "novel corona virus disease" OR "corona virus disease 2019" OR "coronavirus disease 2019" OR "novel coronavirus pneumonia" OR "novel corona virus pneumonia" OR "severe acute respiratory syndrome coronavirus 2") AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:\[2020-01-01 TO 2021-12-31\]) AND (OPEN_ACCESS:y)

## Data extraction and synthesis

We aim to describe the use of five transparency indicators:

-   data availability,

-   code availability,

-   conflicts of interest disclosures,

-   funding disclosures,

-   statements of protocol registration.

To do so, we will use a validated and automated tool developed by Serghiou et al.1 to extract the indicators from open access articles downloaded from the Europe PMC database.

Basic journal- and article-related information (publisher, publication year and journal name) will be retrieved from the Europe PMC database. We will download additional information about journal impact factors and journal subject categories from the Journal Citation Reports.

We will use R 2 for searches, data handling, analysis and reporting. The searches and data export from the Europe PMC will be conducted with the europepmc package3. Transparency indicators from the available full-texts will be extracted with the rtransparent package4. Trends over time in transparency indicators by journal- and article-related information will be reported using descriptive tabulations and graphical illustrations, for instance using the ggplot2 package5.

The protocol of this study will be published beforehand on the Open Science Framework website (OSF, osf.io).

# Limitations

The study sample will be restricted to open access articles in the Europe PMC database which may not correspond to all COVID-19-related studies published in medical journals. However, as the majority of COVID-19-related papers are open access, this will not diminish the strength of our interpretations considerably.

### 2022-02-26 covid_transparency_analysis.Rmd
Some changes to the protocol

---
title: "Transparency of COVID-19-related research - Analysis"
author: Ahmad Sofi-Mahmudi, Eero Raittio
date: "2022-02-11"
output: html_document
---

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = 'C:/Users/ahmad/Google Drive/Meta-Research in Dentistry/FAIRness of COVID-19 data/data/')
```

# Loading required packages

```{r}
pacman::p_load(dplyr,
               ggplot2,
               knitr)
```

# Loading the datasets
First, set working directory to the folder that has all the datasets in it using setwd() function and then run the following code:
```{r}
opendata <- read.csv("covid_transparency_opendata.csv")
```

# Results
Number of all papers (open access and non open access), open-access papers and open-access percentage:
```{r}
kable(data.frame(hits_all = 92752, 
           hits_oa = nrow(opendata),
           oa_percentage = round((nrow(opendata)/92752)*100, 2)))
```

First, adding the real publication year and month to the datasets. The real publication year/month is the year/month the paper was first appear online stored in firstPublicationDate column.
```{r}
library(lubridate)

opendata <- opendata %>%
  mutate(year_firstpub = year(
    as.POSIXlt(firstPublicationDate, 
               format = "%Y-%m-%d")),
    month_firstpub = month(
      as.POSIXlt(firstPublicationDate, 
               format = "%Y-%m-%d")
    )
    )
```

Number of open access papers per year:
```{r}
kable(table(opendata$year_firstpub))
```

The number of journals in our dataset:
```{r}
length(table(opendata$journal))
```
Top 10 journals with the highest number of articles in our dataset, from high to low:
```{r}
kable(head(table(opendata$journal) %>% as.data.frame() %>% arrange(desc(Freq)), 10))
```

The mean and the median of the number of citations to these references:
```{r}
kable(data.frame(Mean = round(mean(opendata$citedByCount),1),
           SD = round(sd(opendata$citedByCount), 2),
           Median = median(opendata$citedByCount),
           IQR = IQR(opendata$citedByCount)
           ))

```

Characteristics of the paper with the highest number of citations:
```{r}
kable(opendata[which.max(opendata$citedByCount),] %>% 
  select(pmid, pmcid, doi.x, title, authorString, journalTitle, pubYear))
```

The number and percentage of papers with a conflict of interest (CoI) disclosure:
```{r}
kable(data.frame(
  number = length(opendata$is_coi_pred[opendata$is_coi_pred == TRUE]),
           percentage = round(length(opendata$is_coi_pred[opendata$is_coi_pred == TRUE])/nrow(opendata)*100, 2)
           ))
```

Confidence interval for CoI:
```{r}
library(epiR)

kable(round(epi.prev(pos = length(opendata$is_coi_pred[opendata$is_coi_pred == TRUE]),
         tested = nrow(opendata),
         se = 0.992,
         sp = 0.995)$ap, 
      2))
```

Comparing CoI disclosure rate in two years:
```{r}
kable(round(prop.table(table(opendata$is_coi_pred, opendata$year_firstpub), 2)*100, 1))
```

Chi-square test:
```{r}
chisq.test(opendata$is_coi_pred, opendata$year_firstpub)
```

The number and percentage of papers with a funding statement:
```{r}
kable(data.frame(number = length(opendata$is_fund_pred[opendata$is_fund_pred == TRUE]),
           percentage = round(length(opendata$is_fund_pred[opendata$is_fund_pred == TRUE])/nrow(opendata)*100, 2)
           ))
```

Confidence interval for funding statement:
```{r}

kable(round(epi.prev(pos = length(opendata$is_fund_pred[opendata$is_fund_pred == TRUE]),
         tested = nrow(opendata),
         se = 0.997,
         sp = 0.981)$ap, 
      2))
```

Comparing funding statement rate in two years:
```{r}
kable(round(prop.table(table(opendata$is_fund_pred, opendata$year_firstpub), 2)*100, 1))
```

Chi-square test:
```{r}
chisq.test(opendata$is_fund_pred, opendata$year_firstpub)
```

The number and percentage of papers that were registered beforehand:
```{r}
kable(data.frame(number = length(opendata$is_register_pred[opendata$is_register_pred == TRUE]),
           percentage = round(length(opendata$is_register_pred[opendata$is_register_pred == TRUE])/nrow(opendata)*100, 2)
           ))
```

Confidence interval for registration:
```{r}

kable(round(epi.prev(pos = length(opendata$is_register_pred[opendata$is_register_pred == TRUE]),
         tested = nrow(opendata),
         se = 0.955,
         sp = 0.997)$ap, 
      2))
```

Comparing registration rate in two years:
```{r}
kable(round(prop.table(table(opendata$is_register_pred, opendata$year_firstpub), 2)*100, 1))
```

Chi-square test:
```{r}
chisq.test(opendata$is_register_pred, opendata$year_firstpub)
```

The number and percentage of papers that shared data (preliminary):
```{r}
kable(data.frame(number = length(opendata$is_open_data[opendata$is_open_data == TRUE]),
           percentage = round(length(opendata$is_open_data[opendata$is_open_data == TRUE])/nrow(opendata)*100, 2)
           ))
```

Comparing data sharing rate in two years (preliminary):
```{r}
kable(round(prop.table(table(opendata$is_open_data, opendata$year_firstpub), 2)*100, 1))
```

The number and percentage of papers that shared code (preliminary):
```{r}
kable(data.frame(number = length(opendata$is_open_code[opendata$is_open_code == TRUE]),
           percentage = round(length(opendata$is_open_code[opendata$is_open_code == TRUE])/nrow(opendata)*100, 2)
           ))
```

Comparing code sharing rate in two years (preliminary):
```{r}
kable(round(prop.table(table(opendata$is_open_code, opendata$year_firstpub), 2)*100, 1))
```

## Journal Impact Factor (JIF) analyses

First, importing jif file:
```{r}
jif <- read.csv("jif_2020.csv")
```

Merging it with opendata:
```{r}
db_jif <- merge(opendata, jif, by.x = "journalIssn", by.y = "ISSN")
db_jif$X2020.JIF <- as.numeric(db_jif$X2020.JIF)
```

Total number of papers published in journals with no JIF:
```{r}
nrow(opendata) - nrow(db_jif)
```
JIF based on COI disclosure:

```{r}
kable(db_jif %>% 
  group_by(is_coi_pred) %>% 
  summarise(Mean = round(mean(X2020.JIF, na.rm = T),3),
            SD = round(sd(X2020.JIF, na.rm = T),4),
            Median = median(X2020.JIF, na.rm = T),
            IQR = IQR(X2020.JIF, na.rm = T)))
```

Test for normality:
```{r}
set.seed(10)

with(db_jif, shapiro.test(sample(X2020.JIF[is_coi_pred == TRUE], 5000)))
with(db_jif, shapiro.test(X2020.JIF[is_coi_pred == FALSE]))
```

Data is not normal, hence we will use unpaired two-samples Wilcoxon test.
```{r}
wilcox.test(X2020.JIF~is_coi_pred, data = db_jif, exact = FALSE)
```

```{r}
kable(db_jif %>% 
  group_by(is_fund_pred) %>% 
  summarise(Mean = round(mean(X2020.JIF, na.rm = T), 3),
            SD = round(sd(X2020.JIF, na.rm = T), 4),
            Median = median(X2020.JIF, na.rm = T),
            IQR = IQR(X2020.JIF, na.rm = T)))
```

Test for normality:
```{r}
set.seed(10)

with(db_jif, shapiro.test(sample(X2020.JIF[is_fund_pred == TRUE], 5000)))
with(db_jif, shapiro.test(X2020.JIF[is_fund_pred == FALSE]))
```

Data is not normal, hence we will use unpaired two-samples Wilcoxon test.
```{r}
wilcox.test(X2020.JIF~is_fund_pred, data = db_jif, exact = FALSE)
```

```{r}
kable(db_jif %>% 
  group_by(is_register_pred) %>% 
  summarise(Mean = round(mean(X2020.JIF, na.rm = T), 3),
            SD = round(sd(X2020.JIF, na.rm = T), 4),
            Median = median(X2020.JIF, na.rm = T),
            IQR = IQR(X2020.JIF, na.rm = T)))
```

Test for normality:
```{r}
set.seed(10)

with(db_jif, shapiro.test(X2020.JIF[is_register_pred == TRUE]))
with(db_jif, shapiro.test(sample(X2020.JIF[is_register_pred == FALSE], 5000)))
```

Data is not normal, hence we will use unpaired two-samples Wilcoxon test.
```{r}
wilcox.test(X2020.JIF~is_register_pred, data = db_jif, exact = FALSE)
```

## Monthly trend
Now, drawing monthly trend for each year:
```{r}
summarybyMonth <- opendata %>%
  group_by(month_firstpub, year_firstpub) %>%
  summarise(counts = n(),
            coi_true = sum(is_coi_pred == TRUE),
            fund_true = sum(is_fund_pred == TRUE),
            reg_true = sum(is_register_pred == TRUE))

summarybyMonth %>%
  ggplot(aes(x = month_firstpub, y = counts)) + 
  geom_col(position = position_dodge()) +
  facet_grid(year_firstpub ~ .) +
  geom_bar(stat = "identity") +
  scale_x_discrete(name="Month", limits = 1:12, labels = 1:12)

coi_summarybyMonth <- opendata %>%
  group_by(month_firstpub, year_firstpub, is_coi_pred) %>%
  summarise(counts = n())

coi_plot <- coi_summarybyMonth %>%
  ggplot(aes(x = month_firstpub, y = counts, fill = is_coi_pred)) + 
  geom_col(position = position_dodge()) +
  facet_grid(. ~ year_firstpub) +
  geom_bar(stat = "identity") +
  scale_x_discrete(name="Month", limits = 1:12, labels = 1:12) +
  theme_minimal() +
  theme(legend.title = element_blank()) +
  labs(title = "Conflict of Interest disclosure")

fund_summarybyMonth <- opendata %>%
  group_by(month_firstpub, year_firstpub, is_fund_pred) %>%
  summarise(counts = n())

fund_plot <- fund_summarybyMonth %>%
  ggplot(aes(x = month_firstpub, y = counts, fill = is_fund_pred)) + 
  geom_col(position = position_dodge()) +
  facet_grid(. ~ year_firstpub) +
  geom_bar(stat = "identity") +
  scale_x_discrete(name="Month", limits = 1:12, labels = 1:12) +
  theme_minimal() +
  theme(legend.title = element_blank()) +
  labs(title = "Funding statement")

reg_summarybyMonth <- opendata %>%
  group_by(month_firstpub, year_firstpub, is_register_pred) %>%
  summarise(counts = n())

reg_plot <- reg_summarybyMonth %>%
  ggplot(aes(x = month_firstpub, y = counts, fill = is_register_pred)) + 
  geom_col(position = position_dodge()) +
  facet_grid(. ~ year_firstpub) +
  geom_bar(stat = "identity") +
  scale_x_discrete(name="Month", limits = 1:12, labels = 1:12) +
  theme_minimal() +
  theme(legend.title = element_blank()) +
  labs(title = "Registration statement")

library(ggpubr)
figure <- ggarrange(coi_plot + rremove("xlab") + theme(axis.text.x=element_blank()), 
                    fund_plot + rremove("xlab") + theme(axis.text.x=element_blank(), strip.text.x = element_blank()),
                    reg_plot + theme(strip.text.x = element_blank()),
                    ncol = 1, nrow = 3, vjust = 1, 
                    align = "hv", common.legend = T, legend = "right")

# tiff("Figure.tiff", width = 6, height = 7, units = "in", res = 300)
figure
#dev.off()
```

### 2022-02-26 covid_transparency_dataset_creation.Rmd
Some changes to the protocol

---
title: "Transparency indicators across the dentistry and oral health literature"
author: Eero Raittio, Ahmad Sofi-Mahmudi, Erfan Shamsoddin, Sergio E Uribe, Julie Farmer
date: "1/8/2022"
output: html_document
---

# Loading required packages

```{r}
pacman::p_load(tidyverse,
               rtransparent, 
               metareadr, 
               europepmc)
```

# Load the datasets

## Load all open access dentistry and oral health papers indexed in PubMed to the end of 2021:

```{r}
COVID_19_keywords <- read.csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vS6ibaTIJ0-4MiOZrLHKwo-Ya3MjlGX4f-dPQwdmWwQZhRGi6C83K1I_WceUJPH9zGyBc4L-lSalET_/pub?output=csv")
```

```{r}
COVID_19_keywords <- COVID_19_keywords %>% mutate(
  search.term.title = paste0("TITLE:", '"', Keyword, '"'),
  search.term.keyword = paste0("KW:", '"', Keyword, '"'),
  search.term.results = paste0("RESULTS:", '"', Keyword, '"'))

TitleQuery <- paste(COVID_19_keywords$search.term.title,
                     collapse = " OR ")

KeywordQuery <- paste(COVID_19_keywords$search.term.keyword,
                     collapse = " OR ")

ResultQuery <- paste(COVID_19_keywords$search.term.results,
                     collapse = " OR ")
```

## Load all open access dentistry and oral health papers indexed in PubMed to the end of 2021:

```{r}

db_2020_1 <- epmc_search(query = paste0("'(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2020-01-01 TO 2020-06-30]) AND ((IN_EPMC:y) OR (OPEN_ACCESS:y)) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")', "'"), limit = 100000, output = "parsed")

db_2020_1 <- epmc_search(query = paste0("'(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2020-01-01 TO 2020-06-30]) AND (OPEN_ACCESS:y) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")', "'"), limit = 100000, output = "parsed")

db_2020_2 <- epmc_search(query = paste0("'(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2020-07-01 TO 2020-12-31]) AND ((IN_EPMC:y) OR (OPEN_ACCESS:y)) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")', "'"), limit = 100000, output = "parsed")

db_2021_1 <- epmc_search(query = paste0("'(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2021-01-01 TO 2021-06-30]) AND ((IN_EPMC:y) OR (OPEN_ACCESS:y)) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")', "'"), limit = 100000, output = "parsed")

db_2020_2 <- epmc_search(query = paste0("'(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2021-07-01 TO 2021-12-31]) AND ((IN_EPMC:y) OR (OPEN_ACCESS:y)) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")', "'"), limit = 100000, output = "parsed")

db <- epmc_search(query = paste0("'(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2020-01-01 TO 2021-12-31]) AND ((IN_EPMC:y) OR (OPEN_ACCESS:y)) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")', "'"), limit = 1000000, output = "parsed")

db <- epmc_search(query = paste0("'(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2020-01-01 TO 2021-12-31]) AND (OPEN_ACCESS:y) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")', "'"), limit = 1000000, output = "parsed")

db <- epmc_search(
  query = paste0(
    "'(",
    TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery,
    ") ",
    'AND (SRC:"MED") 
    AND (LANG:"eng" OR LANG:"en" OR LANG:"us") 
    AND (FIRST_PDATE:[2020-01-01 TO 2021-12-31])
    AND ((IN_EPMC:y) OR (OPEN_ACCESS:y))
    AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")',
    "'"
  ),
  limit = 100000, output = "parsed"
)
```

Let's see how many duplicates are there:
```{r}
table(duplicated(db$pmid))
```

Removing the duplicates:
```{r}
db <- db %>% distinct(pmid, .keep_all = TRUE)
```

Let's see the first six rows of the database:
```{r}
head(db)
```

Removing "PMC" from the cells
```{r}
db_2020_1$pmcid_ <- gsub("PMC", "", as.character(db_2020_1$pmcid))

db_2020_2$pmcid_ <- gsub("PMC", "", as.character(db_2020_2$pmcid))

db_2021_1$pmcid_ <- gsub("PMC", "", as.character(db_2021_1$pmcid))

db_2021_2$pmcid_ <- gsub("PMC", "", as.character(db_2021_2$pmcid))

db$pmcid_ <- gsub("PMC", "", as.character(db$pmcid))
```

Now, we make a folder for xml format articles and switch to that folder:
```{r}
dir.create("pmc")
setwd("pmc")
```

Next, we download xmls in format accessible with rtransparent:
```{r}
db_2020_1 <- subset(db_2020_1, pmcid_!=c(7264604, 7194244, 7323088, 7300787, 7276714, 7262104, 7228371))
db_2020_2 <- subset(db_2020_2, pmcid_!=c(7677857, 7164703, 7678408, 7486161, 7645956, 7486166, 7675701, 7405208, 7753328, 7575215, 7759253, 7568026, 7598225, 7416582, 7649313, 7492687, 7571788, 7675665, 7572235, 7457217))
db_2021_1 <- subset(db_2021_1, pmcid_!=c(8013335, 8013644, 8250125, 8014725, 8251275, 7486807, 8420411, 8251218, 8251305, 7894077))
db_2021_2 <- subset(db_2021_2, pmcid_!=8653067)

db <- subset(db_2021_2, pmcid_!=c(7264604, 8013335))

sapply(db_2020_1$pmcid_, mt_read_pmcoa)

sapply(db_2020_2$pmcid_, mt_read_pmcoa)

sapply(db_2021_1$pmcid_, mt_read_pmcoa)

sapply(db_2021_2$pmcid_, mt_read_pmcoa)

sapply(db$pmcid_, mt_read_pmcoa)

db <- subset(db, !(pmcid_ %in% c(8251275, 8251218, 7486161, 7228371, 8485571)))

```
Error in 8013644 - The metadata format 'pmc' is not supported by the item or by the repository.

Error in 7572235 - The value of the identifier argument is unknown or illegal in this repository.

Now we run rtransparent:
```{r}
filepath = dir(pattern=glob2rx("PMC*.xml"))

results_table_all <- sapply(filepath, rt_all_pmc)

results_table_data <- rt_data_code_pmc_list(
  filepath,
  remove_ns=F,
  specificity = "low")

write.csv(rt_data_code_pmc_list(filepath, remove_ns=F, specificity = "low"), "rt_data.csv")

f1_data <- rt_data_code_pmc_list(f1, remove_ns=F, specificity = "low")

```

A list is created now. We should convert this list to a dataframe:
```{r}
df <- data.table::rbindlist(results_table_all, fill = TRUE)
```

Merge data sharing results to database file:

```{r}
setwd('..')
db <- db[!duplicated(db[, 4]),]
opendata <- merge(db, results_table_data, by = "pmid") %>% merge(df)
View(opendata)
```

### 2022-02-26 protocol_covid_transparency.qmd
Some changes to the protocol

@@
 editor: visual
+bibliography: references.bib
+csl: vancouver.csl
 ---
@@
 
-We will search for open access COVID-19-related articles from dental journals available in the Europe PMC database.
-We will review dental journals based on a list of PubMed-indexed dental journals provided by the National Library of Medicine (NLM) catalog (https://bit.ly/3tz2ScL).
+We will search for open access COVID-19-related articles from medical journals available in the Europe PMC database.
 We will restrict our search to papers in English.
-We will consider papers published from 2020-01-01 to 2021-12-31 with variants of COVID-19 keyword in their title.
+We will consider papers published from 2020-01-01 to 2021-12-31 with variants of COVID-19 keyword in their title, keywords, or results.
 We will download all identified records in XML full-text format (for full-text evaluation) and in CSV format (for descriptive tabulations) from the database.
@@
 
-'(TITLE:COVID OR COVID19 OR "SARS-CoV-2" OR "SARS-CoV2" OR SARSCoV2 OR"SARSCoV-2" OR "SARS coronavirus 2" OR "2019 nCoV" OR "2019nCoV" OR "2019-novel CoV" OR "nCov 2019" OR "nCov 19" OR "severe acute respiratory syndrome coronavirus 2" OR "novel coronavirus disease" OR "novel corona virus disease" OR "corona virus disease 2019" OR "coronavirus disease 2019" OR "novel coronavirus pneumonia" OR "novel corona virus pneumonia" OR "severe acute respiratory syndrome coronavirus 2") AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:\[2020-01-01 TO 2021-12-31\]) AND (OPEN_ACCESS:y)
+(COVID-19 query for title, keywords, and results) AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:\[2020-01-01 TO 2021-12-31\]) AND ((IN_EPMC:y) OR (OPEN_ACCESS:y)) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")'
+
+The COVID-19 query is available in the *covid_keywords.csv* file.
 
@@
 
-To do so, we will use a validated and automated tool developed by Serghiou et al.1 to extract the indicators from open access articles downloaded from the Europe PMC database.
+To do so, we will use a validated and automated tool developed by Serghiou et al. [@serghiou2021] to extract the indicators from open access articles downloaded from the Europe PMC database.
 
 Basic journal- and article-related information (publisher, publication year and journal name) will be retrieved from the Europe PMC database.
-We will download additional information about journal impact factors and journal subject categories from the Journal Citation Reports.
+We will download additional information about journal impact factors from the Journal Citation Reports.
 
-We will use R 2 for searches, data handling, analysis and reporting.
-The searches and data export from the Europe PMC will be conducted with the europepmc package3.
-Transparency indicators from the available full-texts will be extracted with the rtransparent package4.
-Trends over time in transparency indicators by journal- and article-related information will be reported using descriptive tabulations and graphical illustrations, for instance using the ggplot2 package5.
+We will use R [@R-base] for searches, data handling, analysis and reporting.
+The searches and data export from the Europe PMC will be conducted with the europepmc package [@europepmc].
+Transparency indicators from the available full-texts will be extracted with the rtransparent package [@rtransparent].
+Trends over time in transparency indicators by journal- and article-related information will be reported using descriptive tabulations and graphical illustrations, for instance using the ggplot2 package [@ggplot2].
 
@@
 However, as the majority of COVID-19-related papers are open access, this will not diminish the strength of our interpretations considerably.
+
+# References

### 2022-03-06 protocol_covid_transparency.qmd
Update protocol_covid_transparency.qmd
protocol addition uribe plus one reference

@@
 title: "Transparency of COVID-19 medical literature"
+
 author:
@@
   - Eero Raittio*^[Institute of Dentistry, University of Eastern Finland, Kuopio, Finland.; eero.raittio@uef.fi]
-  - Sergio E. Uribe^[Department of Conservative Dentistry and Oral Health & Bioinformatics Research Unit, Riga Stradins University, Riga, Latvia.; sergio.uribe@rsu.lv]
+  - Sergio E.
+Uribe^[Department of Conservative Dentistry and Oral Health & Bioinformatics Research Unit, Riga Stradins University, Riga, Latvia.; sergio.uribe@rsu.lv] ^[School of Dentistry, Universidad Austral de Chile, Valdivia, Chile]
+
 format: html
@@
 
-**Corresponding author:** Eero Raittio; Address: ; Telephone: ; Email: eero.raittio\@uef.fi.
+**Corresponding author:** Eero Raittio; **Address**: ; Telephone: ; Email: eero.raittio\@uef.fi.
 
@@
 
-**Objective and aim:** We aim to assess the use of transparency indicators (data availability, code availability, statements of protocol registration and conflicts of interest and funding disclosures) from open access COVID-19-related articles published in medical journals available from the Europe PubMed Central (PMC) database.
+**Objective and aim:** The recent pandemic has shown how crucial it is to have access to quality medical information and data that allows people to have confidence in healthcare decisions.
+One of these indicators is transparency, i.e., the availability of original data supporting research results.
+We aim to assess the use of transparency indicators (data availability, code availability, statements of protocol registration and conflicts of interest and funding disclosures) from open access COVID-19-related articles published in medical journals available from the Europe PubMed Central (PMC) database.
 We will map transparency in medical articles across different journals, publication years, publishers, specialties and journal impact factors.
 
-**Methods:** We will search and export all open access articles from PubMed-indexed journals available in the Europe PMC database.
+**Methods:** We will search and export all open access articles from PubMed-indexed journals available in the Europe PMC database published between January 2020 and December 2021.
 We will detect transparency indicators with a validated and automated tool developed to extract the indicators from the downloaded articles.
@@
 
-Our aim is to assess the uptake of transparent practices (data sharing, code sharing, conflict of interest disclosures, funding disclosures, statements of protocol registration) from open access full-text COVID-19-related articles published in medical journals available from the Europe PMC.
+Our objective is to evaluate the transparency indicators (data sharing, code, declaration of interest, funding, availability of research protocol) from open access full-text COVID-19-related articles published in medical journals available from the Europe PMC.
 We map transparency across publication years, publishers and journals.
@@
 
-The COVID-19 query is available in the *covid_keywords.csv* file.
+The full COVID-19 query is available in the *covid_keywords.csv* file.
 
@@
 
-We aim to describe the use of five transparency indicators:
+We aim to describe adherence to five transparency practices in the articles:
 
@@
 
+We will also calculate the proportion of articles available as open access via the Europe PMC from the total number of COVID-19-related detected in the database.
+
 We will use R [@R-base] for searches, data handling, analysis and reporting.
@@
 The study sample will be restricted to open access articles in the Europe PMC database which may not correspond to all COVID-19-related studies published in medical journals.
-However, as the majority of COVID-19-related papers are open access, this will not diminish the strength of our interpretations considerably.
+However, as the majority of COVID-19-related papers are open access[@Kiley2020-ek], this will not diminish the strength of our interpretations considerably.
 

### 2022-03-06 protocol_covid_transparency.qmd
Some minor changes
I added a second affiliation to myself and added url to JCR. Corresponding author field was also filled.

@@
 author:
-  - Ahmad Sofi-Mahmudi^[Seqiz Health Network, Kurdistan University of Medical Sciences, Seqiz, Iran.; a.sofimahmudi@gmail.com]
+  - Ahmad Sofi-Mahmudi^[Seqiz Health Network, Kurdistan University of Medical Sciences, Seqiz, Iran.; a.sofimahmudi@gmail.com] ^[Cochrane Iran Associate Centre, National Institute for Medical Research Development (NIMAD), Tehran, Iran.]
   - Eero Raittio*^[Institute of Dentistry, University of Eastern Finland, Kuopio, Finland.; eero.raittio@uef.fi]
@@
 
-format: html
+# format: html
+format:
+  pdf:
+    colorlinks: true
 editor: visual
@@
 
-**Corresponding author:** Eero Raittio; **Address**: ; Telephone: ; Email: eero.raittio\@uef.fi.
+**Corresponding author:** Eero Raittio; **Address**: University of Eastern Finland, Institute of Dentistry, P.O.
+Box 1627, FI-70211 Kuopio, Finland; **Telephone**: +358 294 45 1111; **Email**: [eero.raittio\@uef.fi](mailto:eero.raittio@uef.fi).
 
@@
 Basic journal- and article-related information (publisher, publication year and journal name) will be retrieved from the Europe PMC database.
-We will download additional information about journal impact factors from the Journal Citation Reports.
+We will download additional information about journal impact factors from the Journal Citation Reports (ٰ[jcr.clarivate.com](https://jcr.clarivate.com/)).
 

### 2022-03-07 appendix_covid_transparency.qmd
Manuscript and appendix files

--- 2022-03-06 protocol_covid_transparency.qmd
@@
 ---
-title: "Transparency of COVID-19 medical literature"
+title: "Transparency of COVID-19-related research in medical journals - Appendix"
 
@@
 
-# format: html
-format:
-  pdf:
-    colorlinks: true
+format: html
+#format:
+#  pdf:
+#    colorlinks: true
 editor: visual
-bibliography: references.bib
-csl: vancouver.csl
 ---
@@
 
-**Authors' contributions:** All authors conceived and designed, and wrote and revised the protocol.
+# Appendix 2
 
-# Abstract
-
-**Objective and aim:** The recent pandemic has shown how crucial it is to have access to quality medical information and data that allows people to have confidence in healthcare decisions.
-One of these indicators is transparency, i.e., the availability of original data supporting research results.
-We aim to assess the use of transparency indicators (data availability, code availability, statements of protocol registration and conflicts of interest and funding disclosures) from open access COVID-19-related articles published in medical journals available from the Europe PubMed Central (PMC) database.
-We will map transparency in medical articles across different journals, publication years, publishers, specialties and journal impact factors.
-
-**Methods:** We will search and export all open access articles from PubMed-indexed journals available in the Europe PMC database published between January 2020 and December 2021.
-We will detect transparency indicators with a validated and automated tool developed to extract the indicators from the downloaded articles.
-Basic journal- and article-related information will be retrieved from the PMC database.
-We will use R for the searches (europepmc package), data handling, analysis (rtransparent package) and reporting (ggplot2 package).
-
-**Discussion:** Our results will establish the current state of the rapidly publishing COVID-19 research reporting transparency and may indicate areas for improvement as well as specific steps or requirements to ensure reporting transparency.
-This would influence both scientists' and the public's perception of the credibility of results published in peer-reviewed journals.
-
-***Keywords:*** COVID-19; Dentistry; Data; FAIR data principles; Open Science.
-
-# Aims
-
-Our objective is to evaluate the transparency indicators (data sharing, code, declaration of interest, funding, availability of research protocol) from open access full-text COVID-19-related articles published in medical journals available from the Europe PMC.
-We map transparency across publication years, publishers and journals.
-
-# Methods
-
-## Data sources and study selection
-
-We will search for open access COVID-19-related articles from medical journals available in the Europe PMC database.
-We will restrict our search to papers in English.
-We will consider papers published from 2020-01-01 to 2021-12-31 with variants of COVID-19 keyword in their title, keywords, or results.
-We will download all identified records in XML full-text format (for full-text evaluation) and in CSV format (for descriptive tabulations) from the database.
-The search query will be as follows:
-
-(COVID-19 query for title, keywords, and results) AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:\[2020-01-01 TO 2021-12-31\]) AND ((IN_EPMC:y) OR (OPEN_ACCESS:y)) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")'
-
-The full COVID-19 query is available in the *covid_keywords.csv* file.
-
-## Data extraction and synthesis
-
-We aim to describe adherence to five transparency practices in the articles:
-
--   data availability,
-
--   code availability,
-
--   conflicts of interest disclosures,
-
--   funding disclosures,
-
--   statements of protocol registration.
-
-To do so, we will use a validated and automated tool developed by Serghiou et al. [@serghiou2021] to extract the indicators from open access articles downloaded from the Europe PMC database.
-
-Basic journal- and article-related information (publisher, publication year and journal name) will be retrieved from the Europe PMC database.
-We will download additional information about journal impact factors from the Journal Citation Reports (ٰ[jcr.clarivate.com](https://jcr.clarivate.com/)).
-
-We will also calculate the proportion of articles available as open access via the Europe PMC from the total number of COVID-19-related detected in the database.
-
-We will use R [@R-base] for searches, data handling, analysis and reporting.
-The searches and data export from the Europe PMC will be conducted with the europepmc package [@europepmc].
-Transparency indicators from the available full-texts will be extracted with the rtransparent package [@rtransparent].
-Trends over time in transparency indicators by journal- and article-related information will be reported using descriptive tabulations and graphical illustrations, for instance using the ggplot2 package [@ggplot2].
-
-The protocol of this study will be published beforehand on the Open Science Framework website (OSF, osf.io).
-
-# Limitations
-
-The study sample will be restricted to open access articles in the Europe PMC database which may not correspond to all COVID-19-related studies published in medical journals.
-However, as the majority of COVID-19-related papers are open access[@Kiley2020-ek], this will not diminish the strength of our interpretations considerably.
-
-# References
+```{r}
+sessionInfo()
+```

### 2022-03-07 covid_transparency_analysis.Rmd
Manuscript and appendix files

@@
 title: "Transparency of COVID-19-related research - Analysis"
-author: Ahmad Sofi-Mahmudi, Eero Raittio
+author: Ahmad Sofi-Mahmudi, Eero Raittio, Sergio E Uribe
 date: "2022-02-11"
@@
 ```{r, setup, include=FALSE}
-knitr::opts_knit$set(root.dir = 'C:/Users/ahmad/Google Drive/Meta-Research in Dentistry/FAIRness of COVID-19 data/data/')
+knitr::opts_knit$set(root.dir = '~/Documents/GitHub/covid-transparency/data')
 ```

### 2022-03-07 manuscript_covid_transparency.qmd
Manuscript and appendix files

---
title: "Transparency of COVID-19-related research in medical journals"

author:
  - Ahmad Sofi-Mahmudi^[Seqiz Health Network, Kurdistan University of Medical Sciences, Seqiz, Iran.; a.sofimahmudi@gmail.com] ^[Cochrane Iran Associate Centre, National Institute for Medical Research Development (NIMAD), Tehran, Iran.]
  - Eero Raittio*^[Institute of Dentistry, University of Eastern Finland, Kuopio, Finland.; eero.raittio@uef.fi]
  - Sergio E. Uribe^[Department of Conservative Dentistry and Oral Health & Bioinformatics Research Unit, Riga Stradins University, Riga, Latvia.; sergio.uribe@rsu.lv] ^[School of Dentistry, Universidad Austral de Chile, Valdivia, Chile]

format: html
#format:
#  pdf:
#    colorlinks: true
editor: visual
bibliography: references.bib
csl: vancouver.csl
---

**Corresponding author:** Eero Raittio; **Address**: University of Eastern Finland, Institute of Dentistry, P.O. Box 1627, FI-70211 Kuopio, Finland; **Telephone**: +358 294 45 1111; **Email**: [eero.raittio\@uef.fi](mailto:eero.raittio@uef.fi).

**Authors' contributions:** All authors conceived and designed, and wrote and revised the protocol.

# Abstract

**Objective:** The recent pandemic has shown how crucial it is to have access to quality medical information and data that allows people to have confidence in healthcare decisions. We aimed to assess the use of transparency indicators (data availability, code availability, statements of protocol registration and conflicts of interest and funding disclosures) from open access COVID-19-related articles.

**Methods:** We searched and exported all open-access COVID-19-related articles from PubMed-indexed journals available in the Europe PMC database published between January 2020 and December 2021. We detected transparency indicators with a validated and automated tool developed to extract the indicators from the downloaded articles. Basic journal- and article-related information were retrieved from the PMC database. We used R for the searches (europepmc package), data handling, analysis (rtransparent package) and reporting (ggplot2 package).

**Results:** Of 72,999 available articles published in XX medical journals, XX% provided conflicts of interest disclosure and XX% funding disclosure and X% were preregistered. XX studies shared raw data (XX%) and XX studies shared code. Transparent practices were more common in articles published in journals with XX impact factors, and in 2020 than in 2021.

**Conclusion:** While the majority of the papers had a COI disclosure, the prevalence of the other transparency practices was far from the acceptable level. A much stronger commitment to open science practices, particularly to preregistration, data and code sharing, is needed from all stakeholders.

***Keywords:*** COVID-19; Data; FAIR data principles; Open Science.

# Introduction

The foundation of science is to generate knowledge from reproducible findings (1). Because of the failure to reproduce previous research, the open science movement emerged in recent years (2). The movement's principal focus is on making science more accessible and trustworthy. Data and code sharing, protocol registration and funding and conflicts of interest (COI) disclosures are considered important features of open science and crucial when assessing the credibility of certain scientific findings (2,3). Science without credibility and quality would do a major human, societal and economic disservice (4).

COVID-19 pandemic launched an incredibly fast and voluminous avalanche of scientific publications across the different fields of science. Related to the COVID-19 publications there has been a considerable amount of discussion and examples related to poor or suboptimal data, methods and other scientific practices (5--8). Unsurprisingly, there have been signs of fabricated data, poor data and methods and conflicts of interests in the scientific literature related to the COVID-19 pandemic, which has resulted in a great number of retracted papers (208 as of the end of January 2022)(9). Probably, if transparent scientific practices, like data and code sharing, would be more like the rule than the exception in biomedical research, pre- and post-publication evaluation of the credibility of studies could be much easier (10,11).

The COVID-19 pandemic and its implications for dental services, oral health, oral health-related behaviors and dental education have received great attention from the dental research community. Nevertheless, to date, only one paper related to dentistry or published in dental has been retracted from the Journal of Oral and Maxillofacial Pathology (9). However, meta-research about reporting, methods or reproducibility in COVID-19-related dental research has been limited. Particularly sharing data with adequate quality would allow reanalyses of findings (10).

Our aim was to assess the uptake of transparent scientific practices (data sharing, code sharing, COI disclosures, funding disclosures, and protocol registration) from open access full-text COVID-19-related articles published in dental journals available from the Europe PubMed Central (EPMC) database. We also assessed the FAIRness (Findable, Accessible, Interoperable, and Reusable) of data from those articles that have shared their data. We mapped transparency practices across publication years, publishers and journals.

# Methods

The protocol of this study was published beforehand on the Open Science Framework (OSF) website ([osf.io/5kx2n](https://osf.io/5kx2n)). All code and data related to the study were shared via its OSF repository ([osf.io/x3kb6](https://osf.io/x3kb6/)) at the time of submission of the manuscript.

## Data sources and study selection

We searched for open access COVID-19-related articles from medical journals available in the EPMC database. We restricted our search to papers in English and considered papers published from 2020-01-01 to 2021-12-31 with variants of COVID-19 keyword in their title, keywords, or results. This enabled us to bypass screening for COVID-19-related papers and decrease the human labor for our research. We confirmed our approach by choosing a random sample of 100 papers and of those, all were relevant to COVID-19.

The search query was as follows:

(COVID-19 query for title, keywords, and results) AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:\[2020-01-01 TO 2021-12-31\]) AND ((IN_EPMC:y) OR (OPEN_ACCESS:y)) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")'

The full COVID-19 query is available in Appendix 1.

We downloaded all identified available records in XML full-text format (for full-text evaluation) using the metareadr package [@metareadr] from the database.

## Data extraction and synthesis

We assessed adherence to five transparency practices in the articles:

-   data availability,

-   code availability,

-   conflicts of interest disclosures,

-   funding disclosures,

-   statements of protocol registration.

To do so, we used a validated and automated tool developed by Serghiou et al. [@serghiou2021] suitable to identify these five transparent practices from articles in XML format from the EPMC database.

Basic journal- and article-related information (publisher, publication year and journal name) were retrieved from the EPMC database. We downloaded additional information about journal impact factors from the Journal Citation Reports (ٰ[jcr.clarivate.com](https://jcr.clarivate.com/)).

We also calculated the proportion of articles available as open-access via the Europe PMC from the total number of COVID-19-related detected in the database.

## Data analysis

We used R v4.1.2 [@R-base] for searches, data handling, analysis and reporting. The searches and data export from the Europe PMC were conducted with the europepmc package [@europepmc]. Transparency indicators from the available full-texts were extracted with the rtransparent package [@rtransparent]. Trends over time in transparency indicators by journal- and article-related information were reported using descriptive tabulations and graphical illustrations, for instance using the ggplot2 package [@ggplot2]. We used the numbers for sensitivity and specificity of the rtransparent package (3) to generate 95% confidence intervals (CIs) for our prevalence estimates of the transparency indicators with the epiR package [@epiR]. We also provided monthly trends for transparency indicators, to check whether there were any trends in transparency practices over time. We used Chi-square for comparing indicators between years and the Wilcoxon rank sum test to test whether there was a relationship between transparency indicators and JIFs. The thorough system information for the analysis is available in Appendix 2.

# Results

The total number of COVID-19-related papers (open access and non-open access) was 1038, of which full texts of 657 papers (63.3%) were accessible via the EPMC. Of those, 337 were published in 2020 and 320 in 2021. However, as we could not retrieve seven full texts from the EPMC database due to technical issues, our final sample included 650 full-text articles (Supplementary figure 1).

These articles were published in 59 journals of which the top five were British Dental Journal (n=69), Journal of Dental Education (n=59), British Journal of Oral and Maxillofacial Surgery (n=51), Oral Diseases (n=47), and BMC Oral Health (n=40). The mean and the median of the number of citations to these articles were 7.7 (standard deviation (SD) = 43.15) and 2 (interquartile range (IQR) = 5). The highest cited article with 854 citations was an article published in the International Journal of Oral Sciences in 2020 (20).

About three-fourths of the articles (n=478) had a statement to disclose COI (73.5%, 95% CI: 70.0%--76.8%). Articles published in 2021 had a higher rate of COI disclosure (79.6% vs 67.8%, P \< 0.001). About four-tenths of the references had a funding statement (n=261) (40.2%, 95% CI: 36.5%--43.0%). Articles in 2021 had a funding statement more frequently than 2020 articles (49.1% vs 31.6%, P \< 0.001). Less than 1 in 20 of the articles (n=27) were registered beforehand (4.2%, 95% CI: 2.9%--6.0%) and the rate was six times higher in 2021 compared with 2020 (7.2% vs 1.1%, P \< 0.001). Regarding data and code availability, as we chose the software to have low specificity to be as sensitive as possible, we got some false positives. At first, we got four papers with shared data and none with shared code. After checking manually, we omitted four false positives from the shared data papers. One of them used the journal's template: "The data that support the findings of this study are openly available in \[repository name\] at \[DOI\]" (21). Hence finally, one paper had shared data (0.2%). This article was published in BMC Oral Health in 2021 (22).

A total of 117 articles were published in journals with no impact factor. Papers that had a COI disclosure, funding statement and were registered tend to be published in journals with higher JIF (P = 0.003, \< 0.001, and = 0.007, respectively). Detailed information is illustrated in Table 1.

Figure 1 shows the monthly trend for three transparency indicators. There was an inconsistent trend for the indicators. However, the COI disclosure percentage seemed to be more consistent and showed an increasing trend over time.

# Discussion

## Limitations

The study sample will be restricted to open access articles in the Europe PMC database which may not correspond to all COVID-19-related studies published in medical journals. However, as the majority of COVID-19-related papers are open access[@Kiley2020-ek], this will not diminish the strength of our interpretations considerably.

# Conclusion

# References

### 2022-03-07 protocol_covid_transparency.qmd
Manuscript and appendix files

@@
 ---
-title: "Transparency of COVID-19 medical literature"
+title: "Transparency of COVID-19-related research in medical journals"
 
@@
 **Objective and aim:** The recent pandemic has shown how crucial it is to have access to quality medical information and data that allows people to have confidence in healthcare decisions.
-One of these indicators is transparency, i.e., the availability of original data supporting research results.
-We aim to assess the use of transparency indicators (data availability, code availability, statements of protocol registration and conflicts of interest and funding disclosures) from open access COVID-19-related articles published in medical journals available from the Europe PubMed Central (PMC) database.
-We will map transparency in medical articles across different journals, publication years, publishers, specialties and journal impact factors.
+We aim to assess the use of transparency indicators (data availability, code availability, statements of protocol registration and conflicts of interest and funding disclosures) from open access COVID-19-related articles.
 

### 2022-03-08 manuscript_covid_transparency.qmd
The first part of the results

@@
 
-The total number of COVID-19-related papers (open access and non-open access) was 1038, of which full texts of 657 papers (63.3%) were accessible via the EPMC.
-Of those, 337 were published in 2020 and 320 in 2021.
-However, as we could not retrieve seven full texts from the EPMC database due to technical issues, our final sample included 650 full-text articles (Supplementary figure 1).
+The total number of COVID-19-related papers (open access and non-open access) was 92752, of which full texts of 73042 papers (78.7%) were accessible via the EPMC.
+Of those, 33137 were published in 2020 and 39905 in 2021.
+However, as we could not retrieve 43 full texts from the EPMC database due to technical issues, our final sample included 72999 full-text articles (Appendix 3).
 
-These articles were published in 59 journals of which the top five were British Dental Journal (n=69), Journal of Dental Education (n=59), British Journal of Oral and Maxillofacial Surgery (n=51), Oral Diseases (n=47), and BMC Oral Health (n=40).
-The mean and the median of the number of citations to these articles were 7.7 (standard deviation (SD) = 43.15) and 2 (interquartile range (IQR) = 5).
-The highest cited article with 854 citations was an article published in the International Journal of Oral Sciences in 2020 (20).
+These articles were published in 4892 journals of which the top five were PLoS ONE (n=6825), International Journal of Environmental Research and Public Health (n=2686), Medicine (n=2318), BMJ Open (n=1926), and eLife (n=942).
+The mean and the median of the number of citations to these articles were 3.1 (standard deviation (SD) = 21.76) and 0 (interquartile range (IQR) = 2).
+The highest cited article with 2915 citations was an article published in the Cell in 2020 [@Walls2020].
 
-About three-fourths of the articles (n=478) had a statement to disclose COI (73.5%, 95% CI: 70.0%--76.8%).
-Articles published in 2021 had a higher rate of COI disclosure (79.6% vs 67.8%, P \< 0.001).
-About four-tenths of the references had a funding statement (n=261) (40.2%, 95% CI: 36.5%--43.0%).
-Articles in 2021 had a funding statement more frequently than 2020 articles (49.1% vs 31.6%, P \< 0.001).
-Less than 1 in 20 of the articles (n=27) were registered beforehand (4.2%, 95% CI: 2.9%--6.0%) and the rate was six times higher in 2021 compared with 2020 (7.2% vs 1.1%, P \< 0.001).
+More than nine-tenth of the articles (n=67009) had a statement to disclose COI (91.79%, 95% CI: 91.59%--91.99%).
+Articles published in 2021 had a higher rate of COI disclosure (93.2% vs 90.1	%, P \< 0.001).
+More than eight-tenths of the references had a funding statement (n=60530) (82.92%, 95% CI: 82.64%--83.19%).
+Articles in 2021 had a funding statement more frequently than 2020 articles (83.9% vs 81.7%, P \< 0.001).
+Less than 1 in 6 of the articles (n=11457) were registered beforehand (15.69%, 95% CI: 15.43%--15.96%) and the rate was higher in 2021 compared with 2020 (16.2% vs 15.0%, P \< 0.001).
+
 Regarding data and code availability, as we chose the software to have low specificity to be as sensitive as possible, we got some false positives.
-At first, we got four papers with shared data and none with shared code.
-After checking manually, we omitted four false positives from the shared data papers.
-One of them used the journal's template: "The data that support the findings of this study are openly available in \[repository name\] at \[DOI\]" (21).
-Hence finally, one paper had shared data (0.2%).
-This article was published in BMC Oral Health in 2021 (22).
+Less than 1 in 7 articles had shared their data (n=9493) (13.00%, 95% CI: %--%).
+Articles in 2021 had shared data more frequently than 2020 articles (13.2% vs 12.8%, P \< 0.001).
+Less than 1 in 30 of the articles (n=2264) had shared their code (3.10%, 95% CI: %--%) and the rate was higher in 2021 compared with 2020 (3.3% vs 2.8%, P \< 0.001).
 
-A total of 117 articles were published in journals with no impact factor.
+A total of 51601 articles were published in journals with no impact factor.
 Papers that had a COI disclosure, funding statement and were registered tend to be published in journals with higher JIF (P = 0.003, \< 0.001, and = 0.007, respectively).

### 2022-03-09 manuscript_covid_transparency.qmd
Results are complete

@@
 
-Regarding data and code availability, as we chose the software to have low specificity to be as sensitive as possible, we got some false positives.
-Less than 1 in 7 articles had shared their data (n=9493) (13.00%, 95% CI: %--%).
-Articles in 2021 had shared data more frequently than 2020 articles (13.2% vs 12.8%, P \< 0.001).
-Less than 1 in 30 of the articles (n=2264) had shared their code (3.10%, 95% CI: %--%) and the rate was higher in 2021 compared with 2020 (3.3% vs 2.8%, P \< 0.001).
+Regarding data and code availability, less than 1 in 7 articles had shared their data (n=9493) (13.00%, 95% CI: 12.76%--13.25%).
+Articles in 2021 had shared data more frequently than 2020 articles (13.2% vs 12.8%, P=0.1287).
+Less than 1 in 30 of the articles (n=2264) had shared their code (3.10%, 95% CI: 2.98%--3.23%) and the rate was higher in 2021 compared with 2020 (3.3% vs 2.8%, P \< 0.001).
 
 A total of 51601 articles were published in journals with no impact factor.
-Papers that had a COI disclosure, funding statement and were registered tend to be published in journals with higher JIF (P = 0.003, \< 0.001, and = 0.007, respectively).
+Papers that had a COI disclosure, funding statement and were registered tend to be published in journals with lower JIF (P = 0.003, \< 0.001, and = 0.007, respectively).
 Detailed information is illustrated in Table 1.
 
+| Measurement       | Mean (SD)       |                | Median (IQR)  |               | P-value  |
+|-------------------|-----------------|----------------|---------------|---------------|----------|
+|                   | With            | Without        | With          | Without       |          |
+| COI disclosure    | 4.097 (2.0086)  | 7.442 (2.1377) | 3.709 (2.253) | 8.483 (0.000) | \< 0.001 |
+| Funding statement | 4.085 (1.9627)  | 5.670 (2.8597) | 3.706 (2.253) | 4.773 (4.900) | \< 0.001 |
+| Registration      | 3.808 (1.9459	) | 4.396 (2.1769) | 2.863 (1.973) | 4.003 (2.774) | \< 0.001 |
+| Data sharing      | 4.411 (2.3977)  | 4.251 (2.1221) | 4.057 (2.545) | 3.763 (2.519) | 0.03679  |
+| Code sharing      | 5.891 (4.5176)  | 4.239 (2.0812) | 4.262 (2.927) | 3.800 (2.519) | \< 0.001 |
+
+P-value based on the Wilcoxon rank sum test.
+
+COI: Conflict of interest; SD: Standard deviation; IQR: Inter-quartile range.
+
 Figure 1 shows the monthly trend for three transparency indicators.
-There was an inconsistent trend for the indicators.
-However, the COI disclosure percentage seemed to be more consistent and showed an increasing trend over time.
+According to the figure, there was an inconsistent trend for the indicators.
+However, the COI disclosure and data sharing percentages showed an increasing trend over time.
 

### 2022-03-09 rmarkdown-covid-19-transparent-analysis.Rmd
Manuscript and appendix files
Results are complete

--- 2022-02-26 covid_transparency_analysis.Rmd
@@
 
-The number and percentage of papers that shared data (preliminary):
+The number and percentage of papers that shared data:
 ```{r}
@@
 
-Comparing data sharing rate in two years (preliminary):
+Confidence interval for data sharing:
+```{r}
+
+kable(round(epi.prev(pos = length(opendata$is_open_data[opendata$is_open_data == TRUE]),
+         tested = nrow(opendata),
+         se = 0.758,
+         sp = 0.986)$ap, 
+      2))
+```
+
+Comparing data sharing rate in two years:
 ```{r}
@@
 
-The number and percentage of papers that shared code (preliminary):
+Chi-square test:
+```{r}
+chisq.test(opendata$is_open_data, opendata$year_firstpub)
+```
+
+The number and percentage of papers that shared code:
 ```{r}
@@
 
-Comparing code sharing rate in two years (preliminary):
+Confidence interval for sharing code:
+```{r}
+
+kable(round(epi.prev(pos = length(opendata$is_open_code[opendata$is_open_code == TRUE]),
+         tested = nrow(opendata),
+         se = 0.587,
+         sp = 0.997)$ap, 
+      2))
+```
+
+Comparing sharing code rate in two years:
 ```{r}
 kable(round(prop.table(table(opendata$is_open_code, opendata$year_firstpub), 2)*100, 1))
+```
+
+Chi-square test:
+```{r}
+chisq.test(opendata$is_open_code, opendata$year_firstpub)
 ```
@@
 ```
+
 JIF based on COI disclosure:
-
 ```{r}
@@
 
+JIF based on funding disclosure:
 ```{r}
@@
 
+JIF based on registration:
 ```{r}
@@
 
+JIF based on data sharing:
+```{r}
+kable(db_jif %>% 
+  group_by(is_open_data) %>% 
+  summarise(Mean = round(mean(X2020.JIF, na.rm = T), 3),
+            SD = round(sd(X2020.JIF, na.rm = T), 4),
+            Median = median(X2020.JIF, na.rm = T),
+            IQR = IQR(X2020.JIF, na.rm = T)))
+```
+
+Test for normality:
+```{r}
+set.seed(10)
+
+with(db_jif, shapiro.test(X2020.JIF[is_open_data == TRUE]))
+with(db_jif, shapiro.test(sample(X2020.JIF[is_open_data == FALSE], 5000)))
+```
+
+Data is not normal, hence we will use unpaired two-samples Wilcoxon test.
+```{r}
+wilcox.test(X2020.JIF~is_open_data, data = db_jif, exact = FALSE)
+```
+
+JIF based on code sharing:
+```{r}
+kable(db_jif %>% 
+  group_by(is_open_code) %>% 
+  summarise(Mean = round(mean(X2020.JIF, na.rm = T), 3),
+            SD = round(sd(X2020.JIF, na.rm = T), 4),
+            Median = median(X2020.JIF, na.rm = T),
+            IQR = IQR(X2020.JIF, na.rm = T)))
+```
+
+Test for normality:
+```{r}
+set.seed(10)
+
+with(db_jif, shapiro.test(X2020.JIF[is_open_code == TRUE]))
+with(db_jif, shapiro.test(sample(X2020.JIF[is_open_code == FALSE], 5000)))
+```
+
+Data is not normal, hence we will use unpaired two-samples Wilcoxon test.
+```{r}
+wilcox.test(X2020.JIF~is_open_code, data = db_jif, exact = FALSE)
+```
+
 ## Monthly trend
@@
             fund_true = sum(is_fund_pred == TRUE),
-            reg_true = sum(is_register_pred == TRUE))
-
-summarybyMonth %>%
-  ggplot(aes(x = month_firstpub, y = counts)) + 
-  geom_col(position = position_dodge()) +
-  facet_grid(year_firstpub ~ .) +
-  geom_bar(stat = "identity") +
-  scale_x_discrete(name="Month", limits = 1:12, labels = 1:12)
+            reg_true = sum(is_register_pred == TRUE),
+            data_true = sum(is_open_data == TRUE),
+            code_true = sum(is_open_code == TRUE))
 
@@
   theme_minimal() +
-  theme(legend.title = element_blank()) +
-  labs(title = "Conflict of Interest disclosure")
+  theme(legend.title = element_blank(), 
+        plot.title = element_text(hjust = 0.5)) +
+  labs(title = "Conflict of Interest disclosure",
+       y = "Number of articles") +
+        scale_fill_discrete(labels = c("Without", "With"),
+                            guide = guide_legend(reverse=TRUE))
 
@@
   theme_minimal() +
-  theme(legend.title = element_blank()) +
-  labs(title = "Funding statement")
+  theme(legend.title = element_blank(), 
+        plot.title = element_text(hjust = 0.5)) +
+  labs(title = "Funding statement",
+       y = "Number of articles") +
+        scale_fill_discrete(labels = c("Without", "With"),
+                            guide = guide_legend(reverse=TRUE))
 
@@
   theme_minimal() +
-  theme(legend.title = element_blank()) +
-  labs(title = "Registration statement")
+  theme(legend.title = element_blank(), 
+        plot.title = element_text(hjust = 0.5)) +
+  labs(title = "Registration statement",
+       y = "Number of articles") +
+        scale_fill_discrete(labels = c("Without", "With"),
+                            guide = guide_legend(reverse=TRUE))
+
+data_summarybyMonth <- opendata %>%
+  group_by(month_firstpub, year_firstpub, is_open_data) %>%
+  summarise(counts = n())
+
+data_plot <- data_summarybyMonth %>%
+  ggplot(aes(x = month_firstpub, y = counts, fill = is_open_data)) + 
+  geom_col(position = position_dodge()) +
+  facet_grid(. ~ year_firstpub) +
+  geom_bar(stat = "identity") +
+  scale_x_discrete(name="Month", limits = 1:12, labels = 1:12) +
+  theme_minimal() +
+  theme(legend.title = element_blank(), 
+        plot.title = element_text(hjust = 0.5)) +
+  labs(title = "Data sharing",
+       y = "Number of articles") +
+        scale_fill_discrete(labels = c("Without", "With"),
+                            guide = guide_legend(reverse=TRUE))
+
+code_summarybyMonth <- opendata %>%
+  group_by(month_firstpub, year_firstpub, is_open_code) %>%
+  summarise(counts = n())
+
+code_plot <- code_summarybyMonth %>%
+  ggplot(aes(x = month_firstpub, y = counts, fill = is_open_code)) + 
+  geom_col(position = position_dodge()) +
+  facet_grid(. ~ year_firstpub) +
+  geom_bar(stat = "identity") +
+  scale_x_discrete(name="Month", limits = 1:12, labels = 1:12) +
+  theme_minimal() +
+  theme(legend.title = element_blank(), 
+        plot.title = element_text(hjust = 0.5)) +
+  labs(title = "Code sharing",
+       y = "Number of articles") +
+        scale_fill_discrete(labels = c("Without", "With"),
+                            guide = guide_legend(reverse=TRUE))
 
 library(ggpubr)
+
 figure <- ggarrange(coi_plot + rremove("xlab") + theme(axis.text.x=element_blank()), 
@@
 
-# tiff("Figure.tiff", width = 6, height = 7, units = "in", res = 300)
+figure <- ggarrange(coi_plot + rremove("xlab") + theme(axis.text.x=element_blank()), 
+                    fund_plot + rremove("xlab") + theme(axis.text.x=element_blank(), strip.text.x = element_blank()),
+                    reg_plot + rremove("xlab") + theme(axis.text.x=element_blank(), strip.text.x = element_blank()),
+                    data_plot + rremove("xlab") + theme(axis.text.x=element_blank(), strip.text.x = element_blank()),
+                    code_plot + theme(strip.text.x = element_blank()),
+                    ncol = 1, nrow = 5, vjust = 1, 
+                    align = "hv", common.legend = T, legend = "right")
+
+# tiff("Figure.tiff", width = 6, height = 11, units = "in", res = 300)
 figure

### 2022-03-09 rmarkdown-covid-19-transparent-analysis.Rmd
Update rmarkdown-covid-19-transparent-analysis.Rmd
added here::here() and some code cleaning

@@
 
-```{r, setup, include=FALSE}
-knitr::opts_knit$set(root.dir = '~/Documents/GitHub/covid-transparency/data')
-```
-
 # Loading required packages
@@
 ```{r}
-pacman::p_load(dplyr,
+pacman::p_load(tidyverse,
+               lubridate, 
+               here, 
                ggplot2,
                knitr)
+```
+
+```{r, setup, include=FALSE}
+# knitr::opts_knit$set(root.dir = '~/Documents/GitHub/covid-transparency/data')
 ```
@@
 ```{r}
-opendata <- read.csv("covid_transparency_opendata.csv")
+opendata <- read.csv(here("data", "covid_transparency_opendata.csv"))
 ```
@@
 ```{r}
-library(lubridate)
-
 opendata <- opendata %>%
-  mutate(year_firstpub = year(
-    as.POSIXlt(firstPublicationDate, 
+        mutate(year_firstpub = year(as.POSIXlt(firstPublicationDate,
                                                format = "%Y-%m-%d")),
-    month_firstpub = month(
-      as.POSIXlt(firstPublicationDate, 
-               format = "%Y-%m-%d")
-    )
-    )
+               month_firstpub = month(as.POSIXlt(firstPublicationDate,
+                                                 format = "%Y-%m-%d")))
 ```
@@
 ```{r}
-kable(head(table(opendata$journal) %>% as.data.frame() %>% arrange(desc(Freq)), 10))
+kable(head(table(opendata$journal) 
+           %>% as.data.frame() 
+           %>% arrange(desc(Freq)), 10))
+```
+
+Also, same table, but with all the journals
+
+```{r}
+opendata %>% 
+        mutate(journal = fct_lump_prop(journal, prop = 0.01) ) %>% 
+        count(journal) %>% 
+        arrange(desc(n)) %>% 
+        knitr::kable()
 ```

### 2022-03-16 covid_transparency_dataset_creation.Rmd
Changing all datasets and re-analyse
Due to a small error in the search, we did everything from the scratch again and created new accurate datasets. Then, I re-analysed the results. Results need to be more thorough BTW.

@@
 title: "Transparency indicators across the dentistry and oral health literature"
-author: Eero Raittio, Ahmad Sofi-Mahmudi, Erfan Shamsoddin, Sergio E Uribe, Julie Farmer
+author: Ahmad Sofi-Mahmudi, Eero Raittio, Sergio E Uribe
 date: "1/8/2022"
@@
 ```{r}
-pacman::p_load(tidyverse,
+pacman::p_load(dplyr,
                rtransparent, 
@@
 ```{r}
-COVID_19_keywords <- read.csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vS6ibaTIJ0-4MiOZrLHKwo-Ya3MjlGX4f-dPQwdmWwQZhRGi6C83K1I_WceUJPH9zGyBc4L-lSalET_/pub?output=csv")
+COVID_19_keywords <- read.csv("covid_keywords.csv")
 ```
@@
 
-## Load all open access dentistry and oral health papers indexed in PubMed to the end of 2021:
+## Load all open-access and EPMC COVID-19-related papers indexed in PubMed
+
+As in this step we should download a large file that includes the characteristics of papers of our query, to ease the process, we divided our query to five separate queries based on the date:
+1- first half of 2020;
+2- second half of 2020;
+3- first half of 2021;
+4- second half of 2021;
+5- the first two months of 2022.
 
@@
 
-db_2020_1 <- epmc_search(query = paste0("'(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2020-01-01 TO 2020-06-30]) AND ((IN_EPMC:y) OR (OPEN_ACCESS:y)) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")', "'"), limit = 100000, output = "parsed")
+db_2020_1 <- epmc_search(query = paste0("(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2020-01-01 TO 2020-06-30]) AND (OPEN_ACCESS:y) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")'), limit = 100000, output = "parsed")
 
-db_2020_1 <- epmc_search(query = paste0("'(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2020-01-01 TO 2020-06-30]) AND (OPEN_ACCESS:y) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")', "'"), limit = 100000, output = "parsed")
+total oa and non oa: 19467
 
-db_2020_2 <- epmc_search(query = paste0("'(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2020-07-01 TO 2020-12-31]) AND ((IN_EPMC:y) OR (OPEN_ACCESS:y)) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")', "'"), limit = 100000, output = "parsed")
+db_2020_2 <- epmc_search(query = paste0("(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2020-07-01 TO 2020-12-31]) AND (OPEN_ACCESS:y) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")'), limit = 100000, output = "parsed")
 
-db_2021_1 <- epmc_search(query = paste0("'(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2021-01-01 TO 2021-06-30]) AND ((IN_EPMC:y) OR (OPEN_ACCESS:y)) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")', "'"), limit = 100000, output = "parsed")
+total oa and non oa: 43107
 
-db_2020_2 <- epmc_search(query = paste0("'(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2021-07-01 TO 2021-12-31]) AND ((IN_EPMC:y) OR (OPEN_ACCESS:y)) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")', "'"), limit = 100000, output = "parsed")
+db_2021_1 <- epmc_search(query = paste0("(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2021-01-01 TO 2021-06-30]) AND (OPEN_ACCESS:y) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")'), limit = 100000, output = "parsed")
 
-db <- epmc_search(query = paste0("'(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2020-01-01 TO 2021-12-31]) AND ((IN_EPMC:y) OR (OPEN_ACCESS:y)) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")', "'"), limit = 1000000, output = "parsed")
+total oa and non oa: 50512
 
-db <- epmc_search(query = paste0("'(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2020-01-01 TO 2021-12-31]) AND (OPEN_ACCESS:y) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")', "'"), limit = 1000000, output = "parsed")
+db_2021_2 <- epmc_search(query = paste0("(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2021-07-01 TO 2021-12-31]) AND (OPEN_ACCESS:y) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")'), limit = 100000, output = "parsed")
 
-db <- epmc_search(
-  query = paste0(
-    "'(",
-    TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery,
-    ") ",
-    'AND (SRC:"MED") 
-    AND (LANG:"eng" OR LANG:"en" OR LANG:"us") 
-    AND (FIRST_PDATE:[2020-01-01 TO 2021-12-31])
-    AND ((IN_EPMC:y) OR (OPEN_ACCESS:y))
-    AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")',
-    "'"
-  ),
-  limit = 100000, output = "parsed"
-)
+total oa and non oa: 47975
+
+db_2022 <- epmc_search(query = paste0("(", TitleQuery, " OR ", KeywordQuery, " OR ", ResultQuery, ") ", 'AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2022-01-01 TO 2022-02-28]) AND (OPEN_ACCESS:y) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")'), limit = 100000, output = "parsed")
+
+total oa and non oa: 14272
+
 ```
+
+## We then do the following for each of them
 
@@
 ```{r}
-table(duplicated(db$pmid))
+table(duplicated(db_2020_1$pmid))
+table(duplicated(db_2020_2$pmid))
+table(duplicated(db_2021_1$pmid))
+table(duplicated(db_2021_2$pmid))
+table(duplicated(db_2022$pmid))
+
 ```
@@
 ```{r}
-db <- db %>% distinct(pmid, .keep_all = TRUE)
+db_2020_1 <- db_2020_1 %>% distinct(pmid, .keep_all = TRUE)
+db_2020_2 <- db_2020_2 %>% distinct(pmid, .keep_all = TRUE)
+db_2021_1 <- db_2021_1 %>% distinct(pmid, .keep_all = TRUE)
+db_2021_2 <- db_2021_2 %>% distinct(pmid, .keep_all = TRUE)
+db_2022 <- db_2022 %>% distinct(pmid, .keep_all = TRUE)
+
 ```
 
-Let's see the first six rows of the database:
+Creating a new column from pmcid column and removing "PMC" from the cells:
 ```{r}
-head(db)
+db_2020_1$pmcid_ <- gsub("PMC", "", as.character(db_2020_1$pmcid))
+db_2020_2$pmcid_ <- gsub("PMC", "", as.character(db_2020_2$pmcid))
+db_2021_1$pmcid_ <- gsub("PMC", "", as.character(db_2021_1$pmcid))
+db_2021_2$pmcid_ <- gsub("PMC", "", as.character(db_2021_2$pmcid))
+db_2022$pmcid_ <- gsub("PMC", "", as.character(db_2022$pmcid))
+
 ```
 
-Removing "PMC" from the cells
+Now, we make five folders for xml format articles:
 ```{r}
-db_2020_1$pmcid_ <- gsub("PMC", "", as.character(db_2020_1$pmcid))
-
-db_2020_2$pmcid_ <- gsub("PMC", "", as.character(db_2020_2$pmcid))
-
-db_2021_1$pmcid_ <- gsub("PMC", "", as.character(db_2021_1$pmcid))
-
-db_2021_2$pmcid_ <- gsub("PMC", "", as.character(db_2021_2$pmcid))
-
-db$pmcid_ <- gsub("PMC", "", as.character(db$pmcid))
+dir.create(c("pmc_2020_1", "pmc_2020_2", "pmc_2021_1", "pmc_2021_2", "pmc_2022"))
 ```
 
-Now, we make a folder for xml format articles and switch to that folder:
+Next, we download xmls in format accessible with metareadr To skip errors (i.e., The metadata format 'pmc' is not supported by the item or by the repository.), first define a new function:
+
 ```{r}
-dir.create("pmc")
-setwd("pmc")
+skipping_errors <- function(x) tryCatch(mt_read_pmcoa(x), error = function(e) e)
 ```
@@
 ```{r}
-db_2020_1 <- subset(db_2020_1, pmcid_!=c(7264604, 7194244, 7323088, 7300787, 7276714, 7262104, 7228371))
-db_2020_2 <- subset(db_2020_2, pmcid_!=c(7677857, 7164703, 7678408, 7486161, 7645956, 7486166, 7675701, 7405208, 7753328, 7575215, 7759253, 7568026, 7598225, 7416582, 7649313, 7492687, 7571788, 7675665, 7572235, 7457217))
-db_2021_1 <- subset(db_2021_1, pmcid_!=c(8013335, 8013644, 8250125, 8014725, 8251275, 7486807, 8420411, 8251218, 8251305, 7894077))
-db_2021_2 <- subset(db_2021_2, pmcid_!=8653067)
+setwd("pmc_2020_1")
+sapply(db_2020_1$pmcid_, skipping_errors)
 
-db <- subset(db_2021_2, pmcid_!=c(7264604, 8013335))
+setwd("../pmc_2020_2")
+sapply(db_2020_2$pmcid_, skipping_errors)
 
-sapply(db_2020_1$pmcid_, mt_read_pmcoa)
+setwd("../pmc_2021_1")
+sapply(db_2021_1$pmcid_, skipping_errors)
 
-sapply(db_2020_2$pmcid_, mt_read_pmcoa)
+setwd("../pmc_2021_2")
+sapply(db_2021_2$pmcid_, skipping_errors)
 
-sapply(db_2021_1$pmcid_, mt_read_pmcoa)
-
-sapply(db_2021_2$pmcid_, mt_read_pmcoa)
-
-sapply(db$pmcid_, mt_read_pmcoa)
-
-db <- subset(db, !(pmcid_ %in% c(8251275, 8251218, 7486161, 7228371, 8485571)))
+setwd("../pmc_2022")
+sapply(db_2022$pmcid_, skipping_errors)
 
@@
 ```{r}
-df <- data.table::rbindlist(results_table_all, fill = TRUE)
+df1 <- data.table::rbindlist(results_table_all_f1, fill = TRUE)
+df2 <- data.table::rbindlist(results_table_all_f2, fill = TRUE)
+df3 <- data.table::rbindlist(results_table_all_f3, fill = TRUE)
+df4 <- data.table::rbindlist(results_table_all_f4, fill = TRUE)
+df5 <- data.table::rbindlist(results_table_all_f5, fill = TRUE)
+
 ```
@@
 opendata <- merge(db, results_table_data, by = "pmid") %>% merge(df)
-View(opendata)
 ```
+
+Random sample of search results for validation of our approach (100 refs):
+```{r}
+set.seed(100)
+
+randomsample3 <- db3[sample(nrow(db3), 100), ]
+```

### 2022-03-16 rmarkdown-covid-19-transparent-analysis.Rmd
Changing all datasets and re-analyse
Due to a small error in the search, we did everything from the scratch again and created new accurate datasets. Then, I re-analysed the results. Results need to be more thorough BTW.

@@
 ```{r}
-kable(data.frame(hits_all = 92752, 
+kable(data.frame(hits_all = 175550, 
            hits_oa = nrow(opendata),
-           oa_percentage = round((nrow(opendata)/92752)*100, 2)))
+           oa_percentage = round((nrow(opendata)/175550)*100, 2)))
 ```
@@
 kable(opendata[which.max(opendata$citedByCount),] %>% 
-  select(pmid, pmcid, doi.x, title, authorString, journalTitle, pubYear))
+  select(citedByCount, pmid, pmcid, doi.x, title, authorString, journalTitle, pubYear))
 ```
@@
 ```{r}
-jif <- read.csv("jif_2020.csv")
+jif <- read.csv(here("data", "jif_2020.csv"))
 ```

### 2022-04-01 manuscript_covid_transparency.qmd
Update manuscript_covid_transparency.qmd
The first draft of the results with the figure

--- 2022-03-08 manuscript_covid_transparency.qmd
@@
 
-(COVID-19 query for title, keywords, and results) AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:\[2020-01-01 TO 2021-12-31\]) AND ((IN_EPMC:y) OR (OPEN_ACCESS:y)) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")'
+(COVID-19 query for title, keywords, and results) AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:\[2020-01-01 TO 2021-12-31\]) AND (OPEN_ACCESS:y) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")'
 
@@
 
-The total number of COVID-19-related papers (open access and non-open access) was 92752, of which full texts of 73042 papers (78.7%) were accessible via the EPMC.
-Of those, 33137 were published in 2020 and 39905 in 2021.
-However, as we could not retrieve 43 full texts from the EPMC database due to technical issues, our final sample included 72999 full-text articles (Appendix 3).
+The total number of COVID-19-related papers (open access and non-open access) was 175550, of which full texts of 133470 papers (76.0%) were accessible as open-access via the EPMC.
+However, 852 (0.6%) were not downloadable because of technical issues and thus were excluded from our analyses.
+Of those 132618 remaining articles, 47550 were published in 2020, 75857 in 2021, and 9211 in the first two months of 2022 (Appendix 3).
 
-These articles were published in 4892 journals of which the top five were PLoS ONE (n=6825), International Journal of Environmental Research and Public Health (n=2686), Medicine (n=2318), BMJ Open (n=1926), and eLife (n=942).
-The mean and the median of the number of citations to these articles were 3.1 (standard deviation (SD) = 21.76) and 0 (interquartile range (IQR) = 2).
-The highest cited article with 2915 citations was an article published in the Cell in 2020 [@Walls2020].
+These articles were published in 6084 journals of which the top five were International Journal of Environmental Research and Public Health (n=3711), PLoS ONE (n=3059), Scientific Reports (n=1717), Frontiers in Public Health (n=1351), and Frontiers in Psychology (n=1341).
+The mean and the median of the number of citations to these articles were 12.0 (standard deviation (SD) = 92.27) and 2 (interquartile range (IQR) = 7).
+The highest cited article with 15549 citations was an article published in the Lancet in 2020 [@huang2020].
 
-More than nine-tenth of the articles (n=67009) had a statement to disclose COI (91.79%, 95% CI: 91.59%--91.99%).
-Articles published in 2021 had a higher rate of COI disclosure (93.2% vs 90.1	%, P \< 0.001).
-More than eight-tenths of the references had a funding statement (n=60530) (82.92%, 95% CI: 82.64%--83.19%).
-Articles in 2021 had a funding statement more frequently than 2020 articles (83.9% vs 81.7%, P \< 0.001).
-Less than 1 in 6 of the articles (n=11457) were registered beforehand (15.69%, 95% CI: 15.43%--15.96%) and the rate was higher in 2021 compared with 2020 (16.2% vs 15.0%, P \< 0.001).
+About nine-tenth of the articles (n=115405) had a statement to disclose COI (87.0%, 95% CI: 86.8%--87.2%).
+Articles published in 2022 had a higher rate of COI disclosure (91.4% vs 89.1% for 2021, and 82.9% for 2020, P \< 0.001).
+More than two-thirds of the references had a funding statement (n=91425) (68.9%, 95% CI: 68.7%--69.2%).
+Articles in 2022 had a funding statement more frequently than two other years (79.0% vs 73.1% for 2021, and 60.4% for 2020, P \< 0.001).
+Less than 1 in 25 of the articles (n=4708) were registered beforehand (3.6%, 95% CI: 3.5%--3.7%) and the rate was higher in 2022 (5.4% vs 4.2% for 2021, and 2.2% for 2020, P \< 0.001).
 
-Regarding data and code availability, as we chose the software to have low specificity to be as sensitive as possible, we got some false positives.
-Less than 1 in 7 articles had shared their data (n=9493) (13.00%, 95% CI: %--%).
-Articles in 2021 had shared data more frequently than 2020 articles (13.2% vs 12.8%, P \< 0.001).
-Less than 1 in 30 of the articles (n=2264) had shared their code (3.10%, 95% CI: %--%) and the rate was higher in 2021 compared with 2020 (3.3% vs 2.8%, P \< 0.001).
+Regarding data and code availability, less than 1 in 11 articles had shared their data (n=11552) (8.7%, 95% CI: 8.6%--8.9%).
+Articles in 2022 had shared data more frequently (11.9% vs 9.6% for 2021, and 6.7% for 2020, P \< 0.001).
+Less than 1 in 30 of the articles (n=4055) had shared their code (3.1%, 95% CI: 3.0%--3.2%) and the rate was higher in 2022 (4.0% vs 3.5% for 2021, and 2.2 for 2020, P \< 0.001).
 
-A total of 51601 articles were published in journals with no impact factor.
-Papers that had a COI disclosure, funding statement and were registered tend to be published in journals with higher JIF (P = 0.003, \< 0.001, and = 0.007, respectively).
+A total of 38383 articles were published in journals with a Journal Impact Factor (JIF).
+Papers that had a COI disclosure, funding statement and were registered tend to be published in journals with lower JIF (P = 0.003, \< 0.001, and = 0.007, respectively).
 Detailed information is illustrated in Table 1.
 
+Table 1. Citations and journal impact factor by transparency indicators.
+
+|                    | Citations to article |         |          | Journal Impact Factor |            |          |
+|-----------|-----------|-----------|-----------|:----------|:----------|-----------|
+| Measurement        | Median (IQR)         |         | P-Value  | Median (IQR)          |            | P-value  |
+|                    | With                 | Without |          | With                  | Without    |          |
+| COI disclosure     | 2 (6)                | 2 (7)   | \< 0.001 | 4.1 (2.31)            | 7.2 (5.12) | \< 0.001 |
+| Funding disclosure | 2 (6)                | 2 (8)   | \< 0.001 | 4.1 (2.32)            | 4.1 (2.94) | \< 0.001 |
+| Registration       | 1 (5)                | 2 (6)   | \< 0.001 | 3.8 (2.38)            | 4.2 (2.44) | \< 0.001 |
+| Data sharing       | 2 (8)                | 2 (6)   | \< 0.001 | 4.0 (2.98)            | 4.1 (2.36) | \< 0.001 |
+| Code sharing       | 2 (8)                | 2 (6)   | 0.70     | 4.4 (4.41)            | 4.1 (2.32) | 0.08     |
+
+: P-value based on the Wilcoxon rank sum test.\
+COI: Conflict of interest; SD: Standard deviation; IQR: Inter-quartile range.
+
+Over half of the papers had two transparency indicators (n=74085, 55.9%).
+Less than 0.1% (n=45) had all five indicators while 9.0% (n=11890) had no indicators.
+Among papers with two transparency indicators, the majority had COI and funding disclosures (99.3% and 97.9%, respectively).
+
 Figure 1 shows the monthly trend for three transparency indicators.
-There was an inconsistent trend for the indicators.
-However, the COI disclosure percentage seemed to be more consistent and showed an increasing trend over time.
+According to the figure, all indicators showed increasing monthly trend.
+However, the increasing trends for COI and funding disclosures were steeper compared to three other indicators.
 

### 2022-04-01 rmarkdown-covid-19-transparent-analysis.Rmd
The first draft of the results with the figure

@@
 title: "Transparency of COVID-19-related research - Analysis"
-author: Ahmad Sofi-Mahmudi, Eero Raittio
+author: Ahmad Sofi-Mahmudi, Eero Raittio, Sergio E. Uribe
 date: "2022-02-11"
@@
                ggplot2,
-               knitr)
-```
-
-```{r, setup, include=FALSE}
-# knitr::opts_knit$set(root.dir = '~/Documents/GitHub/covid-transparency/data')
+               knitr,
+               here)
 ```
@@
 kable(data.frame(hits_all = 175550, 
-           hits_oa = nrow(opendata),
-           oa_percentage = round((nrow(opendata)/175550)*100, 2)))
+           hits_oa = 133470,
+           oa_percentage = round((133470/175550)*100, 1)))
+```
+
+Undownloadable papers:
+```{r}
+c(number = 133470 - nrow(opendata), 
+  percentage = round((133470 - nrow(opendata))/133470*100, 1))
 ```
@@
   number = length(opendata$is_coi_pred[opendata$is_coi_pred == TRUE]),
-           percentage = round(length(opendata$is_coi_pred[opendata$is_coi_pred == TRUE])/nrow(opendata)*100, 2)
+           percentage = round(length(opendata$is_coi_pred[opendata$is_coi_pred == TRUE])/nrow(opendata)*100, 1)
            ))
@@
          sp = 0.995)$ap, 
-      2))
+      1))
 ```
 
-Comparing CoI disclosure rate in two years:
+Comparing CoI disclosure rate in three years:
 ```{r}
@@
 kable(data.frame(number = length(opendata$is_fund_pred[opendata$is_fund_pred == TRUE]),
-           percentage = round(length(opendata$is_fund_pred[opendata$is_fund_pred == TRUE])/nrow(opendata)*100, 2)
+           percentage = round(length(opendata$is_fund_pred[opendata$is_fund_pred == TRUE])/nrow(opendata)*100, 1)
            ))
@@
          sp = 0.981)$ap, 
-      2))
+      1))
 ```
 
-Comparing funding statement rate in two years:
+Comparing funding statement rate in three years:
 ```{r}
@@
 kable(data.frame(number = length(opendata$is_register_pred[opendata$is_register_pred == TRUE]),
-           percentage = round(length(opendata$is_register_pred[opendata$is_register_pred == TRUE])/nrow(opendata)*100, 2)
+           percentage = round(length(opendata$is_register_pred[opendata$is_register_pred == TRUE])/nrow(opendata)*100, 1)
            ))
@@
          sp = 0.997)$ap, 
-      2))
+      1))
 ```
 
-Comparing registration rate in two years:
+Comparing registration rate in three years:
 ```{r}
@@
 kable(data.frame(number = length(opendata$is_open_data[opendata$is_open_data == TRUE]),
-           percentage = round(length(opendata$is_open_data[opendata$is_open_data == TRUE])/nrow(opendata)*100, 2)
+           percentage = round(length(opendata$is_open_data[opendata$is_open_data == TRUE])/nrow(opendata)*100, 1)
            ))
@@
          sp = 0.986)$ap, 
-      2))
+      1))
 ```
 
-Comparing data sharing rate in two years:
+Comparing data sharing rate in three years:
 ```{r}
@@
 kable(data.frame(number = length(opendata$is_open_code[opendata$is_open_code == TRUE]),
-           percentage = round(length(opendata$is_open_code[opendata$is_open_code == TRUE])/nrow(opendata)*100, 2)
+           percentage = round(length(opendata$is_open_code[opendata$is_open_code == TRUE])/nrow(opendata)*100, 1)
            ))
@@
          sp = 0.997)$ap, 
-      2))
+      1))
 ```
 
-Comparing sharing code rate in two years:
+Comparing sharing code rate in three years:
 ```{r}
@@
 
-Total number of papers published in journals with no JIF:
+Total number of papers published in journals with a JIF:
 ```{r}
-nrow(opendata) - nrow(db_jif)
+nrow(db_jif)
 ```
@@
   group_by(is_coi_pred) %>% 
-  summarise(Mean = round(mean(X2020.JIF, na.rm = T),3),
-            SD = round(sd(X2020.JIF, na.rm = T),4),
-            Median = median(X2020.JIF, na.rm = T),
-            IQR = IQR(X2020.JIF, na.rm = T)))
+  summarise(Median = round(median(X2020.JIF, na.rm = T), 1),
+            IQR = round(IQR(X2020.JIF, na.rm = T), 2)))
+
+```
+
+Citation based on COI disclosure:
+```{r}
+kable(db_jif %>% 
+  group_by(is_coi_pred) %>% 
+  summarise(Median = round(median(citedByCount, na.rm = T), 1),
+            IQR = round(IQR(citedByCount, na.rm = T), 2)))
 ```
@@
 with(db_jif, shapiro.test(X2020.JIF[is_coi_pred == FALSE]))
+
+with(db_jif, shapiro.test(sample(citedByCount[is_coi_pred == TRUE], 5000)))
+with(db_jif, shapiro.test(citedByCount[is_coi_pred == FALSE]))
 ```
@@
 wilcox.test(X2020.JIF~is_coi_pred, data = db_jif, exact = FALSE)
+
+wilcox.test(citedByCount~is_coi_pred, data = db_jif, exact = FALSE)
 ```
@@
   group_by(is_fund_pred) %>% 
-  summarise(Mean = round(mean(X2020.JIF, na.rm = T), 3),
-            SD = round(sd(X2020.JIF, na.rm = T), 4),
-            Median = median(X2020.JIF, na.rm = T),
-            IQR = IQR(X2020.JIF, na.rm = T)))
+  summarise(Median = round(median(X2020.JIF, na.rm = T), 1),
+            IQR = round(IQR(X2020.JIF, na.rm = T), 2)))
+
+```
+
+Citation based on funding disclosure:
+```{r}
+kable(db_jif %>% 
+  group_by(is_fund_pred) %>% 
+  summarise(Median = round(median(citedByCount, na.rm = T), 1),
+            IQR = round(IQR(citedByCount, na.rm = T), 2)))
 ```
@@
 with(db_jif, shapiro.test(sample(X2020.JIF[is_fund_pred == TRUE], 5000)))
-with(db_jif, shapiro.test(X2020.JIF[is_fund_pred == FALSE]))
+with(db_jif, shapiro.test(sample(X2020.JIF[is_fund_pred == FALSE], 5000)))
+
+with(db_jif, shapiro.test(sample(citedByCount[is_fund_pred == TRUE], 5000)))
+with(db_jif, shapiro.test(sample(citedByCount[is_fund_pred == FALSE], 5000)))
 ```
@@
 wilcox.test(X2020.JIF~is_fund_pred, data = db_jif, exact = FALSE)
+
+wilcox.test(citedByCount~is_fund_pred, data = db_jif, exact = FALSE)
 ```
@@
   group_by(is_register_pred) %>% 
-  summarise(Mean = round(mean(X2020.JIF, na.rm = T), 3),
-            SD = round(sd(X2020.JIF, na.rm = T), 4),
-            Median = median(X2020.JIF, na.rm = T),
-            IQR = IQR(X2020.JIF, na.rm = T)))
+  summarise(Median = round(median(X2020.JIF, na.rm = T), 1),
+            IQR = round(IQR(X2020.JIF, na.rm = T), 2)))
+
+```
+
+Citation based on funding disclosure:
+```{r}
+kable(db_jif %>% 
+  group_by(is_register_pred) %>% 
+  summarise(Median = round(median(citedByCount, na.rm = T), 1),
+            IQR = round(IQR(citedByCount, na.rm = T), 2)))
 ```
@@
 with(db_jif, shapiro.test(sample(X2020.JIF[is_register_pred == FALSE], 5000)))
+
+with(db_jif, shapiro.test(citedByCount[is_register_pred == TRUE]))
+with(db_jif, shapiro.test(sample(citedByCount[is_register_pred == FALSE], 5000)))
 ```
@@
 wilcox.test(X2020.JIF~is_register_pred, data = db_jif, exact = FALSE)
+
+wilcox.test(citedByCount~is_register_pred, data = db_jif, exact = FALSE)
 ```
@@
   group_by(is_open_data) %>% 
-  summarise(Mean = round(mean(X2020.JIF, na.rm = T), 3),
-            SD = round(sd(X2020.JIF, na.rm = T), 4),
-            Median = median(X2020.JIF, na.rm = T),
-            IQR = IQR(X2020.JIF, na.rm = T)))
+  summarise(Median = round(median(X2020.JIF, na.rm = T), 1),
+            IQR = round(IQR(X2020.JIF, na.rm = T), 2)))
+
+```
+
+Citation based on funding disclosure:
+```{r}
+kable(db_jif %>% 
+  group_by(is_open_data) %>% 
+  summarise(Median = round(median(citedByCount, na.rm = T), 1),
+            IQR = round(IQR(citedByCount, na.rm = T), 2)))
 ```
@@
 
-with(db_jif, shapiro.test(X2020.JIF[is_open_data == TRUE]))
+with(db_jif, shapiro.test(sample(X2020.JIF[is_open_data == TRUE], 5000)))
 with(db_jif, shapiro.test(sample(X2020.JIF[is_open_data == FALSE], 5000)))
+
+with(db_jif, shapiro.test(sample(citedByCount[is_open_data == TRUE], 5000)))
+with(db_jif, shapiro.test(sample(citedByCount[is_open_data == FALSE], 5000)))
 ```
@@
 wilcox.test(X2020.JIF~is_open_data, data = db_jif, exact = FALSE)
+
+wilcox.test(citedByCount~is_open_data, data = db_jif, exact = FALSE)
 ```
@@
   group_by(is_open_code) %>% 
-  summarise(Mean = round(mean(X2020.JIF, na.rm = T), 3),
-            SD = round(sd(X2020.JIF, na.rm = T), 4),
-            Median = median(X2020.JIF, na.rm = T),
-            IQR = IQR(X2020.JIF, na.rm = T)))
+  summarise(Median = round(median(X2020.JIF, na.rm = T), 1),
+            IQR = round(IQR(X2020.JIF, na.rm = T), 2)))
+
+```
+
+Citation based on funding disclosure:
+```{r}
+kable(db_jif %>% 
+  group_by(is_open_code) %>% 
+  summarise(Median = round(median(citedByCount, na.rm = T), 1),
+            IQR = round(IQR(citedByCount, na.rm = T), 2)))
 ```
@@
 with(db_jif, shapiro.test(sample(X2020.JIF[is_open_code == FALSE], 5000)))
+
+with(db_jif, shapiro.test(citedByCount[is_open_code == TRUE]))
+with(db_jif, shapiro.test(sample(citedByCount[is_open_code == FALSE], 5000)))
 ```
@@
 wilcox.test(X2020.JIF~is_open_code, data = db_jif, exact = FALSE)
+
+wilcox.test(citedByCount~is_open_code, data = db_jif, exact = FALSE)
+```
+
+## Sum of five indicators
+```{r}
+opendata <- opendata %>% mutate(sumOfIndicators = rowSums(opendata %>% select(is_coi_pred, is_register_pred, is_fund_pred, is_open_data, is_open_code)))
+```
+
+Number of papers with each number of TRUE indicators:
+```{r}
+c(five_ind = nrow(filter(opendata, sumOfIndicators == 5)),
+  four_ind = nrow(filter(opendata, sumOfIndicators == 4)),
+  three_ind = nrow(filter(opendata, sumOfIndicators == 3)),
+  two_ind = nrow(filter(opendata, sumOfIndicators == 2)),
+  one_ind = nrow(filter(opendata, sumOfIndicators == 1)),
+  zero_ind = nrow(filter(opendata, sumOfIndicators == 0)))
+```
+
+Percentage of papers with each number of TRUE indicators:
+```{r}
+
+c(five_ind = round(nrow(filter(opendata, sumOfIndicators == 5))/nrow(opendata)*100, 1),
+  four_ind = round(nrow(filter(opendata, sumOfIndicators == 4))/nrow(opendata)*100, 1),
+  three_ind = round(nrow(filter(opendata, sumOfIndicators == 3))/nrow(opendata)*100, 1),
+  two_ind = round(nrow(filter(opendata, sumOfIndicators == 2))/nrow(opendata)*100, 1),
+  one_ind = round(nrow(filter(opendata, sumOfIndicators == 1))/nrow(opendata)*100, 1),
+  zero_ind = round(nrow(filter(opendata, sumOfIndicators == 0))/nrow(opendata)*100, 1))
+```
+
+To know which of the indicators were TRUE in papers with two TRUE indicators:
+```{r}
+
+nrow_ <- nrow(filter(opendata, sumOfIndicators == 2))
+
+c(coi = round(nrow(filter(opendata, sumOfIndicators == 2 & is_coi_pred))/nrow_*100, 1),
+  fund = round(nrow(filter(opendata, sumOfIndicators == 2 & is_fund_pred))/nrow_*100, 1),
+  reg = round(nrow(filter(opendata, sumOfIndicators == 2 & is_register_pred))/nrow_*100, 1),
+  data = round(nrow(filter(opendata, sumOfIndicators == 2 & is_open_data))/nrow_*100, 1),
+  code = round(nrow(filter(opendata, sumOfIndicators == 2 & is_open_code))/nrow_*100, 1))
 ```
@@
 ## Monthly trend
-Now, drawing monthly trend for each year:
+Figure 1:
 ```{r}
-summarybyMonth <- opendata %>%
-  group_by(month_firstpub, year_firstpub) %>%
-  summarise(counts = n(),
-            coi_true = sum(is_coi_pred == TRUE),
-            fund_true = sum(is_fund_pred == TRUE),
-            reg_true = sum(is_register_pred == TRUE),
-            data_true = sum(is_open_data == TRUE),
-            code_true = sum(is_open_code == TRUE))
+library(tibble)
 
-coi_summarybyMonth <- opendata %>%
-  group_by(month_firstpub, year_firstpub, is_coi_pred) %>%
-  summarise(counts = n())
+proportions <- opendata %>%
+        summarise("COI Disclosure" = sum(is_coi_pred == TRUE),
+                  "Funding disclosure" = sum(is_fund_pred == TRUE),
+                  "Protocol registration" = sum(is_register_pred == TRUE),
+                  "Data sharing" = sum(is_open_data == TRUE),
+                  "Code sharing" = sum(is_open_code == TRUE)) %>%
+        t() %>%
+        as.data.frame() %>%
+        rownames_to_column(var = "indicator") %>%
+        mutate(percentage = round(V1/nrow(opendata)*100, 1))
 
-coi_plot <- coi_summarybyMonth %>%
-  ggplot(aes(x = month_firstpub, y = counts, fill = is_coi_pred)) + 
-  geom_col(position = position_dodge()) +
-  facet_grid(. ~ year_firstpub) +
-  geom_bar(stat = "identity") +
-  scale_x_discrete(name="Month", limits = 1:12, labels = 1:12) +
-  theme_minimal() +
-  theme(legend.title = element_blank(), 
-        plot.title = element_text(hjust = 0.5)) +
-  labs(title = "Conflict of Interest disclosure",
-       y = "Number of articles") +
-        scale_fill_discrete(labels = c("Without", "With"),
-                            guide = guide_legend(reverse=TRUE))
+p1 <- proportions %>% ggplot() +
+        aes(x = reorder(indicator, V1),
+            y = V1,
+            fill = indicator,
+            col = indicator) +
+        geom_col(alpha = 0.75) +
+        geom_text(aes(label = percentage), hjust = -0.1, size = 3) +
+        scale_y_continuous(
+                name = "\nNumber of articles (in ten thousands)",
+                breaks = seq(0, 10E4, 5E4),
+                labels = seq(0, 10, 5),
+                expand = expansion(mult = c(0, 0.2))
+                ) +
+        coord_flip() +
+        xlab(NULL) +
+        theme(
+                legend.position = "none", 
+                panel.grid.major.y = element_blank()
+        )
+```
 
-fund_summarybyMonth <- opendata %>%
-  group_by(month_firstpub, year_firstpub, is_fund_pred) %>%
-  summarise(counts = n())
+For the line grapht, first, we add the real publication year and month to the datasets.
+The real publication year/month is the year/month the paper was first appear online stored in firstPublicationDate column.
+```{r}
+opendata$YearMonth <- format(as.Date(opendata$firstPublicationDate), "%Y-%m")
+```
 
-fund_plot <- fund_summarybyMonth %>%
-  ggplot(aes(x = month_firstpub, y = counts, fill = is_fund_pred)) + 
-  geom_col(position = position_dodge()) +
-  facet_grid(. ~ year_firstpub) +
-  geom_bar(stat = "identity") +
-  scale_x_discrete(name="Month", limits = 1:12, labels = 1:12) +
-  theme_minimal() +
-  theme(legend.title = element_blank(), 
-        plot.title = element_text(hjust = 0.5)) +
-  labs(title = "Funding statement",
-       y = "Number of articles") +
-        scale_fill_discrete(labels = c("Without", "With"),
-                            guide = guide_legend(reverse=TRUE))
-
-reg_summarybyMonth <- opendata %>%
-  group_by(month_firstpub, year_firstpub, is_register_pred) %>%
-  summarise(counts = n())
-
-reg_plot <- reg_summarybyMonth %>%
-  ggplot(aes(x = month_firstpub, y = counts, fill = is_register_pred)) + 
-  geom_col(position = position_dodge()) +
-  facet_grid(. ~ year_firstpub) +
-  geom_bar(stat = "identity") +
-  scale_x_discrete(name="Month", limits = 1:12, labels = 1:12) +
-  theme_minimal() +
-  theme(legend.title = element_blank(), 
-        plot.title = element_text(hjust = 0.5)) +
-  labs(title = "Registration statement",
-       y = "Number of articles") +
-        scale_fill_discrete(labels = c("Without", "With"),
-                            guide = guide_legend(reverse=TRUE))
-
-data_summarybyMonth <- opendata %>%
-  group_by(month_firstpub, year_firstpub, is_open_data) %>%
-  summarise(counts = n())
-
-data_plot <- data_summarybyMonth %>%
-  ggplot(aes(x = month_firstpub, y = counts, fill = is_open_data)) + 
-  geom_col(position = position_dodge()) +
-  facet_grid(. ~ year_firstpub) +
-  geom_bar(stat = "identity") +
-  scale_x_discrete(name="Month", limits = 1:12, labels = 1:12) +
-  theme_minimal() +
-  theme(legend.title = element_blank(), 
-        plot.title = element_text(hjust = 0.5)) +
-  labs(title = "Data sharing",
-       y = "Number of articles") +
-        scale_fill_discrete(labels = c("Without", "With"),
-                            guide = guide_legend(reverse=TRUE))
-
-code_summarybyMonth <- opendata %>%
-  group_by(month_firstpub, year_firstpub, is_open_code) %>%
-  summarise(counts = n())
-
-code_plot <- code_summarybyMonth %>%
-  ggplot(aes(x = month_firstpub, y = counts, fill = is_open_code)) + 
-  geom_col(position = position_dodge()) +
-  facet_grid(. ~ year_firstpub) +
-  geom_bar(stat = "identity") +
-  scale_x_discrete(name="Month", limits = 1:12, labels = 1:12) +
-  theme_minimal() +
-  theme(legend.title = element_blank(), 
-        plot.title = element_text(hjust = 0.5)) +
-  labs(title = "Code sharing",
-       y = "Number of articles") +
-        scale_fill_discrete(labels = c("Without", "With"),
-                            guide = guide_legend(reverse=TRUE))
-
+```{r}
 library(ggpubr)
 
-figure <- ggarrange(coi_plot + rremove("xlab") + theme(axis.text.x=element_blank()), 
-                    fund_plot + rremove("xlab") + theme(axis.text.x=element_blank(), strip.text.x = element_blank()),
-                    reg_plot + theme(strip.text.x = element_blank()),
-                    ncol = 1, nrow = 3, vjust = 1, 
-                    align = "hv", common.legend = T, legend = "right")
+indicator_by_year <- 
+        opendata %>% 
+        select(YearMonth,
+               is_coi_pred,
+               is_fund_pred,
+               is_register_pred,
+               is_open_data,
+               is_open_code) %>%
+        gather("indicator", "value", -YearMonth) %>%
+        count(YearMonth, indicator, value) %>%
+        mutate(indicator = recode(indicator,
+                                  is_coi_pred = "COI Disclosure",
+                                  is_fund_pred = "Funding disclosure",
+                                  is_register_pred = "Protocol registration",
+                                  is_open_data = "Data sharing",
+                                  is_open_code = "Code sharing")) %>%
+        complete(indicator, value, YearMonth, fill = list(n = 0)) %>%
+        group_by(YearMonth, indicator) %>% 
+        mutate(p = n / sum(n)) %>%
+        filter(value) %>%
+        ungroup()
 
-figure <- ggarrange(coi_plot + rremove("xlab") + theme(axis.text.x=element_blank()), 
-                    fund_plot + rremove("xlab") + theme(axis.text.x=element_blank(), strip.text.x = element_blank()),
-                    reg_plot + rremove("xlab") + theme(axis.text.x=element_blank(), strip.text.x = element_blank()),
-                    data_plot + rremove("xlab") + theme(axis.text.x=element_blank(), strip.text.x = element_blank()),
-                    code_plot + theme(strip.text.x = element_blank()),
-                    ncol = 1, nrow = 5, vjust = 1, 
-                    align = "hv", common.legend = T, legend = "right")
+p2 <-
+        indicator_by_year %>% 
+        ggplot() +
+        aes(x = YearMonth, 
+            y = p,
+            group = indicator,
+            color = indicator) +
+        geom_line(size = 0.75) +
+        scale_y_continuous(limits = c(0, 1), 
+                           labels = scales::percent) +
+        scale_color_discrete(name = NULL) +
+        scale_fill_discrete(breaks = c("COI Disclosure",
+                                       "Funding disclosure",
+                                       "Protocol registration",
+                                       "Data sharing",
+                                       "Code sharing")) +
+        labs(y = "Proportion of articles (%)\n", 
+             x = "\nMonth") +
+        theme(panel.grid.minor = element_blank(),
+              legend.position = c(0.2, 0.9),
+              axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1),
+              legend.key.size = unit(0.4, 'cm')
+              )
 
-# tiff("Figure.tiff", width = 6, height = 11, units = "in", res = 300)
+figure <- ggarrange(p1, p2,
+                    ncol = 2, nrow = 1, 
+                    align = "hv", common.legend = F)
+
+# tiff("Figure.tiff", width = 14, height = 7, units = "in", res = 300)
 figure
 #dev.off()
+
 ```

### 2022-04-01 manuscript_covid_transparency.qmd
discussion first edits

@@
 We also provided monthly trends for transparency indicators, to check whether there were any trends in transparency practices over time.
-We used Chi-square for comparing indicators between years and the Wilcoxon rank sum test to test whether there was a relationship between transparency indicators and JIFs.
+We used Chi-square for comparing indicators between years and the Wilcoxon rank sum test to test whether there was a relationship of transparency indicators with citations to articles and JIFs.
 The thorough system information for the analysis is available in Appendix 2.
@@
 |                    | Citations to article |         |          | Journal Impact Factor |            |          |
-|-----------|-----------|-----------|-----------|:----------|:----------|-----------|
+|--------------------|----------------------|---------|----------|:----------------------|:-----------|----------|
 | Measurement        | Median (IQR)         |         | P-Value  | Median (IQR)          |            | P-value  |
@@
 
+Adherence to transparency practices increased over time in COVID-19 related medical literature.
+Adherence to reporting COI disclosure was high throughout study period.
+Code sharing and protocol registration were rare, but improved little over study period.
+
+Adherence to transparency practices was similar in COVID-19 related literature than in other biomedical literature.
+
 ## Limitations

### 2022-04-02 manuscript_covid_transparency.qmd
discussion

@@
 
-Adherence to transparency practices was similar in COVID-19 related literature than in other biomedical literature.
+Adherence to transparency practices was at similar level in COVID-19 related literature than in other biomedical literature.
+This is surprising particularly when considering worldwide, remarkable and noble initiatives related to enhancing open science to tackle the pandemic.
+As early as in January 2020, over 100 organizations, including journals, publishers, funders, universities and other institutions, signed a statement that aimed to ensure free access to research data, tools and other information related to COVID-19 [@vuong2022; @coronavi].
+Later also other initiatives to support the goal emerged, for instance COVID-19 Open Research Dataset (CORD-19)[@cord-19], a free resource of over 280,000 articles about the COVID-19 virus.
+However, it is possible that the algorithms used here did not efficiently detect all different ways of sharing data and material emerged after the pandemic, because the algorithms were validated before the pandemic [@serghiou2021].
 
@@
 
-The study sample will be restricted to open access articles in the Europe PMC database which may not correspond to all COVID-19-related studies published in medical journals.
-However, as the majority of COVID-19-related papers are open access[@Kiley2020-ek], this will not diminish the strength of our interpretations considerably.
+Despite we used representative data and validated methods to investigate transparency practices in COVID-19-related research, our study has some weaknesses.
+The study sample was restricted to open access articles in the EPMC database which may not correspond to all COVID-19-related studies published in medical journals.
+At least in general, the differences in transparency practices between open access and non-open access articles from Europe PMC are rather similar [@wallach2018a].
+However, as the majority of COVID-19-related papers were available via EPMC (76%), this does not diminish the strength of our interpretations considerably.
 

### 2022-04-11 manuscript_covid_transparency.qmd
Update manuscript_covid_transparency.qmd

@@
 
+In comparison to other studies on the transparency of COVID-19 related research is difficult due to different methodology and scale of our research.
+However, it seems that data sharing was little less common than in our sample than what was the proportion detected in the begininning of the pandemic (https://doi.org/10.1007/s11192-021-03971-6).
+
 ## Limitations

### 2022-04-12 manuscript_covid_transparency.qmd
Update manuscript_covid_transparency.qmd

--- 2022-04-02 manuscript_covid_transparency.qmd
@@
 
+In comparison to other studies on the transparency of COVID-19 related research is difficult due to different methodology and scale of our research.
+However, it seems that data sharing was little less common than in our sample than what was the proportion detected in the begininning of the pandemic measured by the investigation of associated data in PubMed Central (https://doi.org/10.1007/s11192-021-03971-6).
+In addition, the proportion of studies in our stidy study was lower than in COVID-19 related preprints shared via medRxiv and bioRxiv (https://link.springer.com/article/10.1007/s11192-022-04346-1#ref-CR9 ).
+
 ## Limitations

### 2022-04-12 manuscript_covid_transparency.qmd
discussion

--- 2022-04-02 manuscript_covid_transparency.qmd
@@
 
-Adherence to transparency practices increased over time in COVID-19 related medical literature.
-Adherence to reporting COI disclosure was high throughout study period.
-Code sharing and protocol registration were rare, but improved little over study period.
+Adherence to transparency practices increased over time in COVID-19 related medical literature during 2020-2021.
+Adherence to reporting COI disclosure was high throughout study period, in addition most of articles had funding disclosure.
+Data sharing, code sharing and protocol registration were rare, but improved little over study period.
+Conflicts of interest disclosures seemed to be more common in articles published in lower impact journals, otherwise differences in transparency practices according to received citations and JIF were small.
 
@@
 However, it is possible that the algorithms used here did not efficiently detect all different ways of sharing data and material emerged after the pandemic, because the algorithms were validated before the pandemic [@serghiou2021].
+
+In comparison to other studies on the transparency of COVID-19 related research is difficult due to different methodology and scale of our research.
+However, it seems that data sharing was little less common than in our sample than what was the proportion detected in the begininning of the pandemic measured by the investigation of associated data in PubMed Central [@lucas-dominguez2021\].
+In addition, the proportion of studies which shared data in our study sample was lower than in COVID-19 related preprints shared via medRxiv and bioRxiv \[@strcic2022\].
+Lacking preregisteration, sharing code and conflicts of interest have go attention during the pandemic \[@Glasziou2020; @besançon2021; @luengo-oroz2020], however we are not aware of the investigation of transparency practices in these aspects in COVID-19 related research.
 
@@
 
+Transparency is crucial to ensure the credibility of science and enable its assessment.
+It has been argued that open science, including transparency practices such as the ones we investigated here, have helped to fight pandemic globally [@besançon2021; @vuong2022].
+However, your findings showed suboptimal adherence to data and code sharing and pre-registeration in COVID-19 related research, whilst majority of researchers had adhered to disclose funding and conflicts of interests related to their COVID-19 related medical research.
+Stronger commitment to open science practices, particularly to preregistration, data and code sharing, is needed from all stakeholders.
+
 # References

### 2022-04-13 manuscript_covid_transparency.qmd
intro, abstract and other minor edits

@@
 **Objective:** The recent pandemic has shown how crucial it is to have access to quality medical information and data that allows people to have confidence in healthcare decisions.
-We aimed to assess the use of transparency indicators (data availability, code availability, statements of protocol registration and conflicts of interest and funding disclosures) from open access COVID-19-related articles.
+We aimed to assess the use of transparency indicators (data availability, code availability, statements of protocol registration and conflicts of interest and funding disclosures) from open access Coronavirus disease 2019 (COVID-19) related articles.
 
@@
 
-**Results:** Of 72,999 available articles published in XX medical journals, XX% provided conflicts of interest disclosure and XX% funding disclosure and X% were preregistered.
-XX studies shared raw data (XX%) and XX studies shared code.
-Transparent practices were more common in articles published in journals with XX impact factors, and in 2020 than in 2021.
+**Results:** Of 72,999 available articles published in XX medical journals, 87.0%, (95% CI: 86.8%--87.2%) provided conflicts of interest disclosure and 68.9% (95% CI: 68.7%--69.2%) funding disclosure and 3.6%, (95% CI: 3.5%--3.7%) were preregistered.
+8.7% (95% CI: 8.6%--8.9%) of studies shared raw data and 3.1% (95% CI: 3.0%--3.2%) of studies shared code.
+Transparent practices were more common in articles published in 2021 than in 2020.
 
 **Conclusion:** While the majority of the papers had a COI disclosure, the prevalence of the other transparency practices was far from the acceptable level.
+Even surprisingly low, particularly in terms of data sharing, taking into account global initiatives to improve open science in COVID-19 related research.
 A much stronger commitment to open science practices, particularly to preregistration, data and code sharing, is needed from all stakeholders.
@@
 
-The foundation of science is to generate knowledge from reproducible findings (1).
-Because of the failure to reproduce previous research, the open science movement emerged in recent years (2).
-The movement's principal focus is on making science more accessible and trustworthy.
-Data and code sharing, protocol registration and funding and conflicts of interest (COI) disclosures are considered important features of open science and crucial when assessing the credibility of certain scientific findings (2,3).
-Science without credibility and quality would do a major human, societal and economic disservice (4).
+COVID-19 pandemic launched an incredibly fast and voluminous avalanche of scientific publications across the different fields of science.
+There has been discussion and many examples related to poor or suboptimal data, methods and other scientific practices in COVID-19-related studies [@Glasziou2020; @besançon2021; @luengo-oroz2020; @schwab2020\].
+However, in addition to fast and successful development of safe and effective vaccination, it is possible that transparent scientific practices, like data and code sharing, have helped to fight pandemic globally \[@besançon2021; @vuong2022\].
+That being said, comprehensive picture of transparency practices in COVID-19-related medical literature is still lacking.
+As an example, it is unknown whether open science initiatives related to COVID-19 pandemic \[@besançon2021; @vuong2022], for instance related to data sharing, led to higher number of articles sharing data in COVID-19 research compared to what has been seen in studies in general in biomedical literature [@serghiou2021].
 
-COVID-19 pandemic launched an incredibly fast and voluminous avalanche of scientific publications across the different fields of science.
-Related to the COVID-19 publications there has been a considerable amount of discussion and examples related to poor or suboptimal data, methods and other scientific practices (5--8).
-Unsurprisingly, there have been signs of fabricated data, poor data and methods and conflicts of interests in the scientific literature related to the COVID-19 pandemic, which has resulted in a great number of retracted papers (208 as of the end of January 2022)(9).
-Probably, if transparent scientific practices, like data and code sharing, would be more like the rule than the exception in biomedical research, pre- and post-publication evaluation of the credibility of studies could be much easier (10,11).
-
-The COVID-19 pandemic and its implications for dental services, oral health, oral health-related behaviors and dental education have received great attention from the dental research community.
-Nevertheless, to date, only one paper related to dentistry or published in dental has been retracted from the Journal of Oral and Maxillofacial Pathology (9).
-However, meta-research about reporting, methods or reproducibility in COVID-19-related dental research has been limited.
-Particularly sharing data with adequate quality would allow reanalyses of findings (10).
-
-Our aim was to assess the uptake of transparent scientific practices (data sharing, code sharing, COI disclosures, funding disclosures, and protocol registration) from open access full-text COVID-19-related articles published in dental journals available from the Europe PubMed Central (EPMC) database.
-We also assessed the FAIRness (Findable, Accessible, Interoperable, and Reusable) of data from those articles that have shared their data.
-We mapped transparency practices across publication years, publishers and journals.
+Our aim was to assess the adherence to transparent scientific practices (data sharing, code sharing, conflict of interest, COI, disclosures, funding disclosures, and protocol registration) from open access full-text COVID-19-related articles published in PubMed indexed journals from the Europe PubMed Central (EPMC) database.
+We mapped adherence to transparency practices across time, journal impact factor (JIF) and citations articles have received
 
@@
 
-We searched for open access COVID-19-related articles from medical journals available in the EPMC database.
+We searched for open access COVID-19-related PubMed-indexed articles available in the EPMC database.
 We restricted our search to papers in English and considered papers published from 2020-01-01 to 2021-12-31 with variants of COVID-19 keyword in their title, keywords, or results.
@@
 
--   conflicts of interest disclosures,
+-   COI disclosures,
 
@@
 
-We also calculated the proportion of articles available as open-access via the Europe PMC from the total number of COVID-19-related detected in the database.
+We also calculated the proportion of articles available as open-access full-texts via the Europe PMC from the total number of COVID-19-related articles in the database.
 
@@
 Data sharing, code sharing and protocol registration were rare, but improved little over study period.
-Conflicts of interest disclosures seemed to be more common in articles published in lower impact journals, otherwise differences in transparency practices according to received citations and JIF were small.
+COI disclosures seemed to be more common in articles published in lower impact journals, otherwise differences in transparency practices according to received citations and JIF were small.
 
-Adherence to transparency practices was at similar level in COVID-19 related literature than in other biomedical literature.
+Adherence to transparency practices was at similar level in COVID-19 related literature than in other biomedical literature analyzed with the same methods (see Fig 3 in [@serghiou2021\]).
 This is surprising particularly when considering worldwide, remarkable and noble initiatives related to enhancing open science to tackle the pandemic.
-As early as in January 2020, over 100 organizations, including journals, publishers, funders, universities and other institutions, signed a statement that aimed to ensure free access to research data, tools and other information related to COVID-19 [@vuong2022; @coronavi].
+As early as in January 2020, over 100 organizations, including journals, publishers, funders, universities and other institutions, signed a statement that aimed to ensure free access to research data, tools and other information related to COVID-19 \[@vuong2022; @coronavi].
 Later also other initiatives to support the goal emerged, for instance COVID-19 Open Research Dataset (CORD-19)[@cord-19], a free resource of over 280,000 articles about the COVID-19 virus.
@@
 
-In comparison to other studies on the transparency of COVID-19 related research is difficult due to different methodology and scale of our research.
-However, it seems that data sharing was little less common than in our sample than what was the proportion detected in the begininning of the pandemic measured by the investigation of associated data in PubMed Central [@lucas-dominguez2021\].
+In comparison to other studies on the transparency of COVID-19 related research is difficult due to different methodologies and scale of our research.
+However, it seems that data sharing was little less common than in our sample than what was the proportion detected in the beginning of the pandemic measured by the investigation of associated data in PubMed Central [@lucas-dominguez2021\].
 In addition, the proportion of studies which shared data in our study sample was lower than in COVID-19 related preprints shared via medRxiv and bioRxiv \[@strcic2022\].
-Lacking preregisteration, sharing code and conflicts of interest have go attention during the pandemic \[@Glasziou2020; @besançon2021; @luengo-oroz2020], however we are not aware of the investigation of transparency practices in these aspects in COVID-19 related research.
-
-## Limitations
+Lacking preregisteration, sharing code and COI have got attention during the pandemic \[@Glasziou2020; @besançon2021; @luengo-oroz2020], however we are not aware of the investigation of transparency practices in these aspects in COVID-19 related research in this scale in medical research.
 
@@
 
-# Conclusion
-
 Transparency is crucial to ensure the credibility of science and enable its assessment.
-It has been argued that open science, including transparency practices such as the ones we investigated here, have helped to fight pandemic globally [@besançon2021; @vuong2022].
-However, your findings showed suboptimal adherence to data and code sharing and pre-registeration in COVID-19 related research, whilst majority of researchers had adhered to disclose funding and conflicts of interests related to their COVID-19 related medical research.
+Transparent scientific practices, like the ones we investigated here, have helped to fight pandemic globally [@besançon2021; @vuong2022].
+However, your findings showed suboptimal adherence to data and code sharing and pre-registeration in COVID-19 related research, whilst majority of researchers had adhered to disclose funding and COI related to their COVID-19 related research.
 Stronger commitment to open science practices, particularly to preregistration, data and code sharing, is needed from all stakeholders.

### 2022-04-13 manuscript_covid_transparency.qmd

@@
 That being said, comprehensive picture of transparency practices in COVID-19-related medical literature is still lacking.
-As an example, it is unknown whether open science initiatives related to COVID-19 pandemic \[@besançon2021; @vuong2022], for instance related to data sharing, led to higher number of articles sharing data in COVID-19 research compared to what has been seen in studies in general in biomedical literature [@serghiou2021].
+As an example, it is unknown whether open science initiatives related to COVID-19 pandemic \[@besançon2021; @vuong2022], for instance to advance data sharing, has led to higher number of articles sharing data in COVID-19 research compared to what has been seen in studies in general in biomedical literature [@serghiou2021].
 
@@
 
--   statements of protocol registration.
+-   protocol registration.
 
@@
 
-Basic journal- and article-related information (publisher, publication year and journal name) were retrieved from the EPMC database.
+Basic journal- and article-related information (publisher, publication year, citations to article and journal name) were retrieved from the EPMC database.
 We downloaded additional information about journal impact factors from the Journal Citation Reports (ٰ[jcr.clarivate.com](https://jcr.clarivate.com/)).
@@
 
-Adherence to transparency practices was at similar level in COVID-19 related literature than in other biomedical literature analyzed with the same methods (see Fig 3 in [@serghiou2021\]).
+Adherence to transparency practices was at similar level in COVID-19 related literature than in other biomedical literature analyzed with the same methods (see Fig 3 in \[@serghiou2021\]).
 This is surprising particularly when considering worldwide, remarkable and noble initiatives related to enhancing open science to tackle the pandemic.
-As early as in January 2020, over 100 organizations, including journals, publishers, funders, universities and other institutions, signed a statement that aimed to ensure free access to research data, tools and other information related to COVID-19 \[@vuong2022; @coronavi].
+As early as in January 2020, over 100 organizations, including journals, publishers, funders, universities and other institutions, signed a statement that aimed to ensure free access to research data, tools and other information related to COVID-19 \[@vuong2022; @coronavi\].
 Later also other initiatives to support the goal emerged, for instance COVID-19 Open Research Dataset (CORD-19)[@cord-19], a free resource of over 280,000 articles about the COVID-19 virus.
@@
 In comparison to other studies on the transparency of COVID-19 related research is difficult due to different methodologies and scale of our research.
-However, it seems that data sharing was little less common than in our sample than what was the proportion detected in the beginning of the pandemic measured by the investigation of associated data in PubMed Central [@lucas-dominguez2021\].
+However, it seems that data sharing was little less common than in our sample than what was the proportion detected in the beginning of the pandemic measured by the investigation of associated data in PubMed Central \[@lucas-dominguez2021\].
 In addition, the proportion of studies which shared data in our study sample was lower than in COVID-19 related preprints shared via medRxiv and bioRxiv \[@strcic2022\].
-Lacking preregisteration, sharing code and COI have got attention during the pandemic \[@Glasziou2020; @besançon2021; @luengo-oroz2020], however we are not aware of the investigation of transparency practices in these aspects in COVID-19 related research in this scale in medical research.
+Lacking preregisteration, sharing code and COI have got attention during the pandemic \[@Glasziou2020; @besançon2021; @luengo-oroz2020\], however we are not aware of the investigation of transparency practices in these aspects in COVID-19 related research in this scale in medical research.
 

### 2022-04-13 manuscript_covid_transparency.qmd
added acknowledgements
Adding "what is new" field to be editted

--- 2022-04-13 manuscript_covid_transparency.qmd
@@
 **Authors' contributions:** All authors conceived and designed, and wrote and revised the protocol.
+
+# What is new
+
+-   key findings
+
+-   what this adds to what is known
+
+-   what is the implication/what should change now
+
+Before completing your submission, please take some time to check that your submission includes "What is new" text, including up to 5 points that address the following areas: key findings, what this adds to what is known, and what is the implication/what should change now.
 
@@
 |                    | Citations to article |         |          | Journal Impact Factor |            |          |
-|--------------------|----------------------|---------|----------|:----------------------|:-----------|----------|
+|-----------|-----------|-----------|-----------|:----------|:----------|-----------|
 | Measurement        | Median (IQR)         |         | P-Value  | Median (IQR)          |            | P-value  |

### 2022-04-14 manuscript_covid_transparency.qmd
edits to results and abstract

--- 2022-04-13 manuscript_covid_transparency.qmd
@@
 
+# What is new
+
+-   key findings
+
+-   what this adds to what is known
+
+-   what is the implication/what should change now
+
+Before completing your submission, please take some time to check that your submission includes "What is new" text, including up to 5 points that address the following areas: key findings, what this adds to what is known, and what is the implication/what should change now.
+
 # Abstract
 
-**Objective:** The recent pandemic has shown how crucial it is to have access to quality medical information and data that allows people to have confidence in healthcare decisions.
-We aimed to assess the use of transparency indicators (data availability, code availability, statements of protocol registration and conflicts of interest and funding disclosures) from open access Coronavirus disease 2019 (COVID-19) related articles.
+**Objective:** We aimed to assess the adherence to five transparency practices (data availability, code availability, statements of protocol registration and conflicts of interest and funding disclosures) from open access Coronavirus disease 2019 (COVID-19) related articles.
 
-**Methods:** We searched and exported all open-access COVID-19-related articles from PubMed-indexed journals available in the Europe PMC database published between January 2020 and December 2021.
+**Methods:** We searched and exported all open-access COVID-19-related articles from PubMed-indexed journals available in the Europe PubMed Central database published between January 2020 and December 2021.
 We detected transparency indicators with a validated and automated tool developed to extract the indicators from the downloaded articles.
-Basic journal- and article-related information were retrieved from the PMC database.
+Basic journal- and article-related information were retrieved from the database.
 We used R for the searches (europepmc package), data handling, analysis (rtransparent package) and reporting (ggplot2 package).
 
-**Results:** Of 72,999 available articles published in XX medical journals, 87.0%, (95% CI: 86.8%--87.2%) provided conflicts of interest disclosure and 68.9% (95% CI: 68.7%--69.2%) funding disclosure and 3.6%, (95% CI: 3.5%--3.7%) were preregistered.
+**Results:** The total number of COVID-19-related articles was 175 550, of which we were able to retrieve full texts of 132 618 (75.6%) articles from the database Of those, 87.0%, (95% CI: 86.8%--87.2%) provided conflicts of interest disclosure and 68.9% (95% CI: 68.7%--69.2%) funding disclosure and 3.6%, (95% CI: 3.5%--3.7%) were preregistered.
 8.7% (95% CI: 8.6%--8.9%) of studies shared raw data and 3.1% (95% CI: 3.0%--3.2%) of studies shared code.
@@
 
-**Conclusion:** While the majority of the papers had a COI disclosure, the prevalence of the other transparency practices was far from the acceptable level.
-Even surprisingly low, particularly in terms of data sharing, taking into account global initiatives to improve open science in COVID-19 related research.
+**Conclusion:** While the majority of the articles were freely available and had a COI disclosure, the adherence to other transparency practices was far from the acceptable level.
+The adherence was even surprisingly low, particularly in terms of data sharing, taking into account global initiatives to improve open science in COVID-19 related research.
 A much stronger commitment to open science practices, particularly to preregistration, data and code sharing, is needed from all stakeholders.
@@
 COVID-19 pandemic launched an incredibly fast and voluminous avalanche of scientific publications across the different fields of science.
-There has been discussion and many examples related to poor or suboptimal data, methods and other scientific practices in COVID-19-related studies [@Glasziou2020; @besançon2021; @luengo-oroz2020; @schwab2020\].
+There has been discussion and many examples related to poor or suboptimal data, methods and other scientific practices in COVID-19-related studies \[@Glasziou2020; @besançon2021; @luengo-oroz2020; @schwab2020\].
 However, in addition to fast and successful development of safe and effective vaccination, it is possible that transparent scientific practices, like data and code sharing, have helped to fight pandemic globally \[@besançon2021; @vuong2022\].
 That being said, comprehensive picture of transparency practices in COVID-19-related medical literature is still lacking.
-As an example, it is unknown whether open science initiatives related to COVID-19 pandemic \[@besançon2021; @vuong2022], for instance to advance data sharing, has led to higher number of articles sharing data in COVID-19 research compared to what has been seen in studies in general in biomedical literature [@serghiou2021].
+As an example, it is unknown whether open science initiatives related to COVID-19 pandemic \[@besançon2021; @vuong2022\], for instance to advance data sharing, has led to higher number of articles sharing data in COVID-19 research compared to what has been seen in studies in general in biomedical literature [@serghiou2021].
 
@@
 
-The total number of COVID-19-related papers (open access and non-open access) was 175550, of which full texts of 133470 papers (76.0%) were accessible as open-access via the EPMC.
+The total number of COVID-19-related articles (open access and non-open access) was 175 550, of which 133 470 (76.0%) were accessible as open-access in the EPMC.
 However, 852 (0.6%) were not downloadable because of technical issues and thus were excluded from our analyses.
@@
 These articles were published in 6084 journals of which the top five were International Journal of Environmental Research and Public Health (n=3711), PLoS ONE (n=3059), Scientific Reports (n=1717), Frontiers in Public Health (n=1351), and Frontiers in Psychology (n=1341).
-The mean and the median of the number of citations to these articles were 12.0 (standard deviation (SD) = 92.27) and 2 (interquartile range (IQR) = 7).
+The mean and the median of the number of citations to these articles were 12.0 (standard deviation, SD = 92.27) and 2 (interquartile range, IQR = 7).
 The highest cited article with 15549 citations was an article published in the Lancet in 2020 [@huang2020].
@@
 
-A total of 38383 articles were published in journals with a Journal Impact Factor (JIF).
+A total of 38383 articles were published in journals with JIF.
 Papers that had a COI disclosure, funding statement and were registered tend to be published in journals with lower JIF (P = 0.003, \< 0.001, and = 0.007, respectively).
@@
 
-Over half of the papers had two transparency indicators (n=74085, 55.9%).
-Less than 0.1% (n=45) had all five indicators while 9.0% (n=11890) had no indicators.
-Among papers with two transparency indicators, the majority had COI and funding disclosures (99.3% and 97.9%, respectively).
+One of eleven (9%, n=11890) articles did not adhere to any of the five transparency practices, 24.0% (n=31 844) adhered to one, 55.9% (n=74085) adhered to two and 9.1% (n=12 110) adhered to three.
+2.0% (n=2644) articles showed adherence to four transparency practices, and less than 0.1% of articles (n=45) had used all five practices.
 
 Figure 1 shows the monthly trend for three transparency indicators.
-According to the figure, all indicators showed increasing monthly trend.
-However, the increasing trends for COI and funding disclosures were steeper compared to three other indicators.
+All indicators showed increasing monthly trend.
+However, the increasing trends for COI and funding disclosures were steeper compared to three other practices.
 
@@
 
+# Acknowledgements
+
+The computational analyzes were performed on servers provided by UEF Bioinformatics Center, University of Eastern Finland, Finland.
+The study was supported by a personal grant to ER from the Finnish Dental Society Apollonia.
+
 # References

### 2022-04-14 manuscript_covid_transparency.qmd
edits

@@
 
-**Objective:** We aimed to assess the adherence to five transparency practices (data availability, code availability, statements of protocol registration and conflicts of interest and funding disclosures) from open access Coronavirus disease 2019 (COVID-19) related articles.
+**Objective:** We aimed to assess the adherence to five transparent practices (data availability, code availability, statements of protocol registration and conflicts of interest and funding disclosures) from open access Coronavirus disease 2019 (COVID-19) related articles.
 
 **Methods:** We searched and exported all open-access COVID-19-related articles from PubMed-indexed journals available in the Europe PubMed Central database published between January 2020 and December 2021.
-We detected transparency indicators with a validated and automated tool developed to extract the indicators from the downloaded articles.
+We detected transparent practices with a validated and automated tool.
 Basic journal- and article-related information were retrieved from the database.
@@
 
-**Results:** The total number of COVID-19-related articles was 175 550, of which we were able to retrieve full texts of 132 618 (75.6%) articles from the database Of those, 87.0%, (95% CI: 86.8%--87.2%) provided conflicts of interest disclosure and 68.9% (95% CI: 68.7%--69.2%) funding disclosure and 3.6%, (95% CI: 3.5%--3.7%) were preregistered.
+**Results:** The total number of COVID-19-related articles was 175 550, of which we were able to retrieve full texts of 132 618 (75.6%) articles from the database Of those, 87.0%, (95% CI: 86.8%--87.2%) provided conflicts of interest disclosure and 68.9% (95% CI: 68.7%--69.2%) funding disclosure and 3.6%, (95% CI: 3.5%--3.7%) had registered protocol.
 8.7% (95% CI: 8.6%--8.9%) of studies shared raw data and 3.1% (95% CI: 3.0%--3.2%) of studies shared code.
@@
 
-**Conclusion:** While the majority of the articles were freely available and had a COI disclosure, the adherence to other transparency practices was far from the acceptable level.
+**Conclusion:** While the majority of the articles were freely available and had a COI disclosure, the adherence to other transparent practices was far from the acceptable level.
 The adherence was even surprisingly low, particularly in terms of data sharing, taking into account global initiatives to improve open science in COVID-19 related research.
-A much stronger commitment to open science practices, particularly to preregistration, data and code sharing, is needed from all stakeholders.
+A much stronger commitment to open science practices, particularly to protocol registration, data and code sharing, is needed from all stakeholders.
 
@@
 Our aim was to assess the adherence to transparent scientific practices (data sharing, code sharing, conflict of interest, COI, disclosures, funding disclosures, and protocol registration) from open access full-text COVID-19-related articles published in PubMed indexed journals from the Europe PubMed Central (EPMC) database.
-We mapped adherence to transparency practices across time, journal impact factor (JIF) and citations articles have received
+We mapped adherence to transparent practices across time, journal impact factor (JIF) and citations articles have received
 
@@
 
-We assessed adherence to five transparency practices in the articles:
+We assessed articles' adherence to five transparent practices:
 
--   data availability,
+-   data sharing,
 
--   code availability,
+-   code sharing,
 
@@
 
-To do so, we used a validated and automated tool developed by Serghiou et al. [@serghiou2021] suitable to identify these five transparent practices from articles in XML format from the EPMC database.
+To do so, we used a validated and automated tool developed by Serghiou et al.
+[@serghiou2021] suitable to identify these five transparent practices from articles in XML format from the EPMC database.
+Regarding data sharing, code sharing and protocol registration the tool identifies whether the article had actually adhered to practice (for instance "data available on request" are not counted as sharing data), whereas COI and funding disclosure merely detect whether articles included the disclosure or not (regardless of its content) [@serghiou2021].
 
@@
 The searches and data export from the Europe PMC were conducted with the europepmc package [@europepmc].
-Transparency indicators from the available full-texts were extracted with the rtransparent package [@rtransparent].
-Trends over time in transparency indicators by journal- and article-related information were reported using descriptive tabulations and graphical illustrations, for instance using the ggplot2 package [@ggplot2].
-We used the numbers for sensitivity and specificity of the rtransparent package (3) to generate 95% confidence intervals (CIs) for our prevalence estimates of the transparency indicators with the epiR package [@epiR].
-We also provided monthly trends for transparency indicators, to check whether there were any trends in transparency practices over time.
-We used Chi-square for comparing indicators between years and the Wilcoxon rank sum test to test whether there was a relationship of transparency indicators with citations to articles and JIFs.
+Indicators of transparent practices from the available full-texts were extracted with the rtransparent package [@rtransparent].
+Trends over time in transparent practices by journal- and article-related information were reported using descriptive tabulations and graphical illustrations, for instance using the ggplot2 package [@ggplot2].
+We used the numbers for sensitivity and specificity of the rtransparent package (3) to generate 95% confidence intervals (CIs) for our prevalence estimates of the transparency practices with the epiR package [@epiR].
+We also provided monthly trends for transparency practices, to check whether there were any trends in transparent practices over time.
+We used Chi-square for comparing adherence to practices between years and the Wilcoxon rank sum test to test whether there was a relationship of transparency practices with citations to articles and JIFs.
 The thorough system information for the analysis is available in Appendix 2.
@@
 
-Table 1. Citations and journal impact factor by transparency indicators.
+Table 1. Citations and journal impact factor by transparent practices.
 
@@
 One of eleven (9%, n=11890) articles did not adhere to any of the five transparency practices, 24.0% (n=31 844) adhered to one, 55.9% (n=74085) adhered to two and 9.1% (n=12 110) adhered to three.
-2.0% (n=2644) articles showed adherence to four transparency practices, and less than 0.1% of articles (n=45) had used all five practices.
+2.0% (n=2644) articles showed adherence to four transparent practices, and less than 0.1% of articles (n=45) had used all five practices.
 
-Figure 1 shows the monthly trend for three transparency indicators.
-All indicators showed increasing monthly trend.
+Figure 1 shows the monthly trend for three transparent practices.
+All practices showed increasing monthly trend.
 However, the increasing trends for COI and funding disclosures were steeper compared to three other practices.
@@
 
-Adherence to transparency practices increased over time in COVID-19 related medical literature during 2020-2021.
+Adherence to transparent practices increased over time in COVID-19 related medical literature during 2020-2021.
 Adherence to reporting COI disclosure was high throughout study period, in addition most of articles had funding disclosure.
 Data sharing, code sharing and protocol registration were rare, but improved little over study period.
-COI disclosures seemed to be more common in articles published in lower impact journals, otherwise differences in transparency practices according to received citations and JIF were small.
+COI disclosures seemed to be more common in articles published in lower impact journals, otherwise differences in transparent practices according to received citations and JIF were small.
 
-Adherence to transparency practices was at similar level in COVID-19 related literature than in other biomedical literature analyzed with the same methods (see Fig 3 in \[@serghiou2021\]).
+Adherence to transparent practices was at similar level in COVID-19 related literature than in other biomedical literature analyzed with the same methods (see Fig 3 in [@serghiou2021\]).
 This is surprising particularly when considering worldwide, remarkable and noble initiatives related to enhancing open science to tackle the pandemic.
-As early as in January 2020, over 100 organizations, including journals, publishers, funders, universities and other institutions, signed a statement that aimed to ensure free access to research data, tools and other information related to COVID-19 \[@vuong2022; @coronavi\].
+As early as in January 2020, over 100 organizations, including journals, publishers, funders, universities and other institutions, signed a statement that aimed to ensure free access to research data, tools and other information related to COVID-19 \[@vuong2022; @coronavi].
 Later also other initiatives to support the goal emerged, for instance COVID-19 Open Research Dataset (CORD-19)[@cord-19], a free resource of over 280,000 articles about the COVID-19 virus.
@@
 In comparison to other studies on the transparency of COVID-19 related research is difficult due to different methodologies and scale of our research.
-However, it seems that data sharing was little less common than in our sample than what was the proportion detected in the beginning of the pandemic measured by the investigation of associated data in PubMed Central \[@lucas-dominguez2021\].
+However, it seems that data sharing was little less common than in our sample than what was the proportion detected in the beginning of the pandemic measured by the investigation of associated data in PubMed Central [@lucas-dominguez2021\].
 In addition, the proportion of studies which shared data in our study sample was lower than in COVID-19 related preprints shared via medRxiv and bioRxiv \[@strcic2022\].
-Lacking preregisteration, sharing code and COI have got attention during the pandemic \[@Glasziou2020; @besançon2021; @luengo-oroz2020\], however we are not aware of the investigation of transparency practices in these aspects in COVID-19 related research in this scale in medical research.
+Lacking protocol registration, sharing code and COI have got attention during the pandemic \[@Glasziou2020; @besançon2021; @luengo-oroz2020], however we are not aware of the investigation of transparency in these aspects in COVID-19 related research in this scale in medical research.
 
-Despite we used representative data and validated methods to investigate transparency practices in COVID-19-related research, our study has some weaknesses.
+Despite we used representative data and validated methods to investigate transparency in COVID-19-related research, our study has some weaknesses.
 The study sample was restricted to open access articles in the EPMC database which may not correspond to all COVID-19-related studies published in medical journals.
@@
 Transparent scientific practices, like the ones we investigated here, have helped to fight pandemic globally [@besançon2021; @vuong2022].
-However, your findings showed suboptimal adherence to data and code sharing and pre-registeration in COVID-19 related research, whilst majority of researchers had adhered to disclose funding and COI related to their COVID-19 related research.
-Stronger commitment to open science practices, particularly to preregistration, data and code sharing, is needed from all stakeholders.
+Whilst majority of COVID-19-related articles were open access and had adhered to disclose funding and COI, your findings showed suboptimal adherence to data and code sharing and protocol registration.
+Stronger commitment to open science practices, particularly to protocol registration, data and code sharing, is needed from all stakeholders.
 

### 2022-04-15 manuscript_covid_transparency.qmd
minor edit to methods

@@
 [@serghiou2021] suitable to identify these five transparent practices from articles in XML format from the EPMC database.
-Regarding data sharing, code sharing and protocol registration the tool identifies whether the article had actually adhered to practice (for instance "data available on request" are not counted as sharing data), whereas COI and funding disclosure merely detect whether articles included the disclosure or not (regardless of its content) [@serghiou2021].
+Regarding data sharing, code sharing and protocol registration the tool identifies whether the article had actually adhered to practice (for instance "data available on request" are not counted as sharing data), whereas COI and funding disclosure merely detect whether articles include the disclosure or not (regardless of its specific content) [@serghiou2021].
 

### 2022-04-21 rmarkdown-covid-19-transparent-analysis.Rmd
edits to results and abstract
delete setwd()

--- 2022-03-09 rmarkdown-covid-19-transparent-analysis.Rmd
@@
 # Loading the datasets
-First, set working directory to the folder that has all the datasets in it using setwd() function and then run the following code:
+
 ```{r}

### 2022-04-21 rmarkdown-covid-19-transparent-analysis.Rmd

--- 2022-04-01 rmarkdown-covid-19-transparent-analysis.Rmd
@@
                knitr,
-               here)
+               here,
+               epiR,
+               ggpubr,
+               tibble)
 ```

### 2022-04-23 manuscript_covid_transparency.qmd
changes in the intro and new references

@@
 
-COVID-19 pandemic launched an incredibly fast and voluminous avalanche of scientific publications across the different fields of science.
-There has been discussion and many examples related to poor or suboptimal data, methods and other scientific practices in COVID-19-related studies \[@Glasziou2020; @besançon2021; @luengo-oroz2020; @schwab2020\].
-However, in addition to fast and successful development of safe and effective vaccination, it is possible that transparent scientific practices, like data and code sharing, have helped to fight pandemic globally \[@besançon2021; @vuong2022\].
-That being said, comprehensive picture of transparency practices in COVID-19-related medical literature is still lacking.
-As an example, it is unknown whether open science initiatives related to COVID-19 pandemic \[@besançon2021; @vuong2022\], for instance to advance data sharing, has led to higher number of articles sharing data in COVID-19 research compared to what has been seen in studies in general in biomedical literature [@serghiou2021].
+During the recent COVID-19 pandemic, part of the public in developed countries justified their lack of adherence to public health measures due to the lack of transparency of the scientific studies used as an argument by national authorities [@Skipper2020-rs\].
+Thus, supranational bodies such as the OECD indicate that transparent communication to citizens is vital to support public health measures and to counter misinformation \[@Bellantoni_undated-tf\].
+On the other hand, the COVID-19 pandemic launched an incredibly fast and voluminous avalanche of scientific publications across the different fields of science.
+This research led to the development of vaccines as well as the evaluation of community interventions.
+On the one hand, it is possible that transparent scientific practices, like data and code sharing, have helped to fight pandemic globally \[@besançon2021; @vuong2022\], while on the other hand, the lack of transparency of some key developments has led the public to mistrust the results of research and public health measures \[@Skipper2020-rs\].
+This shows that there is a lack of awareness about the level of transparency practices in COVID-19-related medical literature.
+As an example, it is unknown whether open science initiatives related to COVID-19 pandemic \[@besançon2021; @vuong2022] to advance data sharing, has led to higher number of articles sharing data in COVID-19 research compared to what has been seen in studies in general in biomedical literature [@serghiou2021].
 
@@
 
-Adherence to transparent practices was at similar level in COVID-19 related literature than in other biomedical literature analyzed with the same methods (see Fig 3 in [@serghiou2021\]).
+Adherence to transparent practices was at similar level in COVID-19 related literature than in other biomedical literature analyzed with the same methods (see Fig 3 in \[@serghiou2021\]).
 This is surprising particularly when considering worldwide, remarkable and noble initiatives related to enhancing open science to tackle the pandemic.
-As early as in January 2020, over 100 organizations, including journals, publishers, funders, universities and other institutions, signed a statement that aimed to ensure free access to research data, tools and other information related to COVID-19 \[@vuong2022; @coronavi].
+As early as in January 2020, over 100 organizations, including journals, publishers, funders, universities and other institutions, signed a statement that aimed to ensure free access to research data, tools and other information related to COVID-19 \[@vuong2022; @coronavi\].
 Later also other initiatives to support the goal emerged, for instance COVID-19 Open Research Dataset (CORD-19)[@cord-19], a free resource of over 280,000 articles about the COVID-19 virus.
@@
 In comparison to other studies on the transparency of COVID-19 related research is difficult due to different methodologies and scale of our research.
-However, it seems that data sharing was little less common than in our sample than what was the proportion detected in the beginning of the pandemic measured by the investigation of associated data in PubMed Central [@lucas-dominguez2021\].
+However, it seems that data sharing was little less common than in our sample than what was the proportion detected in the beginning of the pandemic measured by the investigation of associated data in PubMed Central \[@lucas-dominguez2021\].
 In addition, the proportion of studies which shared data in our study sample was lower than in COVID-19 related preprints shared via medRxiv and bioRxiv \[@strcic2022\].
-Lacking protocol registration, sharing code and COI have got attention during the pandemic \[@Glasziou2020; @besançon2021; @luengo-oroz2020], however we are not aware of the investigation of transparency in these aspects in COVID-19 related research in this scale in medical research.
+Lacking protocol registration, sharing code and COI have got attention during the pandemic \[@Glasziou2020; @besançon2021; @luengo-oroz2020\], however we are not aware of the investigation of transparency in these aspects in COVID-19 related research in this scale in medical research.
 

### 2022-04-27 manuscript/manuscript_covid_transparency.qmd
First draft

--- 2022-04-23 manuscript_covid_transparency.qmd
@@
 
-format: html
+format:
+  docx:
+    section-numbers: true
+    highlight-style: github 
+#format: html
 #format:
@@
 editor: visual
+from: markdown+tex_math_single_backslash
 bibliography: references.bib
@@
 
-**Results:** The total number of COVID-19-related articles was 175 550, of which we were able to retrieve full texts of 132 618 (75.6%) articles from the database Of those, 87.0%, (95% CI: 86.8%--87.2%) provided conflicts of interest disclosure and 68.9% (95% CI: 68.7%--69.2%) funding disclosure and 3.6%, (95% CI: 3.5%--3.7%) had registered protocol.
+**Results:** The total number of COVID-19-related articles was 175,550, of which we were able to retrieve full texts of 132,598 (76%) articles from the database Of those, 87.0%, (95% CI: 86.8%--87.2%) provided conflicts of interest disclosure and 68.9% (95% CI: 68.7%--69.2%) funding disclosure and 3.6%, (95% CI: 3.5%--3.7%) had registered protocol.
 8.7% (95% CI: 8.6%--8.9%) of studies shared raw data and 3.1% (95% CI: 3.0%--3.2%) of studies shared code.
-Transparent practices were more common in articles published in 2021 than in 2020.
+Transparent practices were more common in articles published in 2022 than in 2021 and 2020.
+Apart from COI disclosure, articles that adhered to transparency practices were published in journals with slightly higher median Journal Impact Factor but received equal citations.
 
@@
 
-During the recent COVID-19 pandemic, part of the public in developed countries justified their lack of adherence to public health measures due to the lack of transparency of the scientific studies used as an argument by national authorities [@Skipper2020-rs\].
-Thus, supranational bodies such as the OECD indicate that transparent communication to citizens is vital to support public health measures and to counter misinformation \[@Bellantoni_undated-tf\].
+During the recent COVID-19 pandemic, part of the public in developed countries justified their lack of adherence to public health measures due to the lack of transparency of the scientific studies used as an argument by national authorities [@Skipper2020-rs].
+Thus, supranational bodies such as the OECD indicate that transparent communication to citizens is vital to support public health measures and to counter misinformation [@Bellantoni_undated-tf].
 On the other hand, the COVID-19 pandemic launched an incredibly fast and voluminous avalanche of scientific publications across the different fields of science.
 This research led to the development of vaccines as well as the evaluation of community interventions.
-On the one hand, it is possible that transparent scientific practices, like data and code sharing, have helped to fight pandemic globally \[@besançon2021; @vuong2022\], while on the other hand, the lack of transparency of some key developments has led the public to mistrust the results of research and public health measures \[@Skipper2020-rs\].
+On the one hand, it is possible that transparent scientific practices, like data and code sharing, have helped to fight pandemic globally [@besançon2021; @vuong2022], while on the other hand, the lack of transparency of some key developments has led the public to mistrust the results of research and public health measures [@Skipper2020-rs].
 This shows that there is a lack of awareness about the level of transparency practices in COVID-19-related medical literature.
-As an example, it is unknown whether open science initiatives related to COVID-19 pandemic \[@besançon2021; @vuong2022] to advance data sharing, has led to higher number of articles sharing data in COVID-19 research compared to what has been seen in studies in general in biomedical literature [@serghiou2021].
+As an example, it is unknown whether open science initiatives related to COVID-19 pandemic [@besançon2021; @vuong2022] to advance data sharing, has led to higher number of articles sharing data in COVID-19 research compared to what has been seen in studies in general in biomedical literature [@serghiou2021].
 
@@
 This enabled us to bypass screening for COVID-19-related papers and decrease the human labor for our research.
-We confirmed our approach by choosing a random sample of 100 papers and of those, all were relevant to COVID-19.
+We confirmed our approach by choosing a random sample of 500 papers and of those, all but 10 were relevant to COVID-19.
 
@@
 
-(COVID-19 query for title, keywords, and results) AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:\[2020-01-01 TO 2021-12-31\]) AND (OPEN_ACCESS:y) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")'
+(COVID-19 query for title, keywords, and results) AND (SRC:"MED") AND (LANG:"eng" OR LANG:"en" OR LANG:"us") AND (FIRST_PDATE:[2020-01-01 TO 2021-12-31]) AND (OPEN_ACCESS:y) AND (PUB_TYPE:"Journal Article" OR PUB_TYPE:"research-article" OR PUB_TYPE:"rapid-communication" OR PUB_TYPE:"product-review")'
 
@@
 
-We used R v4.1.2 [@R-base] for searches, data handling, analysis and reporting.
+We used R v4.1.2 [@base] for searches, data handling, analysis and reporting.
 The searches and data export from the Europe PMC were conducted with the europepmc package [@europepmc].
-Indicators of transparent practices from the available full-texts were extracted with the rtransparent package [@rtransparent].
-Trends over time in transparent practices by journal- and article-related information were reported using descriptive tabulations and graphical illustrations, for instance using the ggplot2 package [@ggplot2].
-We used the numbers for sensitivity and specificity of the rtransparent package (3) to generate 95% confidence intervals (CIs) for our prevalence estimates of the transparency practices with the epiR package [@epiR].
-We also provided monthly trends for transparency practices, to check whether there were any trends in transparent practices over time.
-We used Chi-square for comparing adherence to practices between years and the Wilcoxon rank sum test to test whether there was a relationship of transparency practices with citations to articles and JIFs.
-The thorough system information for the analysis is available in Appendix 2.
+Indicators of transparency practices from the available full-texts were extracted with the rtransparent package [@rtransparent].
+Trends over time in transparency practices were reported using descriptive tabulations and graphical illustrations using the ggplot2 package [@ggplot2].
+We used the sensitivity and specificity of the rtransparent package [@serghiou2021] to generate 95% confidence intervals (CIs) for our prevalence estimates of the transparency practices with the epiR package [@epiR].
+
+Additionally, we calculated how many transparency practices articles had adhered to (from 0 to 5 practices).
+We used Chi-square for comparing indicators between years and the Wilcoxon rank sum test to test whether there was a relationship between transparency indicators and JIF or received citations.
+Fisher exact tests with Monte-Carlo simulated p-values for differences in transparency practices by journals and publishers were performed.
+The thorough information, data and code for the analysis are available as supplementary material on the OSF repository ([osf.io/x3kb6](https://osf.io/x3kb6/)).
 
@@
 
-The total number of COVID-19-related articles (open access and non-open access) was 175 550, of which 133 470 (76.0%) were accessible as open-access in the EPMC.
-However, 852 (0.6%) were not downloadable because of technical issues and thus were excluded from our analyses.
-Of those 132618 remaining articles, 47550 were published in 2020, 75857 in 2021, and 9211 in the first two months of 2022 (Appendix 3).
+The total number of COVID-19-related articles (open access and non-open access) was 175,544, of which 133,467 (76.0%) were accessible as open-access in the EPMC.
+However, 869 (0.7%) were not downloadable because of technical issues and thus were excluded from our analyses.
+Of those 132,598 remaining articles, 47,548 were published in 2020, 75,842 in 2021, and 9,208 in the first two months of 2022.
 
-These articles were published in 6084 journals of which the top five were International Journal of Environmental Research and Public Health (n=3711), PLoS ONE (n=3059), Scientific Reports (n=1717), Frontiers in Public Health (n=1351), and Frontiers in Psychology (n=1341).
-The mean and the median of the number of citations to these articles were 12.0 (standard deviation, SD = 92.27) and 2 (interquartile range, IQR = 7).
-The highest cited article with 15549 citations was an article published in the Lancet in 2020 [@huang2020].
+These articles were published in 6,020 journals of which the top five were International Journal of Environmental Research and Public Health (n=3,710), PLoS ONE (n=3,059), Scientific Reports (n=1,716), Frontiers in Public Health (n=1,351), and Frontiers in Psychology (n=1,341).
+The mean and the median of the number of citations to these articles were 12.0 (standard deviation, SD = 92.28) and 2 (interquartile range, IQR = 7).
+The highest cited article with 15,549 citations was an article published in the Lancet in 2020 [@huang2020].
 
-About nine-tenth of the articles (n=115405) had a statement to disclose COI (87.0%, 95% CI: 86.8%--87.2%).
-Articles published in 2022 had a higher rate of COI disclosure (91.4% vs 89.1% for 2021, and 82.9% for 2020, P \< 0.001).
-More than two-thirds of the references had a funding statement (n=91425) (68.9%, 95% CI: 68.7%--69.2%).
-Articles in 2022 had a funding statement more frequently than two other years (79.0% vs 73.1% for 2021, and 60.4% for 2020, P \< 0.001).
-Less than 1 in 25 of the articles (n=4708) were registered beforehand (3.6%, 95% CI: 3.5%--3.7%) and the rate was higher in 2022 (5.4% vs 4.2% for 2021, and 2.2% for 2020, P \< 0.001).
+About nine-tenth of the articles (n=115,387) had a statement to disclose COI (87.0%, 95% CI: 86.8%--87.2%).
+Articles published in 2022 had a higher rate of COI disclosure (91.4% (2022), 89.1% (2021), 82.9% (2020), P\<0.001).
+More than two-thirds of the references had a funding statement (n=91,413) (68.9%, 95% CI: 68.7%--69.2%).
+Articles in 2022 had a funding statement more frequently than two other years (79.0% (2022), 73.1% (2021), 60.4% (2020), P\<0.001).
+Less than 1 in 25 of the articles (n=4708) were registered beforehand (3.6%, 95% CI: 3.5%--3.7%) and the rate was higher in 2022 (5.4% (2022), 4.2% (2021), 2.2% (2020), P\<0.001).
 
-Regarding data and code availability, less than 1 in 11 articles had shared their data (n=11552) (8.7%, 95% CI: 8.6%--8.9%).
-Articles in 2022 had shared data more frequently (11.9% vs 9.6% for 2021, and 6.7% for 2020, P \< 0.001).
-Less than 1 in 30 of the articles (n=4055) had shared their code (3.1%, 95% CI: 3.0%--3.2%) and the rate was higher in 2022 (4.0% vs 3.5% for 2021, and 2.2 for 2020, P \< 0.001).
+Regarding data and code availability, less than 1 in 11 articles had shared their data (n=11551) (8.7%, 95% CI: 8.6%--8.9%).
+Articles in 2022 had shared data more frequently (11.9% (2022), 9.6% (2021), 6.7% (2020), P\<0.001).
+Less than 1 in 30 of the articles (n=4055) had shared their code (3.1%, 95% CI: 3.0%--3.2%) and the rate was higher in 2022 (4.0% (2022), 3.5% (2021), 2.2 (2020), P\<0.001).
 
-A total of 38383 articles were published in journals with JIF.
-Papers that had a COI disclosure, funding statement and were registered tend to be published in journals with lower JIF (P = 0.003, \< 0.001, and = 0.007, respectively).
-Detailed information is illustrated in Table 1.
+There was a consistently increasing monthly trend in the proportion of articles with COI disclosure, funding disclosure and protocol registration, and the proportions were the highest in February 2022 (Figure 1B).
+Increases in data and code sharing were weaker than in other transparency practices over time.
 
-Table 1. Citations and journal impact factor by transparent practices.
+One-eleventh of the articles (9.0%, n=11,889) did not adhere to any of the five transparency practices, 24.0% (n=31,836) adhered to one, 56.0% (n=74,075) adhered to two, and 9.1% (n=12,109) adhered to three.
+Few (2.0%, n=2,644) articles showed adherence to four transparency practices, and less than 0.1% of the articles (n=45) had used all five practices.
 
-|                    | Citations to article |         |          | Journal Impact Factor |            |          |
-|-----------|-----------|-----------|-----------|:----------|:----------|-----------|
-| Measurement        | Median (IQR)         |         | P-Value  | Median (IQR)          |            | P-value  |
-|                    | With                 | Without |          | With                  | Without    |          |
-| COI disclosure     | 2 (6)                | 2 (7)   | \< 0.001 | 4.1 (2.31)            | 7.2 (5.12) | \< 0.001 |
-| Funding disclosure | 2 (6)                | 2 (8)   | \< 0.001 | 4.1 (2.32)            | 4.1 (2.94) | \< 0.001 |
-| Registration       | 1 (5)                | 2 (6)   | \< 0.001 | 3.8 (2.38)            | 4.2 (2.44) | \< 0.001 |
-| Data sharing       | 2 (8)                | 2 (6)   | \< 0.001 | 4.0 (2.98)            | 4.1 (2.36) | \< 0.001 |
-| Code sharing       | 2 (8)                | 2 (6)   | 0.70     | 4.4 (4.41)            | 4.1 (2.32) | 0.08     |
+There were journal and publisher-related statistically significant differences in all five practices (p\<0.001 for all), except for code sharing based on JIF (0.14) (Table 1 and 2, Supplementary material).
 
-: P-value based on the Wilcoxon rank sum test.\
-COI: Conflict of interest; SD: Standard deviation; IQR: Inter-quartile range.
-
-One of eleven (9%, n=11890) articles did not adhere to any of the five transparency practices, 24.0% (n=31 844) adhered to one, 55.9% (n=74085) adhered to two and 9.1% (n=12 110) adhered to three.
-2.0% (n=2644) articles showed adherence to four transparent practices, and less than 0.1% of articles (n=45) had used all five practices.
-
-Figure 1 shows the monthly trend for three transparent practices.
-All practices showed increasing monthly trend.
-However, the increasing trends for COI and funding disclosures were steeper compared to three other practices.
+Apart from COI disclosure, articles that adhered to transparency practices were published in journals with slightly higher median JIF but received equal citations (Table 3).
 
@@
 
-Adherence to transparent practices was at similar level in COVID-19 related literature than in other biomedical literature analyzed with the same methods (see Fig 3 in \[@serghiou2021\]).
+Adherence to transparent practices was at similar level in COVID-19 related literature than in other biomedical literature analyzed with the same methods (see Fig 3 in [@serghiou2021]).
 This is surprising particularly when considering worldwide, remarkable and noble initiatives related to enhancing open science to tackle the pandemic.
-As early as in January 2020, over 100 organizations, including journals, publishers, funders, universities and other institutions, signed a statement that aimed to ensure free access to research data, tools and other information related to COVID-19 \[@vuong2022; @coronavi\].
+As early as in January 2020, over 100 organizations, including journals, publishers, funders, universities and other institutions, signed a statement that aimed to ensure free access to research data, tools and other information related to COVID-19 [@vuong2022; @coronavi].
 Later also other initiatives to support the goal emerged, for instance COVID-19 Open Research Dataset (CORD-19)[@cord-19], a free resource of over 280,000 articles about the COVID-19 virus.
@@
 In comparison to other studies on the transparency of COVID-19 related research is difficult due to different methodologies and scale of our research.
-However, it seems that data sharing was little less common than in our sample than what was the proportion detected in the beginning of the pandemic measured by the investigation of associated data in PubMed Central \[@lucas-dominguez2021\].
-In addition, the proportion of studies which shared data in our study sample was lower than in COVID-19 related preprints shared via medRxiv and bioRxiv \[@strcic2022\].
-Lacking protocol registration, sharing code and COI have got attention during the pandemic \[@Glasziou2020; @besançon2021; @luengo-oroz2020\], however we are not aware of the investigation of transparency in these aspects in COVID-19 related research in this scale in medical research.
+However, it seems that data sharing was little less common than in our sample than what was the proportion detected in the beginning of the pandemic measured by the investigation of associated data in PubMed Central [@lucas-dominguez2021].
+In addition, the proportion of studies which shared data in our study sample was lower than in COVID-19 related preprints shared via medRxiv and bioRxiv [@strcic2022].
+Lacking protocol registration, sharing code and COI have got attention during the pandemic [@Glasziou2020; @besançon2021; @luengo-oroz2020], however we are not aware of the investigation of transparency in these aspects in COVID-19 related research in this scale in medical research.
 
@@
 # References
+
+# Tables
+
+Table 1. Transparency practices in the six most common journals in the sample.
+
+| **Measurement**       | **BMC Oral Health, N = 2503** | **Dental Press J Orthod, N = 530** | **Head Face Med, N = 572** | **J Appl Oral Sci, N = 1440** | **J Indian Prosthodont Soc, N = 495** | **Med Oral Patol Oral Cir Bucal, N = 1174** | **Other, N = 3945** | **p-value** |
+|--------|--------|--------|--------|--------|--------|--------|--------|--------|
+| COI disclosure        | 2489 (99%)                    | 502 (95%)                          | 537 (94%)                  | 303 (21%)                     | 301 (61%)                             | 869 (74%)                                   | 3172 (80%)          | \<0.001     |
+| Funding disclosure    | 2232 (89%)                    | 87 (16%)                           | 280 (49%)                  | 656 (46%)                     | 301 (61%)                             | 497 (42%)                                   | 2505 (63%)          | \<0.001     |
+| Protocol registration | 326 (13%)                     | 12 (2.3%)                          | 23 (4.0%)                  | 31 (2.2%)                     | 4 (0.8%)                              | 43 (3.7%)                                   | 298 (7.6%)          | \<0.001     |
+| Data sharing          | 120 (4.8%)                    | 2 (0.4%)                           | 2 (0.3%)                   | 7 (0.5%)                      | 0 (0%)                                | 5 (0.4%)                                    | 76 (1.9%)           | \<0.001     |
+| Code sharing          | 6 (0.2%)                      | 0 (0%)                             | 0 (0%)                     | 0 (0%)                        | 0 (0%)                                | 0 (0%)                                      | 0 (0%)              | 0.61        |
+
+P-values from Fisher's Exact Test for Count Data with simulated p-value (based on 2000 replicates). COI: Conflict of interest.
+
+Table 2. Transparency practices in the six most common publishers of journals in the sample.
+
+| Measurement           | BioMed Central, N = 3075 | Dental Press Editora, N = 530 | Faculdade de Odontologia de Bauru, N = 1440 | Medicina Oral, Patologia Oral y Cirugia Bucal, N 1174 | Springer Verlag, N = 935 | Wolters Kluwer Medknow Publications, N = 495 | Other, N = 2619 | p-value |
+|--------|--------|--------|--------|--------|--------|--------|--------|--------|
+| COI disclosure        | 3026 (98%)               | 502 (95%)                     | 303 (21%)                                   | 869 (74%)                                             | 904 (97%)                | 301 (61%)                                    | 1878 (72%)      | \<0.001 |
+| Funding disclosure    | 2512 (82%)               | 87 (16%)                      | 656 (46%)                                   | 497 (42%)                                             | 596 (64%)                | 301 (61%)                                    | 1593 (61%)      | \<0.001 |
+| Protocol registration | 349 (11%)                | 12 (2.3%)                     | 31 (2.2%)                                   | 43 (3.7%)                                             | 113 (12%)                | 4 (0.8%)                                     | 144 (5.5%)      | \<0.001 |
+| Data sharing          | 122 (4.0%)               | 2 (0.4%)                      | 7 (0.5%)                                    | 5 (0.4%)                                              | 13 (1.4%)                | 0 (0%)                                       | 54 (2.1%)       | \<0.001 |
+| Code sharing          | 6 (0.2%)                 | 0 (0%)                        | 0 (0%)                                      | 0 (0%)                                                | 1 (0.1%)                 | 0 (0%)                                       | 1 (\<0.1%)      | 0.68    |
+
+P-values from Fisher's Exact Test for Count Data with simulated p-value (based on 2000 replicates).
+391 articles were published in journal without available publisher information.
+COI: Conflict of interest.
+
+Table 3. Citations to article and journal impact factor by transparency practices.
+
+|                    | Citations to article |         |          | Journal Impact Factor |            |          |
+|-----------|-----------|-----------|-----------|:----------|:----------|-----------|
+| Measurement        | Median (IQR)         |         | P-Value  | Median (IQR)          |            | P-value  |
+|                    | With                 | Without |          | With                  | Without    |          |
+| COI disclosure     | 2 (5)                | 5 (9)   | \< 0.001 | 2.8 (0.61)            | 2.7 (0.06) | \< 0.001 |
+| Funding disclosure | 2 (5)                | 3 (7)   | \< 0.001 | 2.8 (0.06)            | 2.7 (0.70) | \< 0.001 |
+| Registration       | 1 (4)                | 2 (7)   | \< 0.001 | 2.8 (0.01)            | 2.7 (0.61) | \< 0.001 |
+| Data sharing       | 2 (4)                | 2 (7)   | 0.01     | 2.8 (0.00)            | 2.8 (0.49) | \< 0.001 |
+| Code sharing       | 1.5 (3)              | 2 (7)   | 0.70     | 2.8 (0.00)            | 2.8 (0.49) | 0.08     |
+
+P-value based on the Wilcoxon rank sum test. 1929 articles were published in journals with no impact factor. COI: Conflict of interest; SD: Standard deviation; IQR: Inter-quartile range.

### 2022-04-27 protocol/protocol_covid_transparency.Rmd
First draft

--- 2022-02-26 manuscript_covid_transparency.Rmd
@@
 ---
-title             : "Transparency of COVID-19-related research in medical journals"
-shorttitle        : "Transparency of COVID-19-related research"
-
+title: "Transparency of COVID-19 medical literature"
 author: 
-    - name        : "Eero Raittio"
-    affiliation   : "1"
-    corresponding : no    # Define only one corresponding author
-    email         : "eero.raittio@uef.fi"
-  - name          : "Ahmad Sofi-Mahmudi"
-    affiliation   : c("2", "3")
-    corresponding : yes    # Define only one corresponding author
-    address       : "41, Shanaz St., Seqiz Health Network, Seqiz, Kurdistan, Iran."
-    email         : "a.sofimahmudi@gmail.com"
-
-affiliation:
-  - id            : "1"
-    institution   : "Institute of Dentistry, University of Eastern Finland, Kuopio, Finland."
-  - id            : "2"
-    institution   : "Seqiz Health Network, Kurdistan University of Medical Sciences, Seqiz, Iran."
-  - id            : "3"
-    institution   : "Cochrane Iran Associate Centre, National Institute for Medical Research Development (NIMAD), Tehran, Iran."
-
-note: "\\clearpage"
-
-authornote:
-
-abstract:
-
-keywords          : "Hypothesis Testing"
-wordcount         : "2,860"
-
-bibliography      : ["references.bib", "r-references.bib"]
-
-floatsintext      : yes #place figures and tables in the text rather than at the end
-numbersections    : no #number sections headings
-figurelist        : no #create list of figure captions
-tablelist         : no #create list of table captions
-footnotelist      : no ##create list of footnotes
-linenumbers       : no #add line numbers in the margins
-mask              : no #omit identifying information from the title page
-draft             : no #add "draft" watermark across all pages
-
-documentclass     : "apa6"
-classoption       : "man"
-output            : papaja::apa6_word
+  - John Doe, Institution One
+  - Jane Doe, Institution Two
+date: "`r Sys.Date()`"
+output: html_document
 ---
-
-```{r setup, include = FALSE}
-library(ggplot2)
-library(papaja)
-library(citr)
-library(rpact)
-library(dplyr)
-r_refs("r-references.bib")
-```
-
-"Whatever else is done about null-hypothesis tests, let us stop viewing statistical analysis as a sanctification process.
-We are awash in a sea of uncertainty, caused by a flood tide of sampling and measurement errors, and there are no objective procedures that avoid human judgment and guarantee correct interpretations of results." [@abelson1997: p.13]
-
-Psychological science is experiencing a period of widespread methodological reflection [@nelson2018].
-Starting 2010, a chain of events caused psychological scientists to doubt the credibility of psychological research [@moshontz2018].
-These events included a case of widespread academic fraud, a series of failed replications of famous findings, and a paper published in psychology's most prestigious journal claiming that extrasensory perception (ESP) is real [@nelson2018].
-
-According to some, one of the root causes of the crisis of confidence is the misunderstanding and misuse of statistical methods [@goodman2019].
-In psychological science, one of the most frequently used approaches to statistical inference is the null-hypothesis significance test (NHST).
-For decades, researchers have been warned about mindless applications of this tool [e.g., @cohen1994; @gigerenzer2004; @nickerson2000; @rozeboom1960].
-
-It has been argued that, in practice, the NHST has taken on a mechanical nature.
-According to Gigerenzer et al.
-[-@gigerenzer1989, p.210-227], scientists have interpreted the NHST as a completely mechanized procedure to arrive at new knowledge while eliminating the need for personal judgment and promising objectivity.
-A normative view has emerged; statistics can speak with one voice and replace personal judgment by means of formal rules.
-
-Among statisticians, there seems to be widespread agreement about what not do with hypothesis tests and *p*-values.
-Unfortunately, there is less agreement about what researchers should be doing instead [@wasserstein2019].
-Crucially, although the founders of statistics disagreed on many things, they seemed to agree on at least one thing: they all rejected mechanical inference and highlighted the importance of human judgment [@abelson1995; @gigerenzer2015; @mayo2018].
-
-Unfortunately, the literature on NHST seems to have lost sight of the important distinction between the tool itself and its thoughtless application [@mayo2018; @lakens2020a].
-It seems the tool has become synonymous with its weakest construal: setting up a nil-null hypothesis and using a single *p*-value below a conventional threshold to infer evidence for a substantive research claim [@mayo2018; @gelman2016].
-As a consequence, numerous scholars have called for banning the tool altogether.
-
-It is important to emphasize, however, that NHST has a place in scientific inference [@abelson1997; @lakens2020a] and that it is likely to stay [@goodman2019].
-Rather than invalidating the tool on the basis of its weakest construal, an important step forward is to highlight how NHSTs can be thoughtfully applied in practice [@wasserstein2019].
-
-In this paper, I examine the way in which hypothesis tests are conducted in psychological science.
-I focus on three decisions central to hypothesis testing: (a) the null hypothesis to be tested, (b) the choice of acceptable error rates, and (c) the determination of sample size.
-I refer to conventional uses of NHST (i.e., choices that are widely shared across the literature, are made without justification, and/or refer to heuristics or conventions) as *defaults*.
-I refer to thoughtful uses of NHST (i.e., choices that are actively justified) as *active choices*.
-In addition to examining common practice, I provide a practical overview of active choices that can be made within the traditional framework to strengthen the practice of null-hypothesis significance testing.
-
-In the first section, I introduce default hypothesis tests and their thoughtful alternatives.
-In the second section, I conduct a literature review to examine the ways in which hypothesis tests are practiced in the psychological research literature.
-My goal is to assess the choices psychological scientists make when testing a hypothesis, the extent to which justifications are provided for these choices, and the extent to which choices are shared across the research literature (i.e., the prevalence of conventions).
-Of special interest are the choice of the null hypothesis (i.e., whether to test a nil-null or, for example, a range prediction), the choices of error rates (i.e., how the balance is struck between type I and type II error rates), and the determination of sample size.
-
-In the third section, I provide an in-depth, practical introduction to active choice hypothesis tests: a selection of thoughtful choices that can be made within the traditional framework to strengthen the practice of null-hypothesis significance testing.
-First, I consider the choice of null hypothesis.
-As an alternative to testing nil-null hypotheses, I introduce several approaches researchers can take to test range predictions.
-
-Second, I consider the choice of error rates.
-As an alternative to using conventional error thresholds, I discuss several approaches that researchers can take to choose and justify their error rates.
-Third, I consider sample size determination.
-As an alternative to using heuristics and unrealistic power estimates (based on effect sizes researchers deem plausible), I discuss several approaches that researchers can take to justify their sample size.
-In addition, I present a tool that allows researchers to efficiently perform well-powered experiments: sequential analysis.
-
-# Default vs. Active Choice Hypothesis Tests
-
-"Knowing what not to do with *p*-values is indeed necessary, but it does not suffice.
-It is as though statisticians were asking users of statistics to tear out the beams and struts holding up the edifice of modern scientific research without offering solid construction materials to replace them.
-Pointing out old, rotting timbers was a good start, but now we need more." [@wasserstein2019: p.
-1]
-
-## On the choice of the null hypothesis
-"This type of problem... is the basis of the tests of significance by which we can examine whether or not the data are in harmony with any suggested hypothesis." [@fisher1925: p. 8]
-
-The meaning of 'null' hypothesis is commonly misunderstood [@bakan1967].
-In the weakest construal of the NHST, the null hypothesis tested is a hypothesis of no difference [i.e., an effect or difference of exactly zero; @mayo2018]; this null is commonly referred to as the *nil-null* or the *straw-man* null hypothesis.
-However, the term null was originally used to refer to any hypothesis to be 'nullified' [@bakan1967; @mayo2018].
-
-The null can be any hypothesis a researcher wishes to test.
-According to Neyman [-@neyman1957: p.10], envisioning possible hypotheses that are relevant to the phenomenon of interest is the work of creative imagination.
-Many of the problems associated with NHST lessen if the null hypothesis is defined as something other than a hypothesized difference, effect, or correlation of precisely zero [@nickerson2000].
-
-A stronger construal of the NHST is one where the researcher tests a range null hypothesis rather than a nil-null hypothesis [@nickerson2000; @lakens2020a]: a move, referred to in the title, from a straw man to a steel woman.
-Instead of testing the nil-null hypothesis that the effect is exactly zero, researchers can test whether effects more extreme than a range of predicted or practically important effect sizes can be rejected.
-Examples of such tests are equivalence tests, minimum-effect tests, and inferiority tests [@lakens2020a; @lakens2018].
-In the third section, I will give a practical introduction to the use of range predictions in hypothesis tests.
-
-## On the choice of error rates
-
-"[N]o scientific worker has a fixed level of significance at which from year to year, and in all circumstances, he rejects hypotheses; he rather gives his mind to each particular case in the light of his evidence and his ideas." [@fisher1956: p.
-44-45]
-
-"From the point of view of mathematical theory all that we can do is to show how the risk of the errors may be controlled and minimized.
-The use of these statistical tools in any given case, in determining just how the balance should be struck, must be left to the investigator." [@neyman1933: p.293]
-
-Prior to conducting a Neyman-Pearson hypothesis test, researchers need to decide on a type I error rate (i.e., the long-run rate at which a true null hypothesis is rejected) and a type II error rate (i.e., given a number of participants, the long-run probability of failing to reject the null for a specified effect size of interest).
-In effect, a researcher needs to decide on the importance of exercising scientific caution (i.e., avoiding a type I error) versus the importance of discovery of an effect (i.e., avoiding a type II error).
-
-However, in practice, the sense of balancing the two types of error has been lost [@oakes1986; @rozeboom1960].
-More commonly, researchers tend to use conventional cut-offs for the two error rates ($\alpha$ =.05 and $\beta$ = .20, implying that type I errors are considered 4 times more costly than type II errors).
-In recent work, researchers have been urged to justify their $\alpha$ [@lakens2018].
-In the third section, I will elaborate on several approaches to the choice and justification of error rates.
-
-## On the choice of sample size
-
-"[I]t is an essential characteristic of experimentation that it is carried out with limited resources, and an essential part of the subject of experimental design to ascertain how these should be best applied." [@fisher1935: p.
-18]
-
-As early as 1962, the perils of conducting small-sample, low-power experiments were introduced to the psychological literature [@cohen1962].
-Unfortunately, more recent work has not been able to paint a rosier picture of the statistical power of hypothesis tests in psychological science [@bakker2012; @morey2016].
-
-A possible reason that hypothesis tests often have low chances of detecting meaningful effects is that researchers use heuristics (e.g., *N* = 50 per cell) or effect size estimates from previous studies (which may very well be inflated) when determining sample size [@lakens2017].
-Alternative approaches that have been suggested are safeguard power analysis [@perugini2014], expected power analysis [@biesanz2017], and powering the test to detect the smallest effect size of interest [@lakens2017].
-
-Researchers are often surprised to find out how many observations are typically required for a test to have appropriate power.
-Sequential analysis, on the other hand, allows researchers to perform well-powered experiments with smaller sample sizes [@wald1973].
-Thus, sequential analysis has the potential to significantly improve the practice of hypothesis testing in psychological science [@beffarabret2018; @lakens2014; @schnuerch2020].
-In the following section, I give a hands-on introduction to sequential hypothesis testing.
-
-## *Sequential Analysis*
-
-In 1929, Harold Dodge and Harry Romig introduced the idea of sequential analysis to the science of statistical quality control [@dodge1929].
-Sequential analysis is a method of statistical inference that allows data to be analyzed several times during its collection, with the goal of optimizing the amount of resources spent in research [@beffarabret2018].
-When Abraham Wald further developed this idea in the '40s, his work was deemed so useful to the United States military that it was classified as Restricted under the Espionage act [@wald1973].
-
-```{r, include = FALSE}
-gs = function(proc, return = "n") {
-  bs = ifelse(proc == "asOF", "bsOF", "bsP")
-
-  # Specify the design
-  design = getDesignGroupSequential(sided = 1,
-                                    alpha = 0.05, 
-                                    beta = 0.2,
-                                    kMax = 3, 
-                                    typeOfDesign = proc,
-                                    typeBetaSpending= bs)
-
-  # Get parameters
-  parameters = getSampleSizeMeans(design = design, groups = 2, alternative = 0.5)
-  n_gs = ceiling(parameters$maxNumberOfSubjects/3/2) # n per look per group
-  alpha = parameters$criticalValuesPValueScale
-  futility = parameters$futilityBoundsPValueScale %>% round(2)
-
-  if(return == "alpha") {return(alpha %>% signif(2) %>% as.character() %>% sub(".", "", .,))}
-  if(return == "n") {return(n_gs)}
-  if(return == "futility") {return(futility %>% as.character() %>% sub(".", "", .,))}
-}
-
-n_p = gs(proc = "asP")
-alpha_p = gs(proc = "asP", return = "alpha")
-futility_p = gs(proc = "asP", return = "futility")
-n_of = gs(proc = "asOF")
-alpha_of = gs(proc = "asOF", return = "alpha")
-futility_of = gs(proc = "asOF", return = "futility")
-```
-
-The most common procedure for sequential hypothesis testing is a group-sequential (GS) design [@proschan2006; @schonbrodt2017].
-The goal of GS procedures is to efficiently reject a null hypothesis in favor of an alternative hypothesis.
-In a prototypical group-sequential procedure [e.g., the Pocock correction or the O'Brien-Fleming correction, @pocock1977; @fleming1984], a researcher is allowed to look at the data several times during the course of a study.
-
-It is well-known that multiple looks at data, conventionally, increase the type I error rate [@simmons2011]: i.e., the long-run rate at which a true null hypothesis is rejected.
-Sequential analysis procedures, however, allow a researcher to take multiple looks at data without inflating the type I error rate.
-Critical $\alpha$ levels are adjusted for each look so that the overall type I error rate is controlled at the desired level (for instance, at an $\alpha$ of 5%).
-
-A general, sophisticated approach is provided by an "$\alpha$*-spending function*" [@lan1983; @wassmer2016].
-This approach allows a researcher to have unequally spaced looks at their data (for instance, after the first 20% and the first 75% of the maximum sample size) and to adapt the number of looks as the experiment continues.
-
-For example, a researcher who wishes to control their overall type I error rate at 5% and take *k* = 3 looks (i.e., two interim looks and one final look) at the data can use the common Pocock-like $\alpha$-spending function to determine what the corrected $\alpha$ level should be for each look (referred to as $\alpha_{1,k}$).
-In this case, $\alpha_{1,1:2}$ = `r alpha_p[1]` for the first two looks and $\alpha_{1,3}$ = `r alpha_p[3]` for the last look (i.e., for a one-tailed test).
-If during any of the interim looks $p \le \alpha_1$, the researcher can reject the null hypothesis and end the experiment; this is known as "*stopping for efficacy*".
-
-```{r, include = FALSE}
-n_fixed = ceiling(power.t.test(delta = 0.5, sig.level = .05, power = 0.8, 
-                                 type = "two.sample", alternative = "one.sided")$n)
-```
-
-Because the use of group-sequential procedures allows for early stopping, it reduces the average sample size needed.
-For example, if a researcher wants to conduct a one-sided, two-sample *t* test ($H_0:~\delta$ = 0 against $H_+:~\delta > 0$) with 80% power to detect a Cohen's *d* (i.e., a standardized mean difference) of 0.5, a conventional Null-Hypothesis Significance Test with an overall $\alpha$ of 5% would require `r n_fixed` participants per group.
-Using the Pocock-like GS design outlined above, the same study can be conducted with only `r n_p` participants per group per look.
-
-In addition to stopping early for efficacy, a researcher can "*stop for futility*" if the observed effect size is too small to warrant further investigation.
-This can be done by manually specifying a lower bound (for example, a lower bound at *d* $\le$ 0 or, equivalently, at *p* $\ge$ .50) or by using a "$\beta$*-spending function*" to determine the critical lower bound for each look (referred to as $\alpha_{0,k}$).
-
-The logic of the $\beta$-spending function is similar to that of the $\alpha$-spending function: critical values that warrant stopping for futility are determined a priori while controlling the overall type II error rate at the desired level (for instance, at a $\beta$ of 20%).
-In other words, the bounds are determined such that the probability of a type II error (i.e., given the number of subjects, the long-run probability of failing to reject the null when the true population effect $\delta$ = 0.5) is $\beta$ %.
-
-Figures 1A and 1B provide a visual overview of two possible group-sequential procedures (the Pocock-like and the O'Brien-Fleming-like designs) using $\alpha$- and $\beta$-spending functions, in the case of a one-tailed, two-sample *t*-test with $\alpha$ = 5%, $\beta$ = 20% for $\delta$ = 0.5, and $k_{max}$ = 3 looks.
-In our example of a Pocock-like design, the futility bounds are $\alpha_{0,1}$ = `r futility_p[1]` for the first look and $\alpha_{0,2}$ = `r futility_p[2]` for the second look.
-The researcher continues their study until the observed *p*-value crosses one of the prespecified $\alpha_1$ or $\alpha_0$ thresholds, or until the maximum *N* has been reached.
-
-Scientific experiments can require substantial time, money, and effort.
-At times, the well-being of research subjects, human or non-human, is at stake [@stefan2020].
-Under the Ethics Code of the American Psychological Association, psychological scientists have the ethical obligation to minimize research subjects' exposure to physical, emotional, or psychological harm [@apa2020].
-Sequential analysis can contribute to this goal, by allowing researchers to stop an experiment when the interim data show clear evidence that effects are positive, negative, or largely absent [@proschan2006].
-Sequential hypothesis testing procedures, it turns out, are highly valuable tools for psychological scientists to add to their statistical toolbox.
-
-\newpage
-
-# In-line code example
-
-```{r}
-n_fixed = ceiling(power.t.test(delta = 0.5, sig.level = .05, power = 0.8, 
-                                 type = "two.sample", alternative = "one.sided")$n)
-```
-
-For example, if a researcher wants to conduct a one-sided, two-sample *t* test ($H_0:~\delta$ = 0 against $H_+:~\delta > 0$) with 80% power to detect a Cohen's *d* (i.e., a standardized mean difference) of 0.5, a conventional Null-Hypothesis Significance Test with an overall $\alpha$ of 5% would require `r n_fixed` participants per group.
-
-\newpage
-
-# Example Table
-
-```{r}
-df = read.csv("MARP_data.csv")
-```
-
-```{r}
-mod1 = lm(wb_overall_mean ~ religiosity + age + gender + ses + education, data = df)
-apa_lm = apa_print(mod1)
-apa_table(apa_lm$table, caption = "Well-being regression table.")
-```
-
-The estimate for religiosity is `r apa_lm$estimate$religiosity`
-
-\newpage
-
-## Acknowledgement
-This manuscript was created using `r cite_r("r-references.bib")`.
-
-\newpage
-
-# References
-
-\begingroup
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
-
-<div id="refs" custom-style="Bibliography"></div>
-\endgroup
-
-\newpage
-
-## Figure 1A
-## *Pocock-like Group-Sequential Design: A worked example* 
-\noindent
-```{r, fig.heigth = 1, fig.width = 12}
-knitr::include_graphics("Figure1A.png", dpi = 1200)
-```
-
-*Note.* Pocock-like group-sequential design for a one-tailed, two-sample *t* test with $\alpha$ = .05, $\beta$ = .20 and $k_{max}$ = 3 looks.
-Critical values are determined on the basis of Pocock-like $\alpha$- and $\beta$-spending functions in *R* package *rpact* (Wassmer & Pahlke, 2020).
-
-\newpage
-
-## Figure 1B
-## *O'Brien-Fleming-like Group-Sequential Design: A worked example* 
-\noindent
-```{r, fig.heigth = 1, fig.width = 12}
-knitr::include_graphics("Figure1B.png", dpi = 1200)
-```
-
-*Note.* O'Brien-Fleming-like group-sequential design for a one-tailed, two-sample *t* test with $\alpha$ = .05, $\beta$ = .20 and $k_{max}$ = 3 looks.
-Critical values are determined on the basis of O'Brien-Fleming-like $\alpha$- and $\beta$-spending functions in *R* package *rpact* (Wassmer & Pahlke, 2020).
 

### 2022-04-27 rmarkdown-covid-19-transparent-analysis.Rmd
First draft

@@
 ```{r}
-pacman::p_load(tidyverse,
-               lubridate, 
-               here, 
+pacman::p_load(dplyr,
                ggplot2,
@@
                here,
+               tidyr,
+               europepmc,
+               lubridate,
+               ggpubr,
                epiR,
-               ggpubr,
+               gtsummary,
+               forcats,
+               nortest,
+               caret,
                tibble)
@@
 ```{r}
-kable(data.frame(hits_all = 175550, 
-           hits_oa = 133470,
-           oa_percentage = round((133470/175550)*100, 1)))
+kable(data.frame(hits_all = 175544, 
+           hits_oa = 133467,
+           oa_percentage = round((133467/175544)*100, 1)))
 ```
@@
 ```{r}
-c(number = 133470 - nrow(opendata), 
-  percentage = round((133470 - nrow(opendata))/133470*100, 1))
+c(number = 133467 - nrow(opendata), 
+  percentage = round((133467 - nrow(opendata))/133467*100, 1))
 ```
@@
 ```{r}
-length(table(opendata$journal))
+length(table(opendata$journalTitle))
 ```
@@
 ```{r}
-kable(head(table(opendata$journal) 
+kable(head(table(opendata$journalTitle) 
            %>% as.data.frame() 
@@
 opendata %>% 
-        mutate(journal = fct_lump_prop(journal, prop = 0.01) ) %>% 
+        mutate(journal = fct_lump_prop(journalTitle, prop = 0.01) ) %>% 
         count(journal) %>% 
@@
 kable(opendata[which.max(opendata$citedByCount),] %>% 
-  select(citedByCount, pmid, pmcid, doi.x, title, authorString, journalTitle, pubYear))
+  select(citedByCount, pmid, pmcid, doi, title, authorString, journalTitle, year_firstpub))
 ```
@@
 ```{r}
-library(epiR)
 
@@
 
-## Journal Impact Factor (JIF) analyses
-
-First, importing jif file:
+Sum of five indicators:
 ```{r}
-jif <- read.csv(here("data", "jif_2020.csv"))
+opendata %>% mutate(sum = rowSums(opendata %>% select(is_coi_pred, is_register_pred, is_fund_pred, is_open_data, is_open_code))) %>% select(sum) %>% tbl_summary()
 ```
 
-Merging it with opendata:
-```{r}
-db_jif <- merge(opendata, jif, by.x = "journalIssn", by.y = "ISSN")
-db_jif$X2020.JIF <- as.numeric(db_jif$X2020.JIF)
-```
+## Journal-related differences in transparency (Table 1)
 
@@
 ```{r}
-nrow(db_jif)
+table(is.na(opendata$jif2020))
 ```
 
-JIF based on COI disclosure:
+Five highest impact journals
 ```{r}
-kable(db_jif %>% 
-  group_by(is_coi_pred) %>% 
-  summarise(Median = round(median(X2020.JIF, na.rm = T), 1),
-            IQR = round(IQR(X2020.JIF, na.rm = T), 2)))
 
+tapply(opendata$jif2020, opendata$journalTitle, max) %>% as.data.frame() %>% arrange(desc(.)) %>% head(5)
+
+opendata$jif2020c = ifelse(opendata$jif2020 < 79.321,"Other", opendata$journalTitle)
 ```
 
-Citation based on COI disclosure:
+Table 1A
 ```{r}
-kable(db_jif %>% 
+set.seed(100)
+opendata %>% 
+  select(journalTitle,is_coi_pred,is_fund_pred,is_register_pred,is_open_data,is_open_code) %>%
+        mutate(journalTitle = fct_lump(journalTitle,n=5)) %>%
+  tbl_summary(by=journalTitle, percent="column",
+              label = c(is_coi_pred ~ "COI disclosure", is_fund_pred ~ "Funding disclosure", is_register_pred ~ "Protocol registration", is_open_data ~ "Data sharing", is_open_code ~ "Code sharing")) %>%
+  add_p(test.args=all_tests("fisher.test") ~ list(simulate.p.value=TRUE)) %>%
+        as_flex_table()
+```
+
+Table 1B
+```{r}
+set.seed(100)
+opendata %>% 
+  select(jif2020c,is_coi_pred,is_fund_pred,is_register_pred,is_open_data,is_open_code) %>%
+  tbl_summary(by=jif2020c, percent="column",
+              label = c(is_coi_pred ~ "COI disclosure", is_fund_pred ~ "Funding disclosure", is_register_pred ~ "Protocol registration", is_open_data ~ "Data sharing", is_open_code ~ "Code sharing")) %>%
+  add_p(test.args=all_tests("fisher.test") ~ list(simulate.p.value=TRUE)) %>%
+        as_flex_table()
+```
+
+## Publisher-related differences in transparency (Table 2)
+
+Table 2
+```{r}
+set.seed(100)
+opendata %>% 
+  select(scimago_publisher,is_coi_pred,is_fund_pred,is_register_pred,is_open_data,is_open_code) %>%
+        mutate(scimago_publisher = fct_lump(scimago_publisher,n=6)) %>%
+  tbl_summary(by=scimago_publisher, percent="column",
+              label = c(is_coi_pred ~ "COI disclosure", is_fund_pred ~ "Funding disclosure", is_register_pred ~ "Protocol registration", is_open_data ~ "Data sharing", is_open_code ~ "Code sharing")) %>%
+  add_p(test.args=all_tests("fisher.test") ~ list(simulate.p.value=TRUE)) %>%
+        as_flex_table()
+```
+
+## Journal Impact Factor (JIF) and citation count analyses (Table 3)
+
+Total number of papers published in journals with no JIF:
+```{r}
+sum(is.na(opendata$jif2020))
+```
+Normality of JIFs and citations
+```{r}
+hist(opendata$jif2020)
+hist(opendata$citedByCount)
+ad.test(opendata$jif2020)
+ad.test(opendata$citedByCount)
+```
+
+JIF and citations based on COI disclosure:
+```{r}
+kable(opendata %>% 
+  group_by(is_coi_pred) %>% 
+  summarise(Median = round(median(jif2020, na.rm = T), 1),
+            IQR = round(IQR(jif2020, na.rm = T), 2)))
+
+kable(opendata %>% 
   group_by(is_coi_pred) %>% 
@@
 
-Test for normality:
+Data was not normally distributed, hence we will use unpaired two-samples Wilcoxon test.
 ```{r}
-set.seed(10)
-
-with(db_jif, shapiro.test(sample(X2020.JIF[is_coi_pred == TRUE], 5000)))
-with(db_jif, shapiro.test(X2020.JIF[is_coi_pred == FALSE]))
-
-with(db_jif, shapiro.test(sample(citedByCount[is_coi_pred == TRUE], 5000)))
-with(db_jif, shapiro.test(citedByCount[is_coi_pred == FALSE]))
+wilcox.test(jif2020~is_coi_pred, data = opendata, exact = FALSE)
+wilcox.test(citedByCount~is_coi_pred, data = opendata, exact = FALSE)
 ```
 
-Data is not normal, hence we will use unpaired two-samples Wilcoxon test.
+JIF and citations based on funding disclosure:
 ```{r}
-wilcox.test(X2020.JIF~is_coi_pred, data = db_jif, exact = FALSE)
+kable(opendata %>% 
+  group_by(is_fund_pred) %>% 
+  summarise(Median = round(median(jif2020, na.rm = T), 1),
+            IQR = round(IQR(jif2020, na.rm = T), 2)))
 
-wilcox.test(citedByCount~is_coi_pred, data = db_jif, exact = FALSE)
-```
-
-JIF based on funding disclosure:
-```{r}
-kable(db_jif %>% 
-  group_by(is_fund_pred) %>% 
-  summarise(Median = round(median(X2020.JIF, na.rm = T), 1),
-            IQR = round(IQR(X2020.JIF, na.rm = T), 2)))
-
-```
-
-Citation based on funding disclosure:
-```{r}
-kable(db_jif %>% 
+kable(opendata %>% 
   group_by(is_fund_pred) %>% 
@@
 
-Test for normality:
+Data was not normally distributed, hence we will use unpaired two-samples Wilcoxon test.
 ```{r}
-set.seed(10)
-
-with(db_jif, shapiro.test(sample(X2020.JIF[is_fund_pred == TRUE], 5000)))
-with(db_jif, shapiro.test(sample(X2020.JIF[is_fund_pred == FALSE], 5000)))
-
-with(db_jif, shapiro.test(sample(citedByCount[is_fund_pred == TRUE], 5000)))
-with(db_jif, shapiro.test(sample(citedByCount[is_fund_pred == FALSE], 5000)))
+wilcox.test(jif2020~is_fund_pred, data = opendata, exact = FALSE)
+wilcox.test(citedByCount~is_fund_pred, data = opendata, exact = FALSE)
 ```
 
-Data is not normal, hence we will use unpaired two-samples Wilcoxon test.
+JIF and citations based on registration:
 ```{r}
-wilcox.test(X2020.JIF~is_fund_pred, data = db_jif, exact = FALSE)
+kable(opendata %>% 
+  group_by(is_register_pred) %>% 
+  summarise(Median = round(median(jif2020, na.rm = T), 1),
+            IQR = round(IQR(jif2020, na.rm = T), 2)))
 
-wilcox.test(citedByCount~is_fund_pred, data = db_jif, exact = FALSE)
-```
-
-JIF based on registration:
-```{r}
-kable(db_jif %>% 
-  group_by(is_register_pred) %>% 
-  summarise(Median = round(median(X2020.JIF, na.rm = T), 1),
-            IQR = round(IQR(X2020.JIF, na.rm = T), 2)))
-
-```
-
-Citation based on funding disclosure:
-```{r}
-kable(db_jif %>% 
+kable(opendata %>% 
   group_by(is_register_pred) %>% 
@@
 
-Test for normality:
+Data was not normally distributed, hence we will use unpaired two-samples Wilcoxon test.
 ```{r}
-set.seed(10)
-
-with(db_jif, shapiro.test(X2020.JIF[is_register_pred == TRUE]))
-with(db_jif, shapiro.test(sample(X2020.JIF[is_register_pred == FALSE], 5000)))
-
-with(db_jif, shapiro.test(citedByCount[is_register_pred == TRUE]))
-with(db_jif, shapiro.test(sample(citedByCount[is_register_pred == FALSE], 5000)))
+wilcox.test(jif2020~is_register_pred, data = opendata, exact = FALSE)
+wilcox.test(citedByCount~is_register_pred, data = opendata, exact = FALSE)
 ```
 
-Data is not normal, hence we will use unpaired two-samples Wilcoxon test.
+JIF and citations based on data sharing:
 ```{r}
-wilcox.test(X2020.JIF~is_register_pred, data = db_jif, exact = FALSE)
+kable(opendata %>% 
+  group_by(is_open_data) %>% 
+  summarise(Median = round(median(jif2020, na.rm = T), 1),
+            IQR = round(IQR(jif2020, na.rm = T), 2)))
 
-wilcox.test(citedByCount~is_register_pred, data = db_jif, exact = FALSE)
-```
-
-JIF based on data sharing:
-```{r}
-kable(db_jif %>% 
-  group_by(is_open_data) %>% 
-  summarise(Median = round(median(X2020.JIF, na.rm = T), 1),
-            IQR = round(IQR(X2020.JIF, na.rm = T), 2)))
-
-```
-
-Citation based on funding disclosure:
-```{r}
-kable(db_jif %>% 
+kable(opendata %>% 
   group_by(is_open_data) %>% 
@@
 
-Test for normality:
+Data was not normally distributed, hence we will use unpaired two-samples Wilcoxon test.
 ```{r}
-set.seed(10)
-
-with(db_jif, shapiro.test(sample(X2020.JIF[is_open_data == TRUE], 5000)))
-with(db_jif, shapiro.test(sample(X2020.JIF[is_open_data == FALSE], 5000)))
-
-with(db_jif, shapiro.test(sample(citedByCount[is_open_data == TRUE], 5000)))
-with(db_jif, shapiro.test(sample(citedByCount[is_open_data == FALSE], 5000)))
+wilcox.test(jif2020~is_open_data, data = opendata, exact = FALSE)
+wilcox.test(citedByCount~is_open_data, data = opendata, exact = FALSE)
 ```
 
-Data is not normal, hence we will use unpaired two-samples Wilcoxon test.
+JIF and citations based on code sharing:
 ```{r}
-wilcox.test(X2020.JIF~is_open_data, data = db_jif, exact = FALSE)
+kable(opendata %>% 
+  group_by(is_open_code) %>% 
+  summarise(Median = round(median(jif2020, na.rm = T), 1),
+            IQR = round(IQR(jif2020, na.rm = T), 2)))
 
-wilcox.test(citedByCount~is_open_data, data = db_jif, exact = FALSE)
-```
-
-JIF based on code sharing:
-```{r}
-kable(db_jif %>% 
-  group_by(is_open_code) %>% 
-  summarise(Median = round(median(X2020.JIF, na.rm = T), 1),
-            IQR = round(IQR(X2020.JIF, na.rm = T), 2)))
-
-```
-
-Citation based on funding disclosure:
-```{r}
-kable(db_jif %>% 
+kable(opendata %>% 
   group_by(is_open_code) %>% 
@@
 
-Test for normality:
+Data was not normally distributed, hence we will use unpaired two-samples Wilcoxon test.
 ```{r}
-set.seed(10)
-
-with(db_jif, shapiro.test(X2020.JIF[is_open_code == TRUE]))
-with(db_jif, shapiro.test(sample(X2020.JIF[is_open_code == FALSE], 5000)))
-
-with(db_jif, shapiro.test(citedByCount[is_open_code == TRUE]))
-with(db_jif, shapiro.test(sample(citedByCount[is_open_code == FALSE], 5000)))
-```
-
-Data is not normal, hence we will use unpaired two-samples Wilcoxon test.
-```{r}
-wilcox.test(X2020.JIF~is_open_code, data = db_jif, exact = FALSE)
-
-wilcox.test(citedByCount~is_open_code, data = db_jif, exact = FALSE)
-```
-
-## Sum of five indicators
-```{r}
-opendata <- opendata %>% mutate(sumOfIndicators = rowSums(opendata %>% select(is_coi_pred, is_register_pred, is_fund_pred, is_open_data, is_open_code)))
-```
-
-Number of papers with each number of TRUE indicators:
-```{r}
-c(five_ind = nrow(filter(opendata, sumOfIndicators == 5)),
-  four_ind = nrow(filter(opendata, sumOfIndicators == 4)),
-  three_ind = nrow(filter(opendata, sumOfIndicators == 3)),
-  two_ind = nrow(filter(opendata, sumOfIndicators == 2)),
-  one_ind = nrow(filter(opendata, sumOfIndicators == 1)),
-  zero_ind = nrow(filter(opendata, sumOfIndicators == 0)))
-```
-
-Percentage of papers with each number of TRUE indicators:
-```{r}
-
-c(five_ind = round(nrow(filter(opendata, sumOfIndicators == 5))/nrow(opendata)*100, 1),
-  four_ind = round(nrow(filter(opendata, sumOfIndicators == 4))/nrow(opendata)*100, 1),
-  three_ind = round(nrow(filter(opendata, sumOfIndicators == 3))/nrow(opendata)*100, 1),
-  two_ind = round(nrow(filter(opendata, sumOfIndicators == 2))/nrow(opendata)*100, 1),
-  one_ind = round(nrow(filter(opendata, sumOfIndicators == 1))/nrow(opendata)*100, 1),
-  zero_ind = round(nrow(filter(opendata, sumOfIndicators == 0))/nrow(opendata)*100, 1))
-```
-
-To know which of the indicators were TRUE in papers with two TRUE indicators:
-```{r}
-
-nrow_ <- nrow(filter(opendata, sumOfIndicators == 2))
-
-c(coi = round(nrow(filter(opendata, sumOfIndicators == 2 & is_coi_pred))/nrow_*100, 1),
-  fund = round(nrow(filter(opendata, sumOfIndicators == 2 & is_fund_pred))/nrow_*100, 1),
-  reg = round(nrow(filter(opendata, sumOfIndicators == 2 & is_register_pred))/nrow_*100, 1),
-  data = round(nrow(filter(opendata, sumOfIndicators == 2 & is_open_data))/nrow_*100, 1),
-  code = round(nrow(filter(opendata, sumOfIndicators == 2 & is_open_code))/nrow_*100, 1))
+wilcox.test(jif2020~is_open_code, data = opendata, exact = FALSE)
+wilcox.test(citedByCount~is_open_code, data = opendata, exact = FALSE)
 ```
@@
 ```{r}
-library(tibble)
 
@@
 ```{r}
-library(ggpubr)
 
@@
 ```
+
+# Appendix
+## Journals
+
+Total number of papers published in journals with an SJR:
+```{r}
+table(is.na(opendata$sjr2020))
+```
+
+Top five journals in SJR:
+```{r}
+tapply(opendata$sjr2020, opendata$journalTitle, max) %>% as.data.frame() %>% arrange(desc(.)) %>% head(5)
+
+opendata$sjr2020c = ifelse(opendata$sjr2020 < 26.304,"Other", opendata$journalTitle)
+```
+
+Table S1
+```{r}
+set.seed(100)
+opendata %>% 
+  select(sjr2020c,is_coi_pred,is_fund_pred,is_register_pred,is_open_data,is_open_code) %>%
+  tbl_summary(by=sjr2020c, percent="column",
+              label = c(is_coi_pred ~ "COI disclosure", is_fund_pred ~ "Funding disclosure", is_register_pred ~ "Protocol registration", is_open_data ~ "Data sharing", is_open_code ~ "Code sharing")) %>%
+  add_p(test.args=all_tests("fisher.test") ~ list(simulate.p.value=TRUE)) %>%
+        as_flex_table()
+```
+
+Top five journals in H-index:
+```{r}
+tapply(opendata$scimago_hindex, opendata$journalTitle, max) %>% as.data.frame() %>% arrange(desc(.)) %>% head(5)
+
+opendata$hindexc = ifelse(opendata$scimago_hindex < 771,"Other", opendata$journalTitle)
+```
+
+Table S2
+```{r}
+set.seed(100)
+opendata %>% 
+  select(hindexc,is_coi_pred,is_fund_pred,is_register_pred,is_open_data,is_open_code) %>%
+  tbl_summary(by=hindexc, percent="column",
+              label = c(is_coi_pred ~ "COI disclosure", is_fund_pred ~ "Funding disclosure", is_register_pred ~ "Protocol registration", is_open_data ~ "Data sharing", is_open_code ~ "Code sharing")) %>%
+  add_p(test.args=all_tests("fisher.test") ~ list(simulate.p.value=TRUE)) %>%
+        as_flex_table()
+```
+
+## Appendix 3:
+Correct p-values (considering all levels)
+```{r}
+opendata %>% 
+  select(journalTitle,is_coi_pred) %>%
+        table() %>%
+        fisher.test(simulate.p.value=TRUE)
+opendata %>% 
+  select(journalTitle,is_fund_pred) %>%
+        table() %>%
+        fisher.test(simulate.p.value=TRUE)
+opendata %>% 
+  select(journalTitle,is_register_pred) %>%
+        table() %>%
+        fisher.test(simulate.p.value=TRUE)
+opendata %>% 
+  select(journalTitle,is_open_data) %>%
+        table() %>%
+        fisher.test(simulate.p.value=TRUE)
+opendata %>% 
+  select(journalTitle,is_open_code) %>%
+        table() %>%
+        fisher.test(simulate.p.value=TRUE)
+```
+
+Code sharing for all journals
+```{r}
+set.seed(100)
+opendata %>% 
+  select(journalTitle,is_open_code) %>%
+  tbl_summary(by=is_open_code, percent="row", sort = list(everything() ~ "frequency"))%>%
+  add_p(test.args=all_tests("fisher.test") ~ list(simulate.p.value=TRUE)) %>% 
+        as_flex_table()
+```
+
+Data sharing for all journals
+```{r}
+set.seed(100)
+opendata %>% 
+  select(journalTitle,is_open_data) %>%
+  tbl_summary(by=is_open_data, percent="row", sort = list(everything() ~ "frequency"))%>%
+  add_p(test.args=all_tests("fisher.test") ~ list(simulate.p.value=TRUE))%>% 
+        as_flex_table()
+```
+
+Protocol registration for all journals
+```{r}
+set.seed(100)
+opendata %>% 
+  select(journalTitle,is_register_pred) %>%
+  tbl_summary(by=is_register_pred, percent="row", sort = list(everything() ~ "frequency"))%>%
+  add_p(test.args=all_tests("fisher.test") ~ list(simulate.p.value=TRUE)) %>% 
+        as_flex_table()
+```
+
+COI disclosure for all journals
+```{r}
+set.seed(100)
+opendata %>% 
+  select(journalTitle,is_coi_pred) %>%
+  tbl_summary(by=is_coi_pred, percent="row", sort = list(everything() ~ "frequency"))%>%
+  add_p(test.args=all_tests("fisher.test") ~ list(simulate.p.value=TRUE))%>% 
+        as_flex_table()
+```
+
+Funding disclosure for all journals
+```{r}
+set.seed(100)
+opendata %>% 
+  select(journalTitle,is_fund_pred) %>%
+  tbl_summary(by=is_fund_pred, percent="row", sort = list(everything() ~ "frequency"))%>%
+  add_p(test.args=all_tests("fisher.test") ~ list(simulate.p.value=TRUE))%>% 
+        as_flex_table()
+```
+
+# Publisher-related differences in transparency
+Correct p-values (considering all levels)
+```{r}
+opendata %>% 
+  select(scimago_publisher,is_coi_pred) %>%
+        table() %>%
+        fisher.test(simulate.p.value=TRUE)
+opendata %>% 
+  select(scimago_publisher,is_fund_pred) %>%
+        table() %>%
+        fisher.test(simulate.p.value=TRUE)
+opendata %>% 
+  select(scimago_publisher,is_register_pred) %>%
+        table() %>%
+        fisher.test(simulate.p.value=TRUE)
+opendata %>% 
+  select(scimago_publisher,is_open_data) %>%
+        table() %>%
+        fisher.test(simulate.p.value=TRUE)
+opendata %>% 
+  select(scimago_publisher,is_open_code) %>%
+        table() %>%
+        fisher.test(simulate.p.value=TRUE)
+```
+
+Data sharing for all publishers
+```{r}
+set.seed(100)
+opendata %>% 
+select(scimago_publisher,is_open_data) %>%
+tbl_summary(by=is_open_data, percent="row", sort = list(everything() ~ "frequency"))%>%
+ add_p(test.args=all_tests("fisher.test") ~ list(simulate.p.value=TRUE))%>% 
+        as_flex_table()
+```
+
+Protocol registeration in all publishers
+```{r}
+set.seed(100)
+opendata %>% 
+select(scimago_publisher,is_register_pred) %>%
+tbl_summary(by=is_register_pred, percent="row", sort = list(everything() ~ "frequency"))%>%
+  add_p(test.args=all_tests("fisher.test") ~ list(simulate.p.value=TRUE))%>% 
+        as_flex_table()
+```
+
+COI disclosure in all publishers
+```{r}
+set.seed(100)
+opendata %>% 
+select(scimago_publisher,is_coi_pred) %>%
+tbl_summary(by=is_coi_pred, percent="row", sort = list(everything() ~ "frequency")) %>%
+  add_p(test.args=all_tests("fisher.test") ~ list(simulate.p.value=TRUE))%>% 
+        as_flex_table()
+```
+
+Funding disclosure in all publishers
+```{r}
+set.seed(100)
+opendata %>% 
+select(scimago_publisher,is_fund_pred) %>%
+tbl_summary(by=is_fund_pred, percent="row", sort = list(everything() ~ "frequency")) %>%
+add_p(test.args=all_tests("fisher.test") ~ list(simulate.p.value=TRUE))%>% 
+        as_flex_table()
+```

### 2022-04-27 r_covid_transparency_dataset_creation.R

@@ -1,0 +1,297 @@
+# Creating dataset and xml files for the research project:
+## Transparency indicators across the dentistry and oral health literature
+## Authors: Ahmad Sofi-Mahmudi, Eero Raittio, KSergio E. Uribe
+## Codes by: Ahmad Sofi-Mahmudi
+
+# Loading required packages
+pacman::p_load(dplyr,
+               rtransparent, 
+               metareadr, 
+               europepmc,
+               here)
+
+# Creating the dataset
+## Getting desired COVID-19 keywords
+COVID_19_keywords <- read.csv(here("data", "covid_transparency_keywords.csv"))
+
+### Some modifications are needed to create proper keywords for Title, Keywords
+### and Results:
+COVID_19_keywords <- COVID_19_keywords %>% mutate(
+        search.term.title = paste0("TITLE:", '"', Keyword, '"'),
+        search.term.keyword = paste0("KW:", '"', Keyword, '"'),
+        search.term.results = paste0("RESULTS:", '"', Keyword, '"'))
+
+TitleQuery <- paste(COVID_19_keywords$search.term.title,
+                    collapse = " OR ")
+
+KeywordQuery <- paste(COVID_19_keywords$search.term.keyword,
+                      collapse = " OR ")
+
+ResultQuery <- paste(COVID_19_keywords$search.term.results,
+                     collapse = " OR ")
+
+### Getting all open-access and EPMC COVID-19-related papers indexed in PubMed
+
+#### As in this step we should download a large file that includes the 
+#### characteristics of papers of our query, to ease the process, we divided our 
+#### query to five separate queries based on the date: 1- first half of 2020;
+#### 2- second half of 2020; 3- first half of 2021; 4- second half of 2021;
+#### 5- the first two months of 2022.
+
+db_2020_1 <- epmc_search(
+        query = paste0("(", TitleQuery, " OR ", 
+                       KeywordQuery, " OR ", 
+                       ResultQuery, ") ", 
+                       'AND (SRC:"MED") 
+                       AND (LANG:"eng" OR LANG:"en" OR LANG:"us") 
+                       AND (FIRST_PDATE:[2020-01-01 TO 2020-06-30]) 
+                       AND (OPEN_ACCESS:y) 
+                       AND (PUB_TYPE:"Journal Article" 
+                            OR PUB_TYPE:"research-article" 
[…]
